### 2021-04-22 outline/outline.tex
first commit

% Created 2021-04-22 Thu 10:45
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[version=3]{mhchem}
\usepackage[numbers,super,sort&compress]{natbib}
\usepackage{natmove}
\usepackage{url}
\usepackage{minted}
\usepackage{underscore}
\usepackage[linktocpage,pdfstartview=FitH,colorlinks,
linkcolor=blue,anchorcolor=blue,
citecolor=blue,filecolor=blue,menucolor=blue,urlcolor=blue]{hyperref}
\usepackage{attachfile}
\usepackage{geometry}
\geometry{margin=1.0in,top=.75in,bottom=.75in}
\usepackage{graphicx}
\usepackage{color}
\usepackage[numbers,super,sort&compress]{natbib}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{font=footnotesize}
\usepackage[version=3]{mhchem}
\usepackage{siunitx}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{mdwlist}
\usepackage{hyperref}
\pagestyle{fancy}
\usepackage{wrapfig}
\usepackage{nopageno}
\fancyhf{}
\fancyhead[LE,RO]{\scriptsize Jerry Crum}
\fancyhead[RE,LO]{\scriptsize ZSE Outline}
%\fancyfoot[CE,CO]{\leftmark}
\fancyfoot[LE,RO]{\thepage}
%\usepackage{subfig}
\usepackage{comment}
\usepackage{titlesec}
\titlespacing*{\section}
{0pt}{0.6\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsection}
{0pt}{0.6\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsubsection}
{0pt}{0.4\baselineskip}{0.1\baselineskip}
\usepackage{parskip}
\usepackage[section]{placeins}
\usepackage{siunitx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\orange}[1]{\textcolor{orange}{#1}}
\usepackage[capitalise]{cleveref}
\author{Jerry T. Crum\(^{\text{a}}\), Justin R. Crum\(^{\text{b}}\), William F. Schneider\(^{\text{a,c}}\)}
\date{}
\title{ZSE Manuscript Outline}
\begin{document}

\begin{OPTIONS}
\def\udesoftecoverride\#1\mainmatter\{\%
  \AfterEndPreamble{#1\mainmatter}
\end{OPTIONS}

\maketitle

\begin{asparaenum}[\expandafter\textsuperscript a ]
\item  Department of Chemical and Biolmolecular Engineering, University of Notre Dame, 250 Nieuwland Science Hall, Notre Dame, IN 46556, USA \\
\item Justin's info \\
\item Department of Chemistry and Biochmeistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
\end{asparaenum}

\newpage
\section*{Introduction}
\label{sec:org9460a91}
\begin{itemize}
\item Topology of zeolite frameworks and associated tetrahedral sites (T-sites) are commonly characterized by their associated rings
\item Rings are defined as a close cycle traversing the T-site and oxygen atoms of the framework, and cannot be decomposed into smaller cycles by a shortcut \cite{goetzke-properties-1991}.
\item A shortcut is defined as a path connecting two nodes of a cycle that is shorter than both the paths connecting those nodes along the cycle.
\item See \Cref{fig:cha-labeled} for example of rings
\begin{itemize}
\item The blue highlighted cycle is a 6-membered ring (6-MR)
\item The green highlighted cycle is a 4-MR
\item The cycle following 1-2-3-4-5-6-7-8 is not an 8-MR because there is a shortcut connecting nodes 3 and 8.
\end{itemize}
\end{itemize}

\begin{center}
\includegraphics[width=0.60\textwidth]{../figures/cha-labeled.pdf}
\captionof{figure}{Cutout of the Chabazite framework showing a 6-MR in blue and a 4-MR in green. Yellow atoms are Si (T-sites), and red atoms are oxygen. \label{fig:cha-labeled}}
\end{center}

\begin{itemize}
\item Rings have been used to identify feasible zeolite topologies \cite{li-why-2014}, to descirbe the similarity and differences between zeolites \cite{curtis-statistical-2003}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints [will get citations for this]

\item Different methods exist to count rings present in a zeolite
\item These methods return different sets of rings
\item Vertex symbols are the set of shortest paths connecting the 6 oxygen-oxygen pairs around a T-site \cite{okeeffe-vertex-1997}
\item Shortest path rings count all the vertex symbol rings that pass through a T-site or an oxygen atom \cite{sastre-topological-2009}
\item Or we can count all the rings that do not have a short cut \cite{goetzke-properties-1991}
\item Differences in ring counts leads to differences in how we describe the topology of zeolites. Therefore, when discussing the rings in a zeolite it is important to also state which method of ring counting is used.
\item Here we report an analysis of rings and T-sites in a large number of zeolite frameworks using Zeolite Simulation Environment, a Python package that implements an efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} for finding rings in arbitrary frameworks.
\item We compare the results of a number of common and new ring definitions applied to a large number of common zeolite frameworks.\cite{li-first-principles-2018}
\end{itemize}

\section*{Software Description}
\label{sec:org9ec0fdb}

\begin{itemize}
\item All of the frameworks listed on the IZA Database of zeolite structures \cite{IZA} are included in a database with ZSE
\item These structures are Atomic Simulation Environment Atoms objects \cite{larsen-atomic-2017}, and can be used with any of the functions in ZSE
\item ZSE also includes CIF tools to read structure files for frameworks not listed in the IZA website, such as hypothetical zeolites, and return an Atoms object that can be used with ZSE
\item ZSE has 3 previously published rules for ring finding implemented
\begin{itemize}
\item All cycles with out a shortcut \cite{goetzke-properties-1991}
\item All shortest path cycles \cite{sastre-topological-2009}
\item Cycles that compose the vertex symbol for a T-site \cite{okeeffe-vertex-1997}
\end{itemize}
\item We have also implemented a new rule that finds all rings with out a shortcut, but excludes rings that are made by traversing a stacked set of rings.
\begin{itemize}
\item Figure showing example of 8-MR in the d6r of CHA and 14-MR in AFI
\end{itemize}
\item Each of the rules: shortest path, vertex symbols, and our new rule are a subset of the no shortcuts rule
\end{itemize}
Process to find rings:
\begin{enumerate}
\item To find rings in a zeolite, ZSE makes a custom connectivity matrix for the Si and O atoms in the framework
\item We use NetworkX \cite{SciPyProceedings-11} to build a shortest path matrix for every atom pair in the zeolite framework
\item We then find all the rings up to some cutoff size base on the algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991}
\item Depending on the rule chosen by the user, ZSE then removes rings from this list that don't meet the qualifications of the rule
\item ZSE returns a list of the rings found, a list of the atom indicies that compose each ring, Atoms objects for each ring that can be further analyzed or visualized by the user
\end{enumerate}

\section*{Results}
\label{sec:org51ff3f9}

\begin{itemize}
\item Reproduce the results from Sastre paper, show ring counts with the other rules
\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature
\item Show the pentasil 6-membered cycle (MFI) and discuss how that is often considered a ring by catalysis researchers, but not a ring by the Goetzke definition
\end{itemize}

\section*{Conclusions}
\label{sec:orga1f2e62}

\begin{itemize}
\item The method used to find rings in a zeolite will provide different ring counts
\item When discussing rings in a zeolite it is import to disclose by which method those rings were found
\item Using ZSE we can find rings based on various methods
\item This provides a foundation for using ring fingerprints in machine learning models to correlate chemical properties and topology
\end{itemize}

\bibliographystyle{unsrtnat}
\bibliography{ref}
\end{document}

### 2021-04-22 outline/fix_ref.py

@@ -1,0 +1,13 @@
+file = open('ref.bib',"r")
+lines = file.readlines()
+file.close()
+for i,line in enumerate(lines):
+    if 'article' in line:
+        if '_' in line:
+            line = line.replace('_','-')
+            print(line)
+        lines[i]=line
+
+file = open('ref.bib',"w")
+file.writelines(lines)
+file.close()

### 2021-04-26 outline/outline.tex
added table of ring counts

@@
-% Created 2021-04-22 Thu 10:45
+% Created 2021-04-26 Mon 08:56
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org9460a91}
+\label{sec:orgd9d2283}
 \begin{itemize}
@@
 \begin{itemize}
-\item The blue highlighted cycle is a 6-membered ring (6-MR)
-\item The green highlighted cycle is a 4-MR
+\item The blue highlighted cycle (8-3-4-5-6-7) is a 6-membered ring (6-MR)
+\item The green highlighted cycle (1-2-3-8) is a 4-MR
 \item The cycle following 1-2-3-4-5-6-7-8 is not an 8-MR because there is a shortcut connecting nodes 3 and 8.
@@
 \section*{Software Description}
-\label{sec:org9ec0fdb}
+\label{sec:orgfc05f55}
 
@@
 \section*{Results}
-\label{sec:org51ff3f9}
+\label{sec:orgded8dc4}
 
@@
 
+\subsection*{Comparison of rings found for the T-site in various uninodal zeolite frameworks}
+\label{sec:orgd5ed0c9}
+
+\begin{center}
+\begin{tabular}{llll}
+Framework & Goetzke & Crum & Sastre \cite{sastre-topological-2009}\\
+\hline
+ABW & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+ACO & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}10\(_{\text{15}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\\
+AFI & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\)\textbullet{}14\(_{\text{7}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\\
+ANA & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\)\\
+ATO & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}8\(_{\text{8}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\\
+BCT & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{12}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+DFT & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\textbullet{}10\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\\
+GIS & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+MER & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\textbullet{}10\(_{\text{10}}\)\textbullet{}14\(_{\text{14}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+MON & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\)\\
+NPO & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+\end{tabular}
+\end{center}
+
 \section*{Conclusions}
-\label{sec:orga1f2e62}
+\label{sec:orgbd83cd3}
 

### 2021-04-26 outline/outline.tex
updated document

--- 2021-04-22 outline/outline.tex
@@
-% Created 2021-04-22 Thu 10:45
+% Created 2021-04-26 Mon 14:20
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org9460a91}
+\label{sec:orgd1ecbe6}
 \begin{itemize}
@@
 \begin{center}
-\includegraphics[width=0.60\textwidth]{../figures/cha-labeled.pdf}
+\includegraphics[width=0.60\textwidth]{../figures/completed-figures/cha-labeled.pdf}
 \captionof{figure}{Cutout of the Chabazite framework showing a 6-MR in blue and a 4-MR in green. Yellow atoms are Si (T-sites), and red atoms are oxygen. \label{fig:cha-labeled}}
@@
 \section*{Software Description}
-\label{sec:org9ec0fdb}
+\label{sec:orgef01691}
 
@@
 \section*{Results}
-\label{sec:org51ff3f9}
+\label{sec:org031b217}
 
 \begin{itemize}
-\item Reproduce the results from Sastre paper, show ring counts with the other rules
-\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature
+\item Reproduce the results from Sastre paper, show ring counts with the other rules, \Cref{tab:ring-counts}
+\begin{itemize}
+\item Results in the Sastre column were found using ZSE but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}
+\end{itemize}
+\end{itemize}
+
+\begin{table}[htbp]
+\caption{Comparison of Ring Indices for the T-sites in Various Uninodal Zeolite Frameworks \label{tab:ring-counts}}
+\centering
+\begin{tabular}{llll}
+Framework & Goetzke & Crum & Sastre \cite{sastre-topological-2009}\\
+\hline
+ABW & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+ACO & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}10\(_{\text{15}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\\
+AFI & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\)\textbullet{}14\(_{\text{7}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\\
+ANA & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\)\\
+ATO & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}8\(_{\text{8}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\\
+BCT & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{12}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+CHA & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
+DFT & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\textbullet{}10\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\\
+GIS & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+GME & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{7}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
+MER & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\textbullet{}10\(_{\text{10}}\)\textbullet{}14\(_{\text{14}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+MON & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\)\\
+NPO & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+\end{tabular}
+\end{table}
+
+\begin{itemize}
+\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
+\begin{itemize}
+\item Looking at results from CHA we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
+\item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
+\item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
+\item The 12-MR is a cycle that circumferences the CHA cage
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
+\captionof{figure}{Chabazite framework with highlighted rings: 4-MR in green, 6-MR in blue, 8-MR in pink, 12-MR in purple.
+The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature.
+This is a placeholder figure, I want to show the rings in CHA for discussion.
+Hoping to get your input on how best to do that.
+\label{fig:cha-rings}}
+\end{center}
+
+\begin{itemize}
 \item Show the pentasil 6-membered cycle (MFI) and discuss how that is often considered a ring by catalysis researchers, but not a ring by the Goetzke definition
@@
 \section*{Conclusions}
-\label{sec:orga1f2e62}
+\label{sec:orgd0b0537}
 

### 2021-04-27 outline/outline.tex
added ring count plot

@@
-% Created 2021-04-26 Mon 14:20
+% Created 2021-04-27 Tue 08:48
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:orgd1ecbe6}
+\label{sec:org7019896}
 \begin{itemize}
@@
 \section*{Software Description}
-\label{sec:orgef01691}
+\label{sec:org37b98dd}
 
@@
 \section*{Results}
-\label{sec:org031b217}
+\label{sec:org0ab2fb0}
+\begin{itemize}
+\item Ring counts frequency plots
+\begin{itemize}
+\item Plot showing how many frameworks on the IZA contain each size ring found using the various ring counting methods
+\item This highlights the differences in the ring rules, and shows that results will vary depending on rule.
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/ring-counts.pdf}
+\captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
+[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website] \label{fig:ring-counts}}
+\end{center}
 
@@
 \item Results in the Sastre column were found using ZSE but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}
+\item This provides an in depth look at some of the frameworks and the differences in rings found by each rule.
+\item Leads into the next section discussing the specific rings of CHA and pentasil that do or don't get counted by each rule.
 \end{itemize}
@@
 \section*{Conclusions}
-\label{sec:orgd0b0537}
+\label{sec:org567b455}
 

### 2021-04-28 outline/outline.tex
added some new T site analysis

@@
-% Created 2021-04-27 Tue 08:48
+% Created 2021-04-27 Tue 16:07
 % Intended LaTeX compiler: pdflatex
@@
 \item Department of Chemical and Biolmolecular Engineering, University of Notre Dame, 250 Nieuwland Science Hall, Notre Dame, IN 46556, USA \\
-\item Justin's info \\
+\item Department of Applied Mathematics, University of Arizona, 617 N Santa Rita Ave, Tucson, AZ 85721, USA\\
 \item Department of Chemistry and Biochmeistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
@@
 \section*{Introduction}
-\label{sec:org7019896}
+\label{sec:org18dc73e}
 \begin{itemize}
@@
 \section*{Software Description}
-\label{sec:org37b98dd}
+\label{sec:orgcf7753e}
 
@@
 \section*{Results}
-\label{sec:org0ab2fb0}
+\label{sec:orgcb7d5e9}
 \begin{itemize}
@@
 \begin{itemize}
+\item Number of unique T-sites
+\begin{itemize}
+\item There are 1460 T-sites through all the frameworks listed on the IZA website.
+\item We can characterize those T-sites by the rings that pass through them
+\item Sastre did this, and called the list of rings, the ring index
+\item If we do this using different rules for ring finding how do the results change?
+\begin{itemize}
+\item See \Cref{fig:unique-ts}
+\end{itemize}
+\item Most common T-site ring index using Goetzke method is: 5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) showing up 23 times through the IZA frameworks.
+\item Most common T-site ring index using Crum method is: 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) showing up 31 times through the IZA frameworks.
+\begin{itemize}
+\item Next most common T-site with Crum method is 5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) showing up 25 times
+\end{itemize}
+\item This raises the question, if you want to use machine learning to correlate T-site rings to chemical properties, which ring method should you use?
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/unique-ts.pdf}
+\captionof{figure}{Number of unique T-sites when classified by the rings passing through them using varrious ring finding rules.
+[Sastre and Vertex method to be updated, still collecting data] \label{fig:unique-ts}}
+\end{center}
+
+\begin{itemize}
+\item Number of unique oxygen sites
+\begin{itemize}
+\item We can repeat this method for the oxygen atoms in all the frameworks
+\item There are 3219 oxygen atoms through all the frameworks listed on the IZA website
+\item Figure here showing counts based on ring finding rules
+\end{itemize}
+
 \item Reproduce the results from Sastre paper, show ring counts with the other rules, \Cref{tab:ring-counts}
@@
 \section*{Conclusions}
-\label{sec:org567b455}
+\label{sec:org0fcf21a}
 

### 2021-05-03 outline/outline.tex
updated some data
added some new plots

@@
-% Created 2021-04-27 Tue 16:07
+% Created 2021-05-03 Mon 09:24
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org18dc73e}
+\label{sec:org4471833}
 \begin{itemize}
@@
 \section*{Software Description}
-\label{sec:orgcf7753e}
+\label{sec:orgb60a0df}
 
@@
 \section*{Results}
-\label{sec:orgcb7d5e9}
+\label{sec:org9070503}
 \begin{itemize}
@@
 \captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
-[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website] \label{fig:ring-counts}}
+[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website.
+Currently the IZA does not show any ring data for the SVY framework, providing one less framework to count.
+\label{fig:ring-counts}}
 \end{center}
@@
 \includegraphics[width=.6\textwidth]{../figures/completed-figures/unique-ts.pdf}
-\captionof{figure}{Number of unique T-sites when classified by the rings passing through them using varrious ring finding rules.
-[Sastre and Vertex method to be updated, still collecting data] \label{fig:unique-ts}}
+\captionof{figure}{Number of unique T-sites when classified by the rings passing through them using varrious ring finding rules. \label{fig:unique-ts}}
 \end{center}
@@
 \item We can repeat this method for the oxygen atoms in all the frameworks
-\item There are 3219 oxygen atoms through all the frameworks listed on the IZA website
-\item Figure here showing counts based on ring finding rules
+\item Counting the symmetry distinct oxygen atoms in each framework on the IZA database leads to a total count of 3219
+\item We can classify those oxygen atoms based on the rings that pass through them, using the various ring counting rules
+\item \Cref{fig:unique-os} shows counts based on ring finding rules
+\item The percentage of unique oxygen sites is much lower than the percentage of unique T-sites for every ring finding method
+\end{itemize}
 \end{itemize}
 
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/unique-os.pdf}
+\captionof{figure}{Number of unique oxygen sites when classified by the rings passing through them using varrious ring finding rules.
+Vertex method not included, since that is a way to classify T-sites only.
+\label{fig:unique-os}}
+\end{center}
+
+\begin{itemize}
 \item Reproduce the results from Sastre paper, show ring counts with the other rules, \Cref{tab:ring-counts}
@@
 \section*{Conclusions}
-\label{sec:org0fcf21a}
+\label{sec:org67648fd}
 

### 2021-05-04 outline/outline.tex
added ring definition matrix

@@
-% Created 2021-05-03 Mon 09:24
+% Created 2021-05-04 Tue 08:36
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org4471833}
+\label{sec:orgb2744a9}
 \begin{itemize}
@@
 \end{center}
+\newpage
+
+\begin{longtable}{l p{2.7cm} p{2.7cm} p{2.7cm} p{2.7cm}}
+\caption{Matrix showing relationship between frameoworks, nodes, paths, cycles, and various ring types. \label{tab:ring-def}}
+\\
+ & Description & Framework & Node (T-Sites) & Node (Oxygen)\\
+\hline
+\endfirsthead
+\multicolumn{5}{l}{Continued from previous page} \\
+\hline
+
+ & Description & Framework & Node (T-Sites) & Node (Oxygen) \\
+
+\hline
+\endhead
+\hline\multicolumn{5}{r}{Continued on next page} \\
+\endfoot
+\endlastfoot
+\hline
+Nodes & T-sites and oxygen atoms & Contains some set of symmetry distinct T-sites and oxygen atoms &  & \\
+Paths & Collection of connected nodes from source to target & Periodic cell contains an infinite number of paths &  & \\
+Cycles & Path that starts and ends at the same node & Periodic cell Contains an infinite number of cycles &  & \\
+Rings & Cycle that contains no shortcuts & Contains a finite number of unique rings & All rings that pass through particular T-site & All rings that pass through particular oxygen atom\\
+Unstacked Rings & Ring that does not traverse two stacked rings & A subset of the Rings above & All unstacked rings that pass through T-site & All unstacked rings that pass through oxygen atom\\
+Shortest Path Rings & Ring that is the shortest ring from at least one set of O-T-O on the cycle & A smaller subset of the rings above & All shortest path rings that pass through T-site & All shortest path rings that pass through oxygen atom\\
+Vertex Symbol & Way to classify the rings around a T-site, shortest ring (and its multiplicity) for each O-O pair around a T-site & Collection of vertex symbols for all symmetry distinct T-sites in framework & Vertex symbol for particular T-site & \\
+\end{longtable}
 
@@
 \section*{Software Description}
-\label{sec:orgb60a0df}
+\label{sec:org5d41ccb}
 
@@
 \section*{Results}
-\label{sec:org9070503}
+\label{sec:orgf798bf0}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:org67648fd}
+\label{sec:orgf3f754a}
 

### 2021-05-05 outline/outline.tex
updated figure 1, add graph theory in intro

@@
-% Created 2021-05-04 Tue 08:36
+% Created 2021-05-05 Wed 14:54
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:orgb2744a9}
+\label{sec:org8d16b16}
 \begin{itemize}
@@
 \begin{itemize}
-\item The blue highlighted cycle (8-3-4-5-6-7) is a 6-membered ring (6-MR)
-\item The green highlighted cycle (1-2-3-8) is a 4-MR
-\item The cycle following 1-2-3-4-5-6-7-8 is not an 8-MR because there is a shortcut connecting nodes 3 and 8.
+\item The pink highlighted cycle (1-2-3-17-20-14-15-16) is a 8-membered ring (8-MR)
+\item The green highlighted cycle (14-20-12-13) is a 4-MR
+\item The red outlined cycle following 3-4-18-19-20-17 is not an 6-MR because there is a shortcut connecting nodes 17 and 18.
+\item Nodes 5-6-7-8-9 outlined in teal represent a path through the framework.
 \end{itemize}
@@
 \begin{center}
-\includegraphics[width=0.60\textwidth]{../figures/completed-figures/cha-labeled.pdf}
-\captionof{figure}{Cutout of the Chabazite framework showing a 6-MR in blue and a 4-MR in green. Yellow atoms are Si (T-sites), and red atoms are oxygen. \label{fig:cha-labeled}}
+\includegraphics[width=0.60\textwidth]{../figures/completed-figures/ring-examples.pdf}
+\captionof{figure}{Cutout of the Chabazite framework showing a path from node 3 to node 9 outlined in teal, a cycle (3-4-18-19-20-17) outlined in red, an 8-MR in pink, and a 4-MR in green.
+Yellow atoms are Si (T-sites), and red atoms are oxygen.
+\label{fig:cha-labeled}}
 \end{center}
@@
 \section*{Software Description}
-\label{sec:org5d41ccb}
+\label{sec:org31f0468}
 
@@
 \section*{Results}
-\label{sec:orgf798bf0}
+\label{sec:orgc227a60}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:orgf3f754a}
+\label{sec:org6e9315d}
 

### 2021-05-06 outline/outline.tex
new figures, more text

@@
-% Created 2021-05-05 Wed 14:54
+% Created 2021-05-06 Thu 16:34
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org8d16b16}
+\label{sec:org061dd26}
 \begin{itemize}
+\item Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges.
+\begin{itemize}
+\item Sometimes in the literature, oxygen atoms are used as bridges instead of nodes, since they only connect to two T-sites
+\end{itemize}
 \item Topology of zeolite frameworks and associated tetrahedral sites (T-sites) are commonly characterized by their associated rings
@@
 \section*{Software Description}
-\label{sec:org31f0468}
+\label{sec:org71ffeda}
 
 \begin{itemize}
-\item All of the frameworks listed on the IZA Database of zeolite structures \cite{IZA} are included in a database with ZSE
+\item All of the frameworks listed on the IZA Database of zeolite structures \cite{baerlocher-database-nodate} are included in a database with ZSE
 \item These structures are Atomic Simulation Environment Atoms objects \cite{larsen-atomic-2017}, and can be used with any of the functions in ZSE
@@
 \item To find rings in a zeolite, ZSE makes a custom connectivity matrix for the Si and O atoms in the framework
-\item We use NetworkX \cite{SciPyProceedings-11} to build a shortest path matrix for every atom pair in the zeolite framework
+\item We use NetworkX \cite{hagberg-exploring-2008} to build a shortest path matrix for every atom pair in the zeolite framework
 \item We then find all the rings up to some cutoff size base on the algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991}
@@
 \section*{Results}
-\label{sec:orgc227a60}
+\label{sec:orge495b06}
 \begin{itemize}
@@
 \begin{itemize}
+\item This ring finding rules often find rings that are not discussed in literature, and are not listed by the IZA
+\item These are classified as untabulated rings by Curtis and Deem \cite{curtis-statistical-2003}
+\item However it is possible that these rings are relevant for describing chemical, catalytic, or topological properties of zeolites
+\item Here we show an example of untabulated rings in the Chabazite framework
 \item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
@@
 \item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
+\begin{itemize}
+\item Crum rule removes these 8-MRs while still finding the 12-MR
+\item Sastre rule does not find the 8-MRs in the d6r or the 12-MR
+\end{itemize}
 \item The 12-MR is a cycle that circumferences the CHA cage
@@
 \includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
-\captionof{figure}{Chabazite framework with highlighted rings: 4-MR in green, 6-MR in blue, 8-MR in pink, 12-MR in purple.
-The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature.
-This is a placeholder figure, I want to show the rings in CHA for discussion.
-Hoping to get your input on how best to do that.
+\captionof{figure}{Chabazite framework with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
+The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure..
 \label{fig:cha-rings}}
@@
 \begin{itemize}
-\item Show the pentasil 6-membered cycle (MFI) and discuss how that is often considered a ring by catalysis researchers, but not a ring by the Goetzke definition
+\item On the other end of the spectrum, there are cycles that would not be classified as a rings by the connectivity rules previously outlined that display properties similar to rings
+\item These shortcut containing cycles can display chemical and/or geometric properties consistent with rings, and are of interest to catalysis researchers even though they are not considered rings by connectivity rules
+\item One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\Cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to  MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor in the cycle.
+Similar to Co\(^{\text{2+}}\) uptake in 3NN Al atoms in 6-MRs in other frameworks such as CHA and AEI.
+\item This 6-membered cycle would not be considered a ring by any of the connectivity rules outlined here due to the shortcut splitting the cycle into two 5-MRs
 \end{itemize}
 
+\begin{center}
+\includegraphics[width=.4\textwidth]{../figures/completed-figures/MFI-6MC.pdf}
+\captionof{figure}{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
+The 6-membered cycle would not be found as a ring by any of the connectivity ring rules (Goetzke, Crum, Sastre, or vertex symbol).
+\label{fig:mfi-6}}
+\end{center}
+
 \section*{Conclusions}
-\label{sec:org6e9315d}
+\label{sec:orgfbe47d4}
 

### 2021-05-06 outline/fix_ref.py

@@ -2,7 +2,7 @@ file = open('ref.bib',"r")
 lines = file.readlines()
 file.close()
 for i,line in enumerate(lines):
-    if 'article' in line:
+    if '@' in line:
         if '_' in line:
             line = line.replace('_','-')
             print(line)

### 2021-05-10 outline/outline.tex
added most common t-site tables

@@
-% Created 2021-05-06 Thu 16:34
+% Created 2021-05-10 Mon 09:54
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Introduction}
-\label{sec:org061dd26}
+\label{sec:orgb4f7112}
 \begin{itemize}
@@
 \section*{Software Description}
-\label{sec:org71ffeda}
+\label{sec:orgb3bb5f3}
 
@@
 \section*{Results}
-\label{sec:orge495b06}
+\label{sec:org2d4602a}
 \begin{itemize}
@@
 \begin{itemize}
-\item This ring finding rules often find rings that are not discussed in literature, and are not listed by the IZA
+\item Here we show the most common ring indices for T-sites in the IZA database using each of the ring finding rules
+\item \Cref{tab:goetzke-ts} shows the five most common ring indices for T-sites using the Goetzke  rule
+\end{itemize}
+\begin{table}[htbp]
+\caption{Most Common Ring Indices Using the Goetzke Rule \label{tab:goetzke-ts}}
+\centering
+\begin{tabular}{lrl}
+Ring Index & Count & Frameworks Containing Index\\
+\hline
+5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) & 23 & IMF(2), MEL(1), MFI(2), PRO(1),\\
+ &  & SVR(2), TUN(2), SFV(13)\\
+4\(_{\text{1}}\)\textbullet{}5\(_{\text{3}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}10\(_{\text{3}}\)\textbullet{}12\(_{\text{4}}\) & 14 & MEL(1), SFV(13)\\
+4\(_{\text{1}}\)\textbullet{}5\(_{\text{3}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{5}}\)\textbullet{}10\(_{\text{1}}\) & 14 & MEL(1), SFV(13)\\
+5\(_{\text{5}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}10\(_{\text{1}}\)\textbullet{}12\(_{\text{1}}\) & 14 & MEL(1), SFV(13)\\
+5\(_{\text{4}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}10\(_{\text{3}}\) & 14 & MEL(1), SFV(13)\\
+\end{tabular}
+\end{table}
+
+\newpage
+\begin{itemize}
+\item \Cref{tab:crum-ts} shows the five most common ring indices for T-sites using the Crum rule
+\end{itemize}
+\begin{table}[htbp]
+\caption{Five Most Common Ring Indices Using the Crum Rule \label{tab:crum-ts}}
+\centering
+\begin{tabular}{lrl}
+Ring Index & Count & Frameworks Containing Index\\
+\hline
+4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 31 & APC(1), GIS(1), MER(1), MWF(13),\\
+ &  & PAU(6), PHI(2), PWN(2), SIV(4)\\
+5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) & 25 & IMF(3), MEL(1), MFI(2), RRO(1),\\
+ &  & SVR(2), TUN(3), SFV(13)\\
+4\(_{\text{2}}\)\textbullet{}6\(_{\text{4}}\) & 17 & FAR(1), FRA(6), GIU(1), LIO(1),\\
+ &  & LOS(2), LTN(2), MAR(1), SOD(1),\\
+ &  & TOL(2)\\
+5\(_{\text{5}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}10\(_{\text{1}}\) & 17 & IMF(1), MEL(1), MFI(1), TUN(1),\\
+ &  & SFV(13)\\
+4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 16 & AFS(1), AFT(3), AFV(1), AFX(2),\\
+ &  & AVL(2), BPH(1), CHA(1), GME(1),\\
+ &  & SBE(1), SFW(3)\\
+\end{tabular}
+\end{table}
+
+\begin{itemize}
+\item \Cref{tab:sastre-ts} shows the five most common ring indices for T-sites using the Sastre rule
+\end{itemize}
+\begin{table}[htbp]
+\caption{Five Most Common Ring Indices Using the Sastre Rule \label{tab:sastre-ts}}
+\centering
+\begin{tabular}{lrl}
+Ring Index & Count & Frameworks Containing Index\\
+\hline
+4\(_{\text{2}}\)\textbullet{}6\(_{\text{4}}\) & 39 & AFG(3), CAN(1), FAR(4), FRA(6),\\
+ &  & GIU(5), LIO(4), LOS(2), LTN(2)\\
+ &  & MAR(4), SOD(1), TOL(7)\\
+5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) & 33 & IMF(3), MEL(2), MFI(2), RRO(1),\\
+ &  & SVR(2), TUN(2), SFV(21)\\
+4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 30 & GIS(1), MER(1), MWF(14), PAU(6),\\
+ &  & PHI(2), PWN(2), SIV(4)\\
+4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\) & 28 & AEI(3), AFT(3), AFV(1), AFX(2),\\
+ &  & AVL(2), CHA(1), GME(1), KFI(1),\\
+ &  & LTF(1), MWF(2), PAU(2), PWN(1),\\
+ &  & RHO(1), SAV(3), SFW(3), TSC(1)\\
+4\(_{\text{3}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{1}}\) & 24 & AFV(1), AVE(2), AVL(1), CLO(2),\\
+ &  & EAB(1), ERI(1), IFY(1), IRN(1),\\
+ &  & LEV(1), LTA(1), LTN(1), MOZ(1),\\
+ &  & OFF(1), SAT(1), SWY(2), TSC(1),\\
+ &  & UFI(1), PTT(1), SYT(3)\\
+\end{tabular}
+\end{table}
+
+\begin{itemize}
+\item \Cref{tab:vertex-ts} shows the five most common ring indices for T-sites using vertex symbols
+\end{itemize}
+\begin{table}[htbp]
+\caption{Five Most Common Ring Indices Using Vertex Symbolscite:bernauer-proton-2016 \label{tab:vertex-ts}}
+\centering
+\begin{tabular}{lrl}
+Vertex Symbol & Count & Frameworks Containing Index\\
+\hline
+4\textbullet{}4\textbullet{}6\textbullet{}6\textbullet{}6\textbullet{}6 & 40 & AFG(3), CAN(1), FAR(4), FRA(6),\\
+ &  & GIU(5), LIO(4), LOS(2), LTN(2),\\
+ &  & MAR(4), RON(1), SOD(1), TOL(7)\\
+4\textbullet{}4\textbullet{}4\textbullet{}6\textbullet{}8\textbullet{}8 & 32 & AEI(3), AFT(3), AFV(1), AFX(2),\\
+ &  & ATT(1), AVL(2), CHA(1), ETV(1),\\
+ &  & GME(1), KFI(1), LTF(1), MRT(2),\\
+ &  & MWF(2), PAU(2), PWN(1), RHO(1),\\
+ &  & SAV(3), SFW(3), TSC(1)\\
+4\textbullet{}4\textbullet{}4\textbullet{}6\textbullet{}6\textbullet{}8 & 30 & AFV(1), AVE(2), AVL(1), CGS(1),\\
+ &  & CLO(2), EAB(1), ERI(1), ETR(1),\\
+ &  & IFY(1), IRN(1), JSW(1), LEV(1),\\
+ &  & LTA(1), LTL(1), LTN(1), MOZ(3),\\
+ &  & OFF(1), PTT(1), SAT(1), SWY(2),\\
+ &  & SYT(3), TSC(1), UFI(1)\\
+4\textbullet{}4\textbullet{}4\textbullet{}8\textbullet{}8\textbullet{}8\(_{\text{2}}\) & 30 & GIS(1), MER(1), MWF(14), PAU(6),\\
+ &  & PHI(2), PWN(2), SIV(4)\\
+5\textbullet{}5\textbullet{}5\textbullet{}5\textbullet{}5\textbullet{}6 & 26 & DDR(1), DOH(2), IHW(1), IMF(1),\\
+ &  & MEL(1), MEP(1), MFI(1), MTN(1),\\
+ &  & SFS(1), SFV(15), TUN(1)\\
+\end{tabular}
+\end{table}
+
+\newpage
+
+\begin{itemize}
+\item These ring finding rules often find rings that are not commonly discussed in literature, and are not listed by the IZA
 \item These are classified as untabulated rings by Curtis and Deem \cite{curtis-statistical-2003}
@@
 \begin{itemize}
-\item Looking at results from CHA we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
+\item Looking at results for CHA in \Cref{tab:ring-counts} we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
 \item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
@@
 \section*{Conclusions}
-\label{sec:orgfbe47d4}
+\label{sec:org5a1fa60}
 

### 2021-05-13 outline/outline.tex
addressing Bill's changes

@@
-% Created 2021-05-10 Mon 09:54
+% Created 2021-05-13 Thu 12:16
 % Intended LaTeX compiler: pdflatex
@@
 \newpage
+\section*{Questions to settle with Bill:}
+\label{sec:org41438e4}
+\begin{enumerate}
+\item Where should vertex symbol go?
+\item Discuss the table and make sure we are in agreement on everything
+\item Naming the ring conventions:
+\begin{enumerate}
+\item Goetzke: ring, primitive ring, fundamental ring \cite{marians-network-1990,guttman-ring-1990,goetzke-properties-1991}
+\item Sastre: Shortest path ring, something else?
+\item Crum: Unstacked rings, something else?
+\end{enumerate}
+\item There are other ring definitions that I have not included in this study: strong ring, and very strong ring. Should we at least introduce those in the intro?
+\end{enumerate}
+
+\section*{Highlights}
+\label{sec:org6acb588}
+What is the main point of the paper? What problem is it addressing? What are the main insights?
+
+\begin{enumerate}
+\item Zeolite framework creates rings, defined as closed paths with no shortcuts.
+\item We provide a Python package that has the ability to find these rings based on previously published and new conventions
+\item We enumerate all rings up to 18 T-sites across all known Fws using each of these conventions
+\begin{enumerate}
+\item Would be very cool to classify Fws according to these rings\ldots{}.
+\begin{enumerate}
+\item Not sure what approach to use for that, but it is a good idea
+\end{enumerate}
+\item Compare to other classifications, eg cage-window vs channel
+\end{enumerate}
+\item Various conventions for associating some subsets of these rings with either a framework or with the T-sites of a framework. Compute and compare conventions across all known frameworks.
+\item Some features treated by convention as rings are not rings in a topological sense but are rather defined by the void space they create.
+\end{enumerate}
+
 \section*{Introduction}
-\label{sec:orgb4f7112}
+\label{sec:orge351f77}
+Introduction needs more structure.  Start with background, identify the problem you want to address.
+
+Zeolites are microporous and crystalline, natural to characterize by the sizes of the features present in the crystal.
+Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites of the zeolite.
+Provide some historical perspective here\ldots{}.
 \begin{itemize}
-\item Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges.
+\item Rings have been used to identify feasible zeolite topologies \cite{li-why-2014}, to descirbe the similarity and differences between zeolites \cite{curtis-statistical-2003}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints [will get citations for this] \red{need more thorough citing}
+
+\item Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges. \red{REFS}
 \begin{itemize}
@@
 \end{itemize}
-\item Topology of zeolite frameworks and associated tetrahedral sites (T-sites) are commonly characterized by their associated rings
-\item Rings are defined as a close cycle traversing the T-site and oxygen atoms of the framework, and cannot be decomposed into smaller cycles by a shortcut \cite{goetzke-properties-1991}.
-\item A shortcut is defined as a path connecting two nodes of a cycle that is shorter than both the paths connecting those nodes along the cycle.
+\item \red{Refer to/use table below} \Cref{tab:ring-def} summarizes the various topological features based on connectivity that can be used to describe a framework, T-site, or oxygen atom
+\item A path is a series of connected nodes from some source node to some target node
+\item A cycle is a path that starts and ends at the same node, and doesn't repeat any other nodes
+\item Rings in zeolites are defined as a cycle traversing the T-site and oxygen atoms (nodes) of the framework, and cannot be decomposed into smaller cycles by a shortcut \cite{goetzke-properties-1991,guttman-ring-1990}.
+\item A shortcut is defined as a path connecting two nodes of a cycle that is shorter than both the paths connecting those nodes along the cycle \cite{goetzke-properties-1991,guttman-ring-1990}
+\item Shortest path ring is a ring that is the shortest ring for at least one set of O-T-O on the cycle \cite{sastre-topological-2009}, this ring will be contained in the vertex symbol for at least one of the T-sites within the framework
 \item See \Cref{fig:cha-labeled} for example of rings
@@
 \begin{longtable}{l p{2.7cm} p{2.7cm} p{2.7cm} p{2.7cm}}
-\caption{Matrix showing relationship between frameoworks, nodes, paths, cycles, and various ring types. \label{tab:ring-def}}
+\caption{Matrix showing relationship between frameoworks, nodes, paths, cycles, and various ring types.
+\red{Vertex symbol doesn't belong in the first column.
+It isn't a topological feature.} \label{tab:ring-def}}
 \\
@@
 Unstacked Rings & Ring that does not traverse two stacked rings & A subset of the Rings above & All unstacked rings that pass through T-site & All unstacked rings that pass through oxygen atom\\
-Shortest Path Rings & Ring that is the shortest ring from at least one set of O-T-O on the cycle & A smaller subset of the rings above & All shortest path rings that pass through T-site & All shortest path rings that pass through oxygen atom\\
-Vertex Symbol & Way to classify the rings around a T-site, shortest ring (and its multiplicity) for each O-O pair around a T-site & Collection of vertex symbols for all symmetry distinct T-sites in framework & Vertex symbol for particular T-site & \\
+Shortest Path Rings & Ring that is the shortest ring for at least one set of O-T-O on the cycle & A smaller subset of the rings above & All shortest path rings that pass through T-site & All shortest path rings that pass through oxygen atom\\
+\red{Vertex Symbol} & Way to classify the rings around a T-site, shortest ring (and its multiplicity) for each O-O pair around a T-site & Collection of vertex symbols for all symmetry distinct T-sites in framework & Vertex symbol for particular T-site & \\
+Geometric rings & A cycle that may contain a shortcut, but has similar geometric/chemical properties to a ring without a shortcut & Contains a finite number of geometric rings & Can be described by the geometric rings that pass through & Can be described by the geometric rings that pass through\\
 \end{longtable}
@@
 \begin{itemize}
-\item Rings have been used to identify feasible zeolite topologies \cite{li-why-2014}, to descirbe the similarity and differences between zeolites \cite{curtis-statistical-2003}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints [will get citations for this]
-
-\item Different methods exist to count rings present in a zeolite
+\item Different conventions exist that can reduce the set of rings to more strictly defined properties
 \item These methods return different sets of rings
+\item We can use rings to characterize oxygen atoms, T-sites, and entire frameworks
+\item T-sites:
+\begin{itemize}
 \item Vertex symbols are the set of shortest paths connecting the 6 oxygen-oxygen pairs around a T-site \cite{okeeffe-vertex-1997}
@@
 \item Or we can count all the rings that do not have a short cut \cite{goetzke-properties-1991}
+\end{itemize}
+\item Oxygen atoms:
+\begin{itemize}
+\item Shortest path rings
+\item All rings with out a shortcut
+\end{itemize}
+\item Framework
+\begin{itemize}
+\item Vertex symbol rings
+\item Shortest path rings
+\item All rings with out a shortcut
+\end{itemize}
 \item Differences in ring counts leads to differences in how we describe the topology of zeolites.
 Therefore, when discussing the rings in a zeolite it is important to also state which method of ring counting is used.
-\item Here we report an analysis of rings and T-sites in a large number of zeolite frameworks using Zeolite Simulation Environment, a Python package that implements an efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} for finding rings in arbitrary frameworks.
-\item We compare the results of a number of common and new ring definitions applied to a large number of common zeolite frameworks.\cite{li-first-principles-2018}
+\item Here we present Zeolite Simulation Environment (ZSE), a Python package that implements the ring finding algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} to find rings up to a user defined cutoff size, and can implement the previously published ring set reduction conventions.
+\item We use ZSE to provide an analysis of rings using each of these conventions on the entire set of IZA zeolite frameworks to compare how they result in different characterizations
 \end{itemize}
 
+\red{Introduction needs to foreshadow the important insights. We captured those in your abstract. They need to appear here too.}
 \section*{Software Description}
-\label{sec:orgb3bb5f3}
+\label{sec:org97556f3}
 
@@
 \begin{itemize}
-\item All cycles with out a shortcut \cite{goetzke-properties-1991}
+\item All cycles without a shortcut \cite{goetzke-properties-1991}
 \item All shortest path cycles \cite{sastre-topological-2009}
@@
 \end{itemize}
-\item We have also implemented a new rule that finds all rings with out a shortcut, but excludes rings that are made by traversing a stacked set of rings.
+\item We have also implemented a new rule that finds all rings with out a shortcut, but excludes rings that are made by traversing a stacked set of rings. \red{Have to define stacked ring.}
 \begin{itemize}
@@
 \section*{Results}
-\label{sec:org2d4602a}
+\label{sec:orgbf8efd2}
 \begin{itemize}
@@
 \begin{itemize}
+\item Ring index was presented by Sastre and Corma as a way to list the rings that pass through a node in a zeolite \cite{sastre-topological-2009}
+\item List rings from smallest to largest, and any multiplicities are shown by a subscript
+\item This is a convenient way to characterize the atoms of a zeolite by the rings they are associated with
 \item Results in the Sastre column were found using ZSE but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}
@@
 \includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
-\captionof{figure}{Chabazite framework with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
+\captionof{figure}{Chabazite cage and d6r with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
 The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure..
@@
 \section*{Conclusions}
-\label{sec:org5a1fa60}
+\label{sec:orgd035302}
+\begin{itemize}
+\item \red{Rings of graph are well defined; here identify all rings up to XXX in YYY frameworks. Find that commonly reported (IZA) ring sizes miss certain rings.}
+\end{itemize}
 
 \begin{itemize}
-\item The method used to find rings in a zeolite will provide different ring counts
+\item The method used to find rings in a zeolite will provide different ring counts \red{unclear}
 \item When discussing rings in a zeolite it is import to disclose by which method those rings were found
@@
 \bibliography{ref}
+
+\section*{Acknowledgments}
+\label{sec:orgc0a6c6b}
+\begin{itemize}
+\item Funding
+\begin{itemize}
+\item CISTAR
+\item Schmitt Fellowship
+\end{itemize}
+\item Discussions
+\begin{itemize}
+\item Christian Baerlocher
+\end{itemize}
+\item Software:
+\begin{itemize}
+\item German Sastre: zeoTsites
+\end{itemize}
+\item Compute Resources
+\begin{itemize}
+\item CRC
+\end{itemize}
+\end{itemize}
 \end{document}

### 2021-05-25 outline/outline.tex
changes based on Bill's comments

@@
-% Created 2021-05-13 Thu 12:16
+% Created 2021-05-25 Tue 11:13
 % Intended LaTeX compiler: pdflatex
@@
 \section*{Questions to settle with Bill:}
-\label{sec:org41438e4}
+\label{sec:org8bf9ac1}
 \begin{enumerate}
-\item Where should vertex symbol go?
 \item Discuss the table and make sure we are in agreement on everything
@@
 \end{enumerate}
-\item There are other ring definitions that I have not included in this study: strong ring, and very strong ring. Should we at least introduce those in the intro?
 \end{enumerate}
@@
 \section*{Highlights}
-\label{sec:org6acb588}
+\label{sec:org389b28f}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \begin{enumerate}
-\item Zeolite framework creates rings, defined as closed paths with no shortcuts.
-\item We provide a Python package that has the ability to find these rings based on previously published and new conventions
-\item We enumerate all rings up to 18 T-sites across all known Fws using each of these conventions
+\item Zeolite framework and nodes commonly classified in terms of rings
+\item Various conventions for associating some subsets of rings with either a framework, the T-sites of a framework, or the oxygens of a framework.
+\item PROBLEM TO ADDRESS: The convention used will provide different ring counts
+\item SOLUTION: We provide a Python package that has the ability to identify and classify these rings based on previously published and new conventions
+\item This software tool uses a node-based search procedure that takes advantage of algorithms developed for general nets
+\item We compute and compare conventions across all known frameworks
+\item Some features treated by convention as rings are not rings in a topological sense but are rather defined by the void space they create.
+\item We enumerate all rings, defined as paths with no short-cut, up to 18 T-sites across all known Fws
 \begin{enumerate}
@@
 \end{enumerate}
-\item Various conventions for associating some subsets of these rings with either a framework or with the T-sites of a framework. Compute and compare conventions across all known frameworks.
-\item Some features treated by convention as rings are not rings in a topological sense but are rather defined by the void space they create.
 \end{enumerate}
@@
 \section*{Introduction}
-\label{sec:orge351f77}
+\label{sec:org645e4ed}
 Introduction needs more structure.  Start with background, identify the problem you want to address.
@@
 Provide some historical perspective here\ldots{}.
+
 \begin{itemize}
@@
 \end{center}
+
 \newpage
@@
 Paths & Collection of connected nodes from source to target & Periodic cell contains an infinite number of paths &  & \\
-Cycles & Path that starts and ends at the same node & Periodic cell Contains an infinite number of cycles &  & \\
+Cycles & Path that starts and ends at the same node & Periodic cell contains an infinite number of cycles &  & \\
 Rings & Cycle that contains no shortcuts & Contains a finite number of unique rings & All rings that pass through particular T-site & All rings that pass through particular oxygen atom\\
 Unstacked Rings & Ring that does not traverse two stacked rings & A subset of the Rings above & All unstacked rings that pass through T-site & All unstacked rings that pass through oxygen atom\\
-Shortest Path Rings & Ring that is the shortest ring for at least one set of O-T-O on the cycle & A smaller subset of the rings above & All shortest path rings that pass through T-site & All shortest path rings that pass through oxygen atom\\
+Shortest Path Rings & Ring that is the shortest ring for at least one set of O-T-O on the cycle & A smaller subset of the rings above & All shortest path rings starting from a T-site (Vertex) & All shortest path rings that pass through oxygen atom\\
 \red{Vertex Symbol} & Way to classify the rings around a T-site, shortest ring (and its multiplicity) for each O-O pair around a T-site & Collection of vertex symbols for all symmetry distinct T-sites in framework & Vertex symbol for particular T-site & \\
@@
 
+\textbf{\textbf{Problem to address}}
 \begin{itemize}
@@
 Therefore, when discussing the rings in a zeolite it is important to also state which method of ring counting is used.
+\end{itemize}
+
+\textbf{\textbf{Solution to problem}}
+\begin{itemize}
 \item Here we present Zeolite Simulation Environment (ZSE), a Python package that implements the ring finding algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} to find rings up to a user defined cutoff size, and can implement the previously published ring set reduction conventions.
@@
 \red{Introduction needs to foreshadow the important insights. We captured those in your abstract. They need to appear here too.}
+
+Using ZSE we show the differences in framework, T-site, and oxygen ring descriptors when using the various ring counting conventions.
+We highlight rings that are found by these conventions but not typically discussed for a number of frameworks.
+We also show that the vertex symbol, a common approach used to characterize T-sites misses important parts of the stereochemistry around a T-site.
+
 \section*{Software Description}
-\label{sec:org97556f3}
+\label{sec:org22a9053}
 
@@
 \section*{Results}
-\label{sec:orgbf8efd2}
+\label{sec:orgf799547}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:orgd035302}
+\label{sec:org57e4218}
 \begin{itemize}
@@
 \section*{Acknowledgments}
-\label{sec:orgc0a6c6b}
+\label{sec:org2e68444}
 \begin{itemize}

### 2021-05-25 outline/outline.tex
changes for Bill

@@
-% Created 2021-05-25 Tue 11:13
+% Created 2021-05-25 Tue 11:15
 % Intended LaTeX compiler: pdflatex
@@
 \newpage
-\section*{Questions to settle with Bill:}
-\label{sec:org8bf9ac1}
+\section*{Discussion With Bill:}
+\label{sec:orgecb832f}
 \begin{enumerate}
-\item Discuss the table and make sure we are in agreement on everything
+\item Go through table again to make sure we agree
+\begin{enumerate}
+\item Will move vertex symbol outside of table
+\end{enumerate}
 \item Naming the ring conventions:
@@
 \end{enumerate}
+\item Look at updated highlights
+\item Look at intro again
 \end{enumerate}
@@
 \section*{Highlights}
-\label{sec:org389b28f}
+\label{sec:orgf8dc093}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section*{Introduction}
-\label{sec:org645e4ed}
+\label{sec:org6606721}
 Introduction needs more structure.  Start with background, identify the problem you want to address.
@@
 \section*{Software Description}
-\label{sec:org22a9053}
+\label{sec:org4b5bf0e}
 
@@
 \section*{Results}
-\label{sec:orgf799547}
+\label{sec:orgfe2594e}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:org57e4218}
+\label{sec:orgcbcd1fd}
 \begin{itemize}
@@
 \section*{Acknowledgments}
-\label{sec:org2e68444}
+\label{sec:org48e2b73}
 \begin{itemize}

### 2021-08-19 outline/outline.tex
added abstract

@@
-% Created 2021-05-25 Tue 11:15
+% Created 2021-08-19 Thu 09:06
 % Intended LaTeX compiler: pdflatex
@@
 \newpage
+\section*{Abstract}
+\label{sec:org05df739}
+The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
+These ring descriptors have been used to identify feasible zeolite topologies, to describe the similarity and differences between zeolites, to identify sites or voids of catalytic relevance, and as machine learning fingerprints.
+Numerous definitions and algorithms for finding zeolite rings have been proposed and applied throughout the literature.
+Here we report an analysis of rings and T-sites in a large number of zeolite frameworks using Zeolite Simulation Environment, a Python package that implements an efficient algorithm presented by Goetzke and Klein for finding rings in arbitrary frameworks.
+We compare the result of a number of common and new ring definitions applied to a large number of common zeolite frameworks.
+We discover previously unrecognized rings in a number of frameworks.
+We show that the vertex symbol, a common approach used to characterize T-sites, misses important parts of the stereochemistry around a T-site, and propose an alternative definition.
+This tool provides an effective platform for characterizing zeolite and T-site structures useful for building models and doing machine learning.
+
 \section*{Discussion With Bill:}
-\label{sec:orgecb832f}
+\label{sec:org9bd16bb}
 \begin{enumerate}
@@
 \section*{Highlights}
-\label{sec:orgf8dc093}
+\label{sec:org61c4a30}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section*{Introduction}
-\label{sec:org6606721}
+\label{sec:org566a252}
 Introduction needs more structure.  Start with background, identify the problem you want to address.
@@
 \section*{Software Description}
-\label{sec:org4b5bf0e}
+\label{sec:orgb3ae4e5}
 
@@
 \section*{Results}
-\label{sec:orgfe2594e}
+\label{sec:orgc51c807}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:orgcbcd1fd}
+\label{sec:org8668819}
 \begin{itemize}
@@
 \section*{Acknowledgments}
-\label{sec:org48e2b73}
+\label{sec:orgf0794ce}
 \begin{itemize}

### 2021-08-19 outline/outline.tex
updated authors

@@
-% Created 2021-08-19 Thu 09:06
+% Created 2021-08-19 Thu 09:12
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage[capitalise]{cleveref}
-\author{Jerry T. Crum\(^{\text{a}}\), Justin R. Crum\(^{\text{b}}\), William F. Schneider\(^{\text{a,c}}\)}
+\author{Jerry T. Crum\(^{\text{a}}\), Justin R. Crum\(^{\text{b}}\), Cameron Taylor\(^{\text{a}}\), William F. Schneider\(^{\text{a,c}}\)}
 \date{}
@@
 \section*{Abstract}
-\label{sec:org05df739}
+\label{sec:orgfa1e0c6}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 \section*{Discussion With Bill:}
-\label{sec:org9bd16bb}
+\label{sec:org3c6ab20}
 \begin{enumerate}
@@
 \section*{Highlights}
-\label{sec:org61c4a30}
+\label{sec:org0b8f319}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section*{Introduction}
-\label{sec:org566a252}
+\label{sec:orgfd15178}
 Introduction needs more structure.  Start with background, identify the problem you want to address.
@@
 \section*{Software Description}
-\label{sec:orgb3ae4e5}
+\label{sec:org22fbc9a}
 
@@
 \section*{Results}
-\label{sec:orgc51c807}
+\label{sec:org7944fc3}
 \begin{itemize}
@@
 \section*{Conclusions}
-\label{sec:org8668819}
+\label{sec:orga80270e}
 \begin{itemize}
@@
 \section*{Acknowledgments}
-\label{sec:orgf0794ce}
+\label{sec:org6a03fc5}
 \begin{itemize}

### 2021-08-22 outline/outline.tex
sorted outline into paragraph

@@
-% Created 2021-08-19 Thu 09:12
+% Created 2021-08-22 Sun 13:26
 % Intended LaTeX compiler: pdflatex
@@
 \newpage
-\section*{Abstract}
-\label{sec:orgfa1e0c6}
+\section{Abstract}
+\label{sec:org598be1c}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 
-\section*{Discussion With Bill:}
-\label{sec:org3c6ab20}
+\section{Discussion With Bill:}
+\label{sec:org9624c98}
 \begin{enumerate}
@@
 
-\section*{Highlights}
-\label{sec:org0b8f319}
+\section{Highlights}
+\label{sec:orgcb8af34}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 
-\section*{Introduction}
-\label{sec:orgfd15178}
-Introduction needs more structure.  Start with background, identify the problem you want to address.
+\section{Introduction}
+\label{sec:org5fd2860}
 
-Zeolites are microporous and crystalline, natural to characterize by the sizes of the features present in the crystal.
+\textbf{\textbf{Paragraph 1: background on uses for zeolite rings}}
+\begin{itemize}
+\item Zeolites are microporous and crystalline, natural to characterize by the sizes of the features present in the crystal.
 Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites of the zeolite.
-Provide some historical perspective here\ldots{}.
 
+\item Rings have been used to identify feasible zeolite topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints [will get citations for this]
+\end{itemize}
+
+\textbf{\textbf{Paragraph 2: Explain features of graph representing zeolites}}
 \begin{itemize}
-\item Rings have been used to identify feasible zeolite topologies \cite{li-why-2014}, to descirbe the similarity and differences between zeolites \cite{curtis-statistical-2003}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints [will get citations for this] \red{need more thorough citing}
-
 \item Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges. \red{REFS}
@@
 \end{itemize}
-\item \red{Refer to/use table below} \Cref{tab:ring-def} summarizes the various topological features based on connectivity that can be used to describe a framework, T-site, or oxygen atom
+\item \Cref{tab:ring-def} summarizes the various topological features based on connectivity that can be used to describe a framework, T-site, or oxygen atom
 \item A path is a series of connected nodes from some source node to some target node
 \item A cycle is a path that starts and ends at the same node, and doesn't repeat any other nodes
+\end{itemize}
+
+\textbf{\textbf{Paragraph 3: Explain the various ring counting methodologies we intend to cover}}
+\begin{itemize}
 \item Rings in zeolites are defined as a cycle traversing the T-site and oxygen atoms (nodes) of the framework, and cannot be decomposed into smaller cycles by a shortcut \cite{goetzke-properties-1991,guttman-ring-1990}.
@@
 
-\textbf{\textbf{Problem to address}}
+\textbf{\textbf{Paragraph 4: Problem to address}}
 \begin{itemize}
@@
 
-\textbf{\textbf{Solution to problem}}
+\textbf{\textbf{Paragraph 5: Our solution to the problem}}
 \begin{itemize}
@@
 
-\red{Introduction needs to foreshadow the important insights. We captured those in your abstract. They need to appear here too.}
-
 Using ZSE we show the differences in framework, T-site, and oxygen ring descriptors when using the various ring counting conventions.
@@
 
-\section*{Software Description}
-\label{sec:org22fbc9a}
+\section{Software Description}
+\label{sec:orgbff414f}
 
+\textbf{\textbf{Paragraph 1: Basics of ZSE tootls}}
 \begin{itemize}
@@
 \item ZSE also includes CIF tools to read structure files for frameworks not listed in the IZA website, such as hypothetical zeolites, and return an Atoms object that can be used with ZSE
+\end{itemize}
+
+\textbf{\textbf{Paragraph 2: Implementation of ring counting methodologies}}
+\begin{itemize}
 \item ZSE has 3 previously published rules for ring finding implemented
@@
 \end{itemize}
+
+\textbf{\textbf{Paragraph 3: Process to find rings}}
+
 Process to find rings:
@@
 
-\section*{Results}
-\label{sec:org7944fc3}
+\section{Results}
+\label{sec:org6a10ab6}
+\textbf{\textbf{Paragraph 1: Differences in frameworks}}
 \begin{itemize}
@@
 
+\textbf{\textbf{Paragraph 2: Differences in T-sites}}
 \begin{itemize}
@@
 \begin{itemize}
+\item Include a heat map based on Cameron's results showing the coefficient of similarity that he defined
+\end{itemize}
+
+\textbf{\textbf{Paragraph 3: Differences in O-sites}}
+\begin{itemize}
 \item Number of unique oxygen sites
@@
 
+\textbf{\textbf{Paragraph 4: Comparison to previously published results by Sastre}}
 \begin{itemize}
@@
 
+\textbf{\textbf{Paragraph 5: Rings that are found using Goetzke method, but not discussed in most literature}}
+\begin{itemize}
+\item These ring finding rules often find rings that are not commonly discussed in literature, and are not listed by the IZA
+\item These are classified as untabulated rings by Curtis and Deem \cite{curtis-statistical-2003}
+\item However it is possible that these rings are relevant for describing chemical, catalytic, or topological properties of zeolites
+\item Here we show an example of untabulated rings in the Chabazite framework
+\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
+\begin{itemize}
+\item Looking at results for CHA in \Cref{tab:ring-counts} we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
+\item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
+\item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
+\begin{itemize}
+\item Crum rule removes these 8-MRs while still finding the 12-MR
+\item Sastre rule does not find the 8-MRs in the d6r or the 12-MR
+\end{itemize}
+\item The 12-MR is a cycle that circumferences the CHA cage
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
+\captionof{figure}{Chabazite cage and d6r with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
+The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure..
+\label{fig:cha-rings}}
+\end{center}
+
+\textbf{\textbf{Paragraph 6: Geometric rings}}
+\begin{itemize}
+\item On the other end of the spectrum, there are cycles that would not be classified as a rings by the connectivity rules previously outlined that display properties similar to rings
+\item These shortcut containing cycles can display chemical and/or geometric properties consistent with rings, and are of interest to catalysis researchers even though they are not considered rings by connectivity rules
+\item One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\Cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to  MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor in the cycle.
+Similar to Co\(^{\text{2+}}\) uptake in 3NN Al atoms in 6-MRs in other frameworks such as CHA and AEI.
+\item This 6-membered cycle would not be considered a ring by any of the connectivity rules outlined here due to the shortcut splitting the cycle into two 5-MRs
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=.4\textwidth]{../figures/completed-figures/MFI-6MC.pdf}
+\captionof{figure}{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
+The 6-membered cycle would not be found as a ring by any of the connectivity ring rules (Goetzke, Crum, Sastre, or vertex symbol).
+\label{fig:mfi-6}}
+\end{center}
+
+\textbf{\textbf{Paragraph 7: Stereochemistry concerns when using Ring Index}}
+\begin{itemize}
+\item The ring index for a T-site or O-site only list the size and counts of rings, but does not include any useful information about the orientation of those rings in space
+\item Because of this, multiple T-sites could have the same ring index, but with very different stereochemisty of those rings
+\item Stereochemistry could influence the chemical properties we care about, such as deprotonation energy, T-site substitution energy, or catalytic properties
+\item This would indicate that a ring index is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
+\item \textbf{\textbf{Here we insert a figure of two T-sites with the same ring index, but very different orientations}}
+\end{itemize}
+
+\textbf{\textbf{Not sure if this is relevant to include or not, but it provides some context for us too discuss}}
 \begin{itemize}
@@
 
-\begin{itemize}
-\item These ring finding rules often find rings that are not commonly discussed in literature, and are not listed by the IZA
-\item These are classified as untabulated rings by Curtis and Deem \cite{curtis-statistical-2003}
-\item However it is possible that these rings are relevant for describing chemical, catalytic, or topological properties of zeolites
-\item Here we show an example of untabulated rings in the Chabazite framework
-\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
-\begin{itemize}
-\item Looking at results for CHA in \Cref{tab:ring-counts} we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
-\item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
-\item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
-\begin{itemize}
-\item Crum rule removes these 8-MRs while still finding the 12-MR
-\item Sastre rule does not find the 8-MRs in the d6r or the 12-MR
-\end{itemize}
-\item The 12-MR is a cycle that circumferences the CHA cage
-\end{itemize}
-\end{itemize}
-\begin{center}
-\includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
-\captionof{figure}{Chabazite cage and d6r with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
-The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure..
-\label{fig:cha-rings}}
-\end{center}
-
-\begin{itemize}
-\item On the other end of the spectrum, there are cycles that would not be classified as a rings by the connectivity rules previously outlined that display properties similar to rings
-\item These shortcut containing cycles can display chemical and/or geometric properties consistent with rings, and are of interest to catalysis researchers even though they are not considered rings by connectivity rules
-\item One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\Cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to  MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor in the cycle.
-Similar to Co\(^{\text{2+}}\) uptake in 3NN Al atoms in 6-MRs in other frameworks such as CHA and AEI.
-\item This 6-membered cycle would not be considered a ring by any of the connectivity rules outlined here due to the shortcut splitting the cycle into two 5-MRs
-\end{itemize}
-
-\begin{center}
-\includegraphics[width=.4\textwidth]{../figures/completed-figures/MFI-6MC.pdf}
-\captionof{figure}{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
-The 6-membered cycle would not be found as a ring by any of the connectivity ring rules (Goetzke, Crum, Sastre, or vertex symbol).
-\label{fig:mfi-6}}
-\end{center}
-
-\section*{Conclusions}
-\label{sec:orga80270e}
+\section{Conclusions}
+\label{sec:org97a358f}
 \begin{itemize}
@@
 
-\section*{Acknowledgments}
-\label{sec:org6a03fc5}
+\section{Acknowledgments}
+\label{sec:org6d5013b}
 \begin{itemize}

### 2021-08-27 outline/outline.tex
updated some figures

@@
-% Created 2021-08-22 Sun 13:26
+% Created 2021-08-27 Fri 09:55
 % Intended LaTeX compiler: pdflatex
@@
 \section{Abstract}
-\label{sec:org598be1c}
+\label{sec:org9ce2e1a}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 \section{Discussion With Bill:}
-\label{sec:org9624c98}
+\label{sec:org38f32bc}
 \begin{enumerate}
@@
 \section{Highlights}
-\label{sec:orgcb8af34}
+\label{sec:org732966a}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section{Introduction}
-\label{sec:org5fd2860}
+\label{sec:orgdb6474c}
 
@@
 \begin{center}
-\includegraphics[width=0.60\textwidth]{../figures/completed-figures/ring-examples.pdf}
-\captionof{figure}{Cutout of the Chabazite framework showing a path from node 3 to node 9 outlined in teal, a cycle (3-4-18-19-20-17) outlined in red, an 8-MR in pink, and a 4-MR in green.
+\includegraphics[width=0.60\textwidth]{../figures/completed-figures/ring-examples2.pdf}
+\captionof{figure}{Cutout of the Chabazite framework showing a path (5-6-7-8-9) highlighted with purple bonds, a cycle (3-4-18-19-20-17) highlighted with blue bonds, an 8-MR filled in with pink, and a 4-MR filled in with green.
 Yellow atoms are Si (T-sites), and red atoms are oxygen.
@@
 \section{Software Description}
-\label{sec:orgbff414f}
+\label{sec:org249dbf1}
 
@@
 \section{Results}
-\label{sec:org6a10ab6}
+\label{sec:org64d529f}
 \textbf{\textbf{Paragraph 1: Differences in frameworks}}
@@
 
+Note that in \Cref{fig:tsite-frequency} a bar for IZA rings is not present. This is because the IZA does not list rings that pass through specific T-sites.
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/tsite-ring-counts.pdf}
+\captionof{figure}{Frequency of T-sites across all IZA frameworks containing ring sizes between 3- and 18-MR. \label{fig:tsite-frequency}}
+\end{center}
+
 \begin{itemize}
@@
 \label{fig:unique-os}}
+\end{center}
+
+Note that in \Cref{fig:osite-frequency} a bar for Vertex rings is not present. This is because the Vertex symbol only defines rings that pass through a specific T-site.
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/osite-ring-counts.pdf}
+\captionof{figure}{Frequency of O-sites across all IZA frameworks containing ring sizes between 3- and 18-MR. \label{fig:osite-frequency}}
 \end{center}
@@
 \section{Conclusions}
-\label{sec:org97a358f}
+\label{sec:org36e2194}
 \begin{itemize}
@@
 \section{Acknowledgments}
-\label{sec:org6d5013b}
+\label{sec:org1f2a844}
 \begin{itemize}

### 2022-01-09 outline/outline.tex
major revisions

@@
-% Created 2021-08-27 Fri 09:55
+% Created 2022-01-09 Sun 14:34
 % Intended LaTeX compiler: pdflatex
@@
 \item Department of Applied Mathematics, University of Arizona, 617 N Santa Rita Ave, Tucson, AZ 85721, USA\\
-\item Department of Chemistry and Biochmeistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
+\item Department of Chemistry and Biochemistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
 \end{asparaenum}
@@
 \section{Abstract}
-\label{sec:org9ce2e1a}
+\label{sec:orgef0dc6e}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 
-\section{Discussion With Bill:}
-\label{sec:org38f32bc}
-\begin{enumerate}
-\item Go through table again to make sure we agree
-\begin{enumerate}
-\item Will move vertex symbol outside of table
-\end{enumerate}
-\item Naming the ring conventions:
-\begin{enumerate}
-\item Goetzke: ring, primitive ring, fundamental ring \cite{marians-network-1990,guttman-ring-1990,goetzke-properties-1991}
-\item Sastre: Shortest path ring, something else?
-\item Crum: Unstacked rings, something else?
-\end{enumerate}
-\item Look at updated highlights
-\item Look at intro again
-\end{enumerate}
-
 \section{Highlights}
-\label{sec:org732966a}
+\label{sec:orgc351069}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \item This software tool uses a node-based search procedure that takes advantage of algorithms developed for general nets
-\item We compute and compare conventions across all known frameworks
+\item We compute and compare conventions across all known frameworks, highlighting differences in which rings get counted by which convention
 \item Some features treated by convention as rings are not rings in a topological sense but are rather defined by the void space they create.
-\item We enumerate all rings, defined as paths with no short-cut, up to 18 T-sites across all known Fws
-\begin{enumerate}
-\item Would be very cool to classify Fws according to these rings\ldots{}.
-\begin{enumerate}
-\item Not sure what approach to use for that, but it is a good idea
-\end{enumerate}
-\item Compare to other classifications, eg cage-window vs channel
-\end{enumerate}
+\item We show that the common method of describing a T-site by it's vertex symbol does not incorporate important information about the stereochemistry of the rings, and provide a new method for listing vertex symbol rings that compensates for that deficiency
 \end{enumerate}
@@
 \section{Introduction}
-\label{sec:orgdb6474c}
+\label{sec:orgc8ebded}
 
@@
 \section{Software Description}
-\label{sec:org249dbf1}
+\label{sec:org7e0878c}
 
@@
 \section{Results}
-\label{sec:org64d529f}
-\textbf{\textbf{Paragraph 1: Differences in frameworks}}
+\label{sec:org664ded4}
+
+\textbf{\textbf{Paragraph 1: IZA doesn't list all rings in a framework}}
 \begin{itemize}
-\item Ring counts frequency plots
-\begin{itemize}
-\item Plot showing how many frameworks on the IZA contain each size ring found using the various ring counting methods
-\item This highlights the differences in the ring rules, and shows that results will vary depending on rule.
+\item IZA is a common reference used to identify all the rings in a zeolite framework
+\item It only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the vertex symbol of a T-site
+\item These rings have been called tabulated rings in previous literature \cite{curtis-statistical-2003}
+\item In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the tabulated rings
+\item These rings still may provide important information about the topology of a zeolite framework, or the local void environment around a T-site
+\item \cref{fig:framework-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database
+\item At larger sized rings (>8-MR) we see the two counts diverge
 \end{itemize}
-\end{itemize}
+
 \begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/ring-counts.pdf}
-\captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
-[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website.
-Currently the IZA does not show any ring data for the SVY framework, providing one less framework to count.
-\label{fig:ring-counts}}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/rings-vs-iza-rings.pdf}
+\captionof{figure}{Counts of frameworks containing each size rings between 3 and 18-MR using the Goetzke algorithm and the tabulated rings on the IZA database. \label{fig:framework-counts}}
 \end{center}
@@
 \begin{itemize}
+\item Starting with T-sites we can find all the rings that pass through each T-site using a modified version of the Goetzke algorithm.
+\item We can prune this list of rings using other definitions, such as the shortest path rings presented by Sastre et al.
+\cite{sastre-topological-2009}, by counting only the vertex symbol rings \cite{okeeffe-vertex-1997}, or by our new algorithm that removes any ring that transcends a stacked set of rings.
 \item Number of unique T-sites
@@
 
-Note that in \Cref{fig:tsite-frequency} a bar for IZA rings is not present. This is because the IZA does not list rings that pass through specific T-sites.
-
 \begin{center}
@@
 \begin{itemize}
-\item Include a heat map based on Cameron's results showing the coefficient of similarity that he defined
+\item Here we use a similarity score to determine how similar or different the rings found by each counting method are to each other
+\item The similarity score is determined by \cref{eq:similarity}, where \(sr\) is the number of similar rings found by each method, and \(mr\) is the max number of rings found by either method.
 \end{itemize}
+\begin{equation}\label{eq:similarity}
+s = \frac{sr}{mr}
+\end{equation}
+\begin{itemize}
+\item We compare each method for every T-site in the IZA database, and average the similarity scores for each method pair
+\item The results are shown in the following heat map
+\item We see that down the diagonal is each method compared to itself, and thus has a similarity score of 1
+\item The remainder of the table follows intuition
+\item the most restrictive counting method (vertex symbol rings) compared to the least restrictive method (rings) has the lowest similarity score
+\item The two most similar ring counting methods are this work compared to the shortest path rings
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/similarity-score.png}
+\captionof{figure}{Heat map showing the similarity score for four ring counting methods.
+Similarity score of 1 means identical, and a similarity score of zero implies no similar rings found.
+\label{fig:similarity}}
+\end{center}
 
@@
 
-\textbf{\textbf{Paragraph 7: Stereochemistry concerns when using Ring Index}}
+\textbf{\textbf{Paragraph 7: Stereochemistry concerns when using vertex symbols}}
 \begin{itemize}
-\item The ring index for a T-site or O-site only list the size and counts of rings, but does not include any useful information about the orientation of those rings in space
-\item Because of this, multiple T-sites could have the same ring index, but with very different stereochemisty of those rings
-\item Stereochemistry could influence the chemical properties we care about, such as deprotonation energy, T-site substitution energy, or catalytic properties
-\item This would indicate that a ring index is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
-\item \textbf{\textbf{Here we insert a figure of two T-sites with the same ring index, but very different orientations}}
+\item Vertex symbols are common way to identify similar T-sites in zeolite frameworks
+\item The vertex symbol as defined lists the rings included by finding pairs of opposite edge rings, and listing them from smallest pair to largest pair
+\item This method does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments
+\item For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\textbullet{}5\(_{\text{2}}\)\textbullet{}5\textbullet{}8\(_{\text{2}}\)\textbullet{}5\textbullet{}8\(_{\text{2}}\)
+\begin{itemize}
+\item So each T-site is a member of one 4-MR, four 5-MRs, and four 8-MRs
+\item However, the orientation of those rings around each of those T-sites are not identical
+\item FIGURE below shows a cutout of each of these frameworks only including the atoms that make up the vertex symbol rings around the specified T-sites
+\item We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up MON T1
+\item \red{I can make this figure better with more angles, and maybe include a line diagram explaining the differences}
+\end{itemize}
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/stereo.png}
+\captionof{figure}{Cutout of MOR, EON, and MON that only shows the rings associated with the vertex symbol of T3, T9, and T1 respectively. \label{fig:stereo}}
+\end{center}
+
+\begin{itemize}
+\item This has lead us to create a new method for listing the rings in the vertex symbol that takes into account the structural connection of the rings
+\begin{itemize}
+\item This is accomplished by assigning a weight to each of the oxygen atoms around the T-site in question based on the size of rings that connect to that oxygen
+\item \red{need to work out how to describe the method here}
+\item With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}5\(_{\text{2}}\)\textbullet{}5\textbullet{}4\textbullet{}5 and MON T1 as: 8\(_{\text{2}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}4\textbullet{}5\textbullet{}5\(_{\text{2}}\)\textbullet{}5
+\item The difference is subtle, but highlights the distinct structural  difference between the two types of T-sites that is not otherwise captured by a vertex symbol
+\end{itemize}
+\item \red{Here we would include a table of some of the most common vertex symbols, and how many different variations they have with this new descriptor}
+
+\item Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about, such as deprotonation energy, T-site substitution energy, or catalytic properties
+\item This would indicate that a vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
 \end{itemize}
@@
 \newpage
+\red{Removed paragraph, because we don't want to describe frameworks using these other ring finding methods. Keeping text in case we want to bring this back.}
+\textbf{\textbf{Paragraph 1: Differences in frameworks}}
+\begin{itemize}
+\item Ring counts frequency plots
+\begin{itemize}
+\item Plot showing how many frameworks on the IZA contain each size ring found using the various ring counting methods
+\item This highlights the differences in the ring rules, and shows that results will vary depending on rule.
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/ring-counts.pdf}
+\captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
+[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website.
+Currently the IZA does not show any ring data for the SVY framework, providing one less framework to count.
+\label{fig:ring-counts}}
+\end{center}
 
 \section{Conclusions}
-\label{sec:org36e2194}
+\label{sec:orgf48f218}
 \begin{itemize}
@@
 \section{Acknowledgments}
-\label{sec:org1f2a844}
+\label{sec:org3dde92a}
 \begin{itemize}

### 2022-01-22 outline/outline.tex
updated outline

@@
-% Created 2022-01-09 Sun 14:34
+% Created 2022-01-22 Sat 12:56
 % Intended LaTeX compiler: pdflatex
@@
 \section{Abstract}
-\label{sec:orgef0dc6e}
+\label{sec:org14a7c72}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 \section{Highlights}
-\label{sec:orgc351069}
+\label{sec:orga9a0704}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section{Introduction}
-\label{sec:orgc8ebded}
+\label{sec:org9bb9b10}
 
@@
 \section{Software Description}
-\label{sec:org7e0878c}
+\label{sec:org2e90fc2}
 
@@
 \section{Results}
-\label{sec:org664ded4}
+\label{sec:orge3c702c}
 
@@
 
-\textbf{\textbf{Paragraph 2: Differences in T-sites}}
+\textbf{\textbf{Paragraph 2: Example of untabulated rings in CHA shows 12-MR belt on cage, and 8-MRs in the D6R}}
 \begin{itemize}
-\item Starting with T-sites we can find all the rings that pass through each T-site using a modified version of the Goetzke algorithm.
+\item Taking a closer look at some of these untabulated rings, highlights rings not typically but listed for some frameworks, but still relevant to describing their topology
+\item Here we show an example of untabulated rings in the Chabazite framework
+\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
+\begin{itemize}
+\item Looking at results for CHA in \Cref{tab:ring-counts} we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
+\item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
+\item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
+\item The 12-MR is a cycle that circumferences the CHA cage
+\end{itemize}
+\end{itemize}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
+\captionof{figure}{Chabazite cage and d6r with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
+The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure.
+\label{fig:cha-rings}}
+\end{center}
+
+\textbf{\textbf{Paragraph 3: Untabulated rings in AFI also reveals rings traversing a pair of stacked rings}}
+\begin{itemize}
+\item AFI has one unique T-site
+\item According to the IZA it contains 4, 6, and 12-MRs
+\item When we search using the Goetzke algorithm, we also find that it contains 14-MRs that traverses two stacked 12-MRs shown in \cref{fig:afi-14}
+\item These types of rings may not be of interest depending on what you want to describe
+\item Using the method outlined in section X.Y we can remove these types of rings from the Goetzke ring counts
+\end{itemize}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{../figures/completed-figures/afi-14.pdf}
+\captionof{figure}{12-MR channel in AFI with a 14-MR traversing 7 T-sites of each 12-MR highlighted in purple. \label{fig:afi-14}}
+\end{center}
+
+\textbf{\textbf{Paragraph 4: Comparing all 4 ringing finding conventions and IZA listed rings by counting the number of frameworks containing each size ring}}
+\begin{itemize}
+\item Plot showing how many frameworks on the IZA contain each size ring found using the various ring counting methods
+\item This highlights the differences in the conventions, and shows that results will vary depending on method used.
+\item In general a hierarchy of rings sizes found by each method is rings>this work>shortest path rings>vertex symbol rings
+\item While the IZA lists all vertex symbol rings, and a selection of general rings
+\end{itemize}
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/ring-counts.pdf}
+\captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
+[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website.
+Currently the IZA does not show any ring data for the SVY framework, providing one less framework to count.
+\label{fig:ring-counts}}
+\end{center}
+
+\textbf{\textbf{Paragraph 5: Some cycles behave like rings while not be considered rings by connectivity rules}} 
+\begin{itemize}
+\item On the other end of the spectrum, there are cycles that would not be classified as a rings by the connectivity rules previously outlined, but display properties similar to rings
+\item The trade-off to using well defined connectivity based ring definitions leaves out these particular void environments that may still be of interest
+\item These shortcut containing cycles can display chemical and/or geometric properties consistent with rings, and are of interest to catalysis researchers even though they are not considered rings by connectivity rules
+\item One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\Cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to  MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor in the cycle.
+Similar to Co\(^{\text{2+}}\) uptake in 3NN Al atoms in 6-MRs in other frameworks such as CHA and AEI.
+\item This 6-membered cycle would not be considered a ring by any of the connectivity rules outlined here due to the shortcut splitting the cycle into two 5-MRs
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=.4\textwidth]{../figures/completed-figures/MFI-6MC.pdf}
+\captionof{figure}{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
+The 6-membered cycle would not be found as a ring by any of the connectivity ring rules (Goetzke, Crum, Sastre, or vertex symbol).
+\label{fig:mfi-6}}
+\end{center}
+
+\textbf{\textbf{Paragraph 6: We can describe T-sites of a framework by the rings of that framework that pass through it AFI Example}}
+\begin{itemize}
+\item Take for example the AFI framework, made of 4, 6, 12, and 14-MRs and containing one symmetry distinct T-site
+\item We can describe that T-site by counting how many rings of each size pass through it
 \item We can prune this list of rings using other definitions, such as the shortest path rings presented by Sastre et al.
-\cite{sastre-topological-2009}, by counting only the vertex symbol rings \cite{okeeffe-vertex-1997}, or by our new algorithm that removes any ring that transcends a stacked set of rings.
+\cite{sastre-topological-2009}, by counting only the vertex symbol rings \cite{okeeffe-vertex-1997}, or by our new algorithm that removes any ring that traverses a stacked set of rings
+\item Listing these rings by using a ring index (size\(_{\text{count}}\) from smallest to largest rings) \cite{sastre-topological-2009} shows the differences in counts by each method
+\begin{itemize}
+\item Rings: 4\textbullet{}6\(_{\text{13}}\)\textbullet{}12\textbullet{}14\(_{\text{7}}\)
+\item This work: 4\textbullet{}6\(_{\text{13}}\)\textbullet{}12
+\item Shortest Path Rings: 4\textbullet{}6\(_{\text{13}}\)
+\item Vertex Symbol Rings: 4\textbullet{}6\(_{\text{11}}\)
+\end{itemize}
+\item Each method returns a different count, this is highlighted in figure\ldots{}
+\end{itemize}
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/afi-funnel.pdf}
+\captionof{figure}{Diagram showing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions.
+\label{fig:afi-funnel}}
+\end{center}
+
+\textbf{\textbf{Paragraph 7: Comparison to previously published results by Sastre}}
+\begin{table}[htbp]
+\caption{Comparison of Ring Indices for the T-sites in Various Uninodal Zeolite Frameworks \label{tab:ring-counts}}
+\centering
+\begin{tabular}{lllll}
+Framework & Rings & This Work & Shortest Path Rings \cite{sastre-topological-2009} & Vertex Symbol Rings \cite{baerlocher-database-nodate}\\
+\hline
+ABW & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{2}}\)\\
+ACO & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}10\(_{\text{15}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\\
+AFI & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\)\textbullet{}14\(_{\text{7}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{11}}\)\\
+ANA & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{8}}\)\\
+ATO & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}8\(_{\text{8}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\\
+BCT & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{12}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+CHA & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
+DFT & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\textbullet{}10\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{4}}\)\textbullet{}8\(_{\text{6}}\)\\
+GIS & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+GME & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{7}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
+MER & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\textbullet{}10\(_{\text{10}}\)\textbullet{}14\(_{\text{14}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
+MON & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{4}}\)\textbullet{}8\(_{\text{4}}\)\\
+NPO & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
+\end{tabular}
+\end{table}
+
+\begin{itemize}
+\item Reproduce the results from Sastre paper, show ring counts with the other rules, \Cref{tab:ring-counts}
+\begin{itemize}
+\item Results in the Sastre column were found using ZSE but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}
+\item Results in the Vertex symbol rings column were also found with ZSE but agree directly will the vertex symbols listed on the IZA website
+\begin{itemize}
+\item These vertex symbols have been compressed into a ring index to compare with the other methods
+\end{itemize}
+\item This provides an in depth look at some of the frameworks and the differences in rings found by each rule.
+\item Leads into the next section discussing the specific rings of CHA and pentasil that do or don't get counted by each rule.
+\end{itemize}
+\end{itemize}
+
+\textbf{\textbf{Paragraph 8: Example of T-site ring counts for 6 symmetry distinct T-sites in MOZ shows limitations of shortest path and vertex symbol rings}}
+\begin{itemize}
+\item MOZ framework is made of 4, 6, 8, 10, 12, 14, and 18-MRs, and contains 6 symmetry distinct T-sites
+\item \cref{table:moz} shows the ring index for each T-site using each ring finding method
+\end{itemize}
+\begin{table}[htbp]
+\caption{Ring indices for each distinct T-site in the MOZ framework using each ring counting convention. \label{table:moz}}
+\centering
+\begin{tabular}{lllll}
+T-Site & Rings & This Work & Shortest Path Rings & Vertex Symbol Rings\\
+\hline
+T1 & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{7}}\)•10\(_{\text{7}}\)•18\(_{\text{5}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{3}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{3}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\\
+T2 & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{7}}\)•10\(_{\text{7}}\)•14\(_{\text{5}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{3}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{3}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\\
+T3 & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\(_{\text{5}}\)•10\(_{\text{4}}\)•12\(_{\text{4}}\)•14\(_{\text{5}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8•12\(_{\text{4}}\) & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8 & 4\(_{\text{3}}\)•6\(_{\text{2}}\)•8\\
+T4 & 4\(_{\text{2}}\)•6•8\(_{\text{6}}\)•10\(_{\text{6}}\)•12•18\(_{\text{26}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{6}}\)•12 & 4\(_{\text{2}}\)•6•8\(_{\text{6}}\)•12 & 4\(_{\text{2}}\)•6•8\(_{\text{6}}\)•12\\
+T5 & 4\(_{\text{2}}\)•6•8\(_{\text{7}}\)•10\(_{\text{6}}\)•14\(_{\text{18}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{7}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{7}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{7}}\)\\
+T6 & 4\(_{\text{2}}\)•6•8\(_{\text{3}}\)•10\(_{\text{2}}\)•12\(_{\text{8}}\)•14\(_{\text{18}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{3}}\)•12\(_{\text{8}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{3}}\) & 4\(_{\text{2}}\)•6•8\(_{\text{3}}\)\\
+\end{tabular}
+\end{table}
+
+\begin{itemize}
+\item \cref{fig:moz} shows the T-site locations inside the framework
+\item If you were interested in which T-sites have access to the 12-MR channels, the shortest path rings and vertex symbol rings would only say T3
+\item However general rings, and this work would tell you T4 and T6 also have access to the 12-MR channels as highlighted in Figure ..
+\end{itemize}
+\begin{center}
+\includegraphics[width=.5\textwidth]{../figures/completed-figures/moz.pdf}
+\captionof{figure}{Cutout of the MOZ framework showing two 12-MR channels, with distinct T-sites highlighted. T1: navy, T2: green, T3: orange, T4: purple, T5: blue, T6: red. \label{fig:moz}}
+\end{center}
+
+\textbf{\textbf{Paragraph 9: Comparing ring counts for every T-site across the IZA database highlights differences in conventions}}
+\begin{itemize}
 \item Number of unique T-sites
@@
 \item We can characterize those T-sites by the rings that pass through them
-\item Sastre did this, and called the list of rings, the ring index
-\item If we do this using different rules for ring finding how do the results change?
-\begin{itemize}
-\item See \Cref{fig:unique-ts}
-\end{itemize}
 \item Most common T-site ring index using Goetzke method is: 5\(_{\text{6}}\)\textbullet{}10\(_{\text{4}}\) showing up 23 times through the IZA frameworks.
@@
 \end{itemize}
-\item This raises the question, if you want to use machine learning to correlate T-site rings to chemical properties, which ring method should you use?
+\item This raises the question, if you want to ascertain chemical or physical properties about a T-site based on it's ring count, the results will depend on the convention used
 \end{itemize}
@@
 \begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/tsite-ring-counts.pdf}
-\captionof{figure}{Frequency of T-sites across all IZA frameworks containing ring sizes between 3- and 18-MR. \label{fig:tsite-frequency}}
+\includegraphics[width=.8\textwidth]{../figures/completed-figures/dist-cumudist.pdf}
+\captionof{figure}{Frequency of T-sites across all IZA frameworks containing ring sizes between 3- and 18-MR (left), and cumulative distribution of T-sites containing each ring size normalized to final rings value (right).
+\label{fig:tsite-frequency}}
 \end{center}
 
+\textbf{\textbf{Paragraph 10: Similarity of results returned by each ring finding convention}}
 \begin{itemize}
@@
 \begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/similarity-score.png}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/similarity-heat-map.pdf}
 \captionof{figure}{Heat map showing the similarity score for four ring counting methods.
@@
 
-\textbf{\textbf{Paragraph 3: Differences in O-sites}}
-\begin{itemize}
-\item Number of unique oxygen sites
-\begin{itemize}
-\item We can repeat this method for the oxygen atoms in all the frameworks
-\item Counting the symmetry distinct oxygen atoms in each framework on the IZA database leads to a total count of 3219
-\item We can classify those oxygen atoms based on the rings that pass through them, using the various ring counting rules
-\item \Cref{fig:unique-os} shows counts based on ring finding rules
-\item The percentage of unique oxygen sites is much lower than the percentage of unique T-sites for every ring finding method
-\end{itemize}
-\end{itemize}
-
-\begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/unique-os.pdf}
-\captionof{figure}{Number of unique oxygen sites when classified by the rings passing through them using varrious ring finding rules.
-Vertex method not included, since that is a way to classify T-sites only.
-\label{fig:unique-os}}
-\end{center}
-
-Note that in \Cref{fig:osite-frequency} a bar for Vertex rings is not present. This is because the Vertex symbol only defines rings that pass through a specific T-site.
-
-\begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/osite-ring-counts.pdf}
-\captionof{figure}{Frequency of O-sites across all IZA frameworks containing ring sizes between 3- and 18-MR. \label{fig:osite-frequency}}
-\end{center}
-
-\textbf{\textbf{Paragraph 4: Comparison to previously published results by Sastre}}
-\begin{itemize}
-\item Reproduce the results from Sastre paper, show ring counts with the other rules, \Cref{tab:ring-counts}
-\begin{itemize}
-\item Ring index was presented by Sastre and Corma as a way to list the rings that pass through a node in a zeolite \cite{sastre-topological-2009}
-\item List rings from smallest to largest, and any multiplicities are shown by a subscript
-\item This is a convenient way to characterize the atoms of a zeolite by the rings they are associated with
-\item Results in the Sastre column were found using ZSE but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}
-\item This provides an in depth look at some of the frameworks and the differences in rings found by each rule.
-\item Leads into the next section discussing the specific rings of CHA and pentasil that do or don't get counted by each rule.
-\end{itemize}
-\end{itemize}
-
-\begin{table}[htbp]
-\caption{Comparison of Ring Indices for the T-sites in Various Uninodal Zeolite Frameworks \label{tab:ring-counts}}
-\centering
-\begin{tabular}{llll}
-Framework & Goetzke & Crum & Sastre \cite{sastre-topological-2009}\\
-\hline
-ABW & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
-ACO & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}10\(_{\text{15}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{6}}\)\\
-AFI & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\)\textbullet{}14\(_{\text{7}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{13}}\)\\
-ANA & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{2}}\)\textbullet{}8\(_{\text{16}}\)\\
-ATO & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}8\(_{\text{8}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\textbullet{}12\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{9}}\)\\
-BCT & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{20}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{12}}\) & 4\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
-CHA & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
-DFT & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\textbullet{}10\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\) & 4\(_{\text{2}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}8\(_{\text{10}}\)\\
-GIS & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
-GME & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{7}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\textbullet{}12\(_{\text{1}}\) & 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{2}}\)\\
-MER & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\textbullet{}10\(_{\text{10}}\)\textbullet{}14\(_{\text{14}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\) & 4\(_{\text{3}}\)\textbullet{}8\(_{\text{4}}\)\\
-MON & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\) & 4\(_{\text{1}}\)\textbullet{}5\(_{\text{5}}\)\textbullet{}8\(_{\text{6}}\)\\
-NPO & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\textbullet{}12\(_{\text{40}}\) & 3\(_{\text{1}}\)\textbullet{}6\(_{\text{6}}\)\\
-\end{tabular}
-\end{table}
-
-\textbf{\textbf{Paragraph 5: Rings that are found using Goetzke method, but not discussed in most literature}}
-\begin{itemize}
-\item These ring finding rules often find rings that are not commonly discussed in literature, and are not listed by the IZA
-\item These are classified as untabulated rings by Curtis and Deem \cite{curtis-statistical-2003}
-\item However it is possible that these rings are relevant for describing chemical, catalytic, or topological properties of zeolites
-\item Here we show an example of untabulated rings in the Chabazite framework
-\item Show the cage belts results for CHA, AFT, etc\ldots{} and discuss how those rings don't show up in previous literature, \Cref{fig:cha-rings}
-\begin{itemize}
-\item Looking at results for CHA in \Cref{tab:ring-counts} we see the Goetzke method finds 4\(_{\text{3}}\)\textbullet{}6\(_{\text{1}}\)\textbullet{}8\(_{\text{6}}\)\textbullet{}12\(_{\text{1}}\)
-\item This is different from the results in the Sastre paper \cite{sastre-topological-2009}, in that they only show 2 8-MRs and no 12-MRs
-\item The extra 8-MRs result from cycles traversing nodes in both 6-MRs of the d6r
-\begin{itemize}
-\item Crum rule removes these 8-MRs while still finding the 12-MR
-\item Sastre rule does not find the 8-MRs in the d6r or the 12-MR
-\end{itemize}
-\item The 12-MR is a cycle that circumferences the CHA cage
-\end{itemize}
-\end{itemize}
-\begin{center}
-\includegraphics[width=0.45\textwidth]{../figures/completed-figures/cha-all-rings.pdf}
-\captionof{figure}{Chabazite cage and d6r with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
-The 8-MR in the d6r and the 12-MR are rings not typically discussed in literature, Si atoms have been replaced with Al atoms to help identify those rings in the overall cage structure..
-\label{fig:cha-rings}}
-\end{center}
-
-\textbf{\textbf{Paragraph 6: Geometric rings}}
-\begin{itemize}
-\item On the other end of the spectrum, there are cycles that would not be classified as a rings by the connectivity rules previously outlined that display properties similar to rings
-\item These shortcut containing cycles can display chemical and/or geometric properties consistent with rings, and are of interest to catalysis researchers even though they are not considered rings by connectivity rules
-\item One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\Cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to  MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor in the cycle.
-Similar to Co\(^{\text{2+}}\) uptake in 3NN Al atoms in 6-MRs in other frameworks such as CHA and AEI.
-\item This 6-membered cycle would not be considered a ring by any of the connectivity rules outlined here due to the shortcut splitting the cycle into two 5-MRs
-\end{itemize}
-
-\begin{center}
-\includegraphics[width=.4\textwidth]{../figures/completed-figures/MFI-6MC.pdf}
-\captionof{figure}{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
-The 6-membered cycle would not be found as a ring by any of the connectivity ring rules (Goetzke, Crum, Sastre, or vertex symbol).
-\label{fig:mfi-6}}
-\end{center}
-
-\textbf{\textbf{Paragraph 7: Stereochemistry concerns when using vertex symbols}}
+\textbf{\textbf{Paragraph 11: Stereochemistry concerns when using vertex symbols}}
 \begin{itemize}
@@
 
-\textbf{\textbf{Not sure if this is relevant to include or not, but it provides some context for us too discuss}}
+\textbf{\textbf{Everything below here is not intended to be included in the main text of the paper}}
+
+\textbf{\textbf{Paragraph 3: Differences in O-sites}}
+\begin{itemize}
+\item Number of unique oxygen sites
+\begin{itemize}
+\item We can repeat this method for the oxygen atoms in all the frameworks
+\item Counting the symmetry distinct oxygen atoms in each framework on the IZA database leads to a total count of 3219
+\item We can classify those oxygen atoms based on the rings that pass through them, using the various ring counting rules
+\item \Cref{fig:unique-os} shows counts based on ring finding rules
+\item The percentage of unique oxygen sites is much lower than the percentage of unique T-sites for every ring finding method
+\end{itemize}
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/unique-os.pdf}
+\captionof{figure}{Number of unique oxygen sites when classified by the rings passing through them using varrious ring finding rules.
+Vertex method not included, since that is a way to classify T-sites only.
+\label{fig:unique-os}}
+\end{center}
+
+Note that in \Cref{fig:osite-frequency} a bar for Vertex rings is not present. This is because the Vertex symbol only defines rings that pass through a specific T-site.
+
+\begin{center}
+\includegraphics[width=.6\textwidth]{../figures/completed-figures/osite-ring-counts.pdf}
+\captionof{figure}{Frequency of O-sites across all IZA frameworks containing ring sizes between 3- and 18-MR. \label{fig:osite-frequency}}
+\end{center}
 \begin{itemize}
@@
 \textbf{\textbf{Paragraph 1: Differences in frameworks}}
-\begin{itemize}
-\item Ring counts frequency plots
-\begin{itemize}
-\item Plot showing how many frameworks on the IZA contain each size ring found using the various ring counting methods
-\item This highlights the differences in the ring rules, and shows that results will vary depending on rule.
-\end{itemize}
-\end{itemize}
-\begin{center}
-\includegraphics[width=.6\textwidth]{../figures/completed-figures/ring-counts.pdf}
-\captionof{figure}{Number of IZA frameworks containing each size ring, using the various ring counting rules.
-[This will be updated with the Sastre method, vertex method, and the rings listed on  the IZA website.
-Currently the IZA does not show any ring data for the SVY framework, providing one less framework to count.
-\label{fig:ring-counts}}
-\end{center}
 
 \section{Conclusions}
-\label{sec:orgf48f218}
+\label{sec:org2135ac7}
 \begin{itemize}
@@
 \section{Acknowledgments}
-\label{sec:org3dde92a}
+\label{sec:org7b1d020}
 \begin{itemize}

### 2022-02-14 manuscript/manuscript.tex
wrote manuscript

% Created 2022-02-14 Mon 09:23
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[version=3]{mhchem}
\usepackage[numbers,super,sort&compress]{natbib}
\usepackage{natmove}
\usepackage{url}
\usepackage{minted}
\usepackage{underscore}
\usepackage[linktocpage,pdfstartview=FitH,colorlinks,
linkcolor=blue,anchorcolor=blue,
citecolor=blue,filecolor=blue,menucolor=blue,urlcolor=blue]{hyperref}
\usepackage{attachfile}
\usepackage{geometry}
\geometry{margin=1.0in,top=.75in,bottom=.75in}
\usepackage{graphicx}
\usepackage{color}
\usepackage[numbers,super,sort&compress]{natbib}
\usepackage{caption}
\usepackage{subcaption}
\captionsetup{font=footnotesize}
\usepackage[version=3]{mhchem}
\usepackage{siunitx}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{mdwlist}
\usepackage{hyperref}
\pagestyle{fancy}
\usepackage{wrapfig}
\usepackage{nopageno}
\fancyhf{}
\fancyhead[LE,RO]{\scriptsize Jerry Crum}
\fancyhead[RE,LO]{\scriptsize ZSE Outline}
%\fancyfoot[CE,CO]{\leftmark}
\fancyfoot[LE,RO]{\thepage}
%\usepackage{subfig}
\usepackage{comment}
\usepackage{titlesec}
\titlespacing*{\section}
{0pt}{0.6\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsection}
{0pt}{0.6\baselineskip}{0.2\baselineskip}
\titlespacing*{\subsubsection}
{0pt}{0.4\baselineskip}{0.1\baselineskip}
\usepackage{parskip}
\usepackage[section]{placeins}
\usepackage{siunitx}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\green}[1]{\textcolor{green}{#1}}
\newcommand{\orange}[1]{\textcolor{orange}{#1}}
\usepackage[capitalise]{cleveref}
\author{Jerry T. Crum\(^{\text{a}}\), Justin R. Crum\(^{\text{b}}\), Cameron Taylor\(^{\text{a}}\), William F. Schneider\(^{\text{a,c}}\)}
\date{}
\title{ZSE Manuscript Outline---Needs a Title!!!!}
\begin{document}

\begin{OPTIONS}
\def\udesoftecoverride\#1\mainmatter\{\%
  \AfterEndPreamble{#1\mainmatter}
\end{OPTIONS}

\maketitle

\begin{asparaenum}[\expandafter\textsuperscript a ]
\item Department of Chemical and Biolmolecular Engineering, University of Notre Dame, 250 Nieuwland Science Hall, Notre Dame, IN 46556, USA \\
\item Department of Applied Mathematics, University of Arizona, 617 N Santa Rita Ave, Tucson, AZ 85721, USA\\
\item Department of Chemistry and Biochemistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
\end{asparaenum}

\newpage

\section{Introduction}
\label{sec:orgf4a8429}
Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges. The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}. It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal. Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite. Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/ring-examples2.pdf}
\caption{Cutout of the Chabazite framework showing a path (5-6-7-8-9) highlighted with purple bonds, a cycle (3-4-18-19-20-17) highlighted with blue bonds, an 8-MR filled in with pink, and a 4-MR filled in with green. Yellow atoms are Si (T-sites), and red atoms are oxygen. \label{fig:3.1}}
\end{figure}
\end{figure}

Many researchers have presented methods to define and count the rings present in zeolites \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001}. Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges. In some cases researchers will use T-sites as nodes and oxygen atoms as edges, since the oxygen atoms are only connected to two T-sites \cite{goetzke-properties-1991}. Some basic definitions of graphs that will be used going forward are highlighted in a cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}, and described in \cref{tab:graphs}. The size of each of these features can either be defined by the total number of atoms contained in them, or by the number of T-sites contained in them. The latter method is the standard convention which we will use. A path is a sequence of edges which joins a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}). A cycle is a path that starts and ends at the same node, and does not repeat any other nodes (shown by blue bonds in \cref{fig:3.1}). In the most basic form a ring is a cycle that does not contain any shortcuts \cite{guttman-ring-1990,goetzke-properties-1991}. \cref{fig:3.1} shows an 8-membered ring and a 4-MR filled in with pink and green respectively.

\begin{table}
\centering
\begin{threeparttable}
\caption{List of graph based features, their descriptions, and how they apply to frameworks, T-sites, and oxygen atoms. \label{tab:graphs}}
{\tiny
\begin{tabular}{lp{2cm}p{2cm}p{2cm}p{2cm}}
\hline
Feature & Description & Application to Frameworks & Application to T-sites & Application to Oxygen Atoms\\
\hline
Node & T-site or oxygen atom & Contains some set of symmetry distinct nodes &  & \\
Path & A sequence of nodes that are connected to each other & Traverses through a framework & T-sites and oxygen atoms connected in series forms a path & T-sites and oxygen atoms connected in series forms a path\\
Cycle & A path that starts and ends at the same node, and does not repeat any other nodes & Contains some set of cycles & Formed by alternating connected T-sites and oxygen atoms & Formed by alternating connected T-sites and oxygen atoms\\
Rings & A cycle that does not contain a shortcut \cite{guttman-ring-1990,goetzke-properties-1991} & Can describe channels of a framework, or other distinct void environments & Can be described by counting the number of rings that pass through it & Can be described by the number of rings that pass through it\\
Modified Shortcut Rings & A cycle that does not contain a modified shortcut & A subset of the rings will be found in the framework & Used to describe a T-site & Used to describe an oxygen atom\\
Shortest Path Rings & A ring that is the shortest ring for at one least one set of O-T-O along the cycle & Another subset of the rings of a framework & Used to describe a T-site & Used to describe an oxygen atom\\
Vertex Symbol Rings & The rings making up the vertex symbol of a T-site & Yet another subset of the rings of the framework & Used to describe a T-site & Not applicable to oxygen atoms\\
\hline
\end{tabular}

\begin{tablenotes}
}
\end{tablenotes}
\end{threeparttable}
\end{table}

A shortcut is defined as a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}, or in other words, the path is a shortcut if it is shorter than both paths between these nodes along the cycle. An example is the path connecting T17 and T18 in \cref{fig:3.1}, which is a shortcut of the blue cycle. This path contains two T-sites, splitting the 8-MR cycle into two 4-MRs. Many definitions build off this idea that rings do not contain shortcuts. In a zeolite, rings are used to define the topology of the framework, such as the channels and void environments that are present. Counting the rings that pass through a T-site can also be used to describe and differentiate the local void environment around symmetry distinct T-sites. Like T-sites, ring counts can be used to differentiate symmetry distinct oxygen atoms.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/cha-vertex.pdf}
\caption{Cutout of the CHA framework, showing the rings that make up the vertex symbol of the single symmetry distinct T-site. a) Example of the tetrahedron formed by the T-site and four connected oxygen atoms, with labeled edges of the tetrahedron. b) Rings associated with opposite edges E1 and E3. c) Rings associated with opposite edges E5 and E6. d) Ring associated with opposited edges E2 and E4. Rings are colored as: 4-MR (green), 6-MR (blue), and 8-MR (pink). \label{fig:cha-vertex}}
\end{figure}
\end{figure}

Vertex symbols are a common way to describe the T-sites of a zeolite framework, and were first used to do so in 1997 by O'Keeffe and Hyde in 1997 \cite{okeeffe-vertex-1997}. The vertex symbol of a T-site contains the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms. The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together. These grouped pairs are listed from smallest to largest forming the vertex symbol. \cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework. The edges of the tetrahedron are labeled to aid in identification. \cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges. The vertex symbol of T1 in CHA is thus able to be determined as 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8. For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol. An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.

Another ring counting convention presented by Sastre and Corma is to count only the shortest path connecting an O-T-O in a framework. With this definition, they can find and count all the rings in a framework that are the shortest path for at least one set of O-T-O along the cycle \cite{sastre-topological-2009}. Using AFI as an example again (\cref{fig:3.3}), we see that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom. The 12-MR would not be included because for every O-T-O along the 12-MR, the shortest path connecting them is not the 12-MR. The difference between this method and the vertex symbol rings is subtle, but with this shortest path convention any ring belonging to the vertex symbol of any T-site in a framework will be included in the ring count for each of the T-sites that ring passes through. \cref{fig:3.4} shows a cutout of the TON framework including a 6- (blue) and 10-MR (orange). For T1, only the 6-MR would be counted in the vertex symbol because it is the shortest path connecting O2 and O3. The 10-MR is part of the vertex symbol for T3 because it is the shortest path connecting O2 and O14. Since this 10-MR is the shortest path for at least one set of O-T-O along the ring, and passes through T1, it does get counted in the shortest path rings for T1.

\begin{figure}
\begin{figure}[!H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/afi-12-6.pdf}
\caption{Cutout of the 12-MR channel in AFI highlighting a 12-MR in purple, and a 6-MR in blue. The 6-MR is included in the vertex symbol of labeled T1 because it is the shortest path connecting O1 and O2. The 12-MR would not be included in the vertex symbol or shortest path ring list because for each O-T-O along the 12-MR there is a shorter path connecting them. \label{fig:3.3}}
\end{figure}
\end{figure}

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/ton-6-10.pdf}
\caption{Cutout of the TON framework showing a 6- (blue) and 10-MR (orange). The 10-MR is the shortest path connecting O14-T3-O2, and passes through T1, so it is counted in the shortest path rings for T1. \label{fig:3.4}}
\end{figure}
\end{figure}

We can use ring counts to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites. Since various conventions exist that can reduce the set of rings in a zeolite to more strictly defined properties, the ring counts returned by the various conventions will differ. Differences in ring counts leads to differences in how we might describe the topological environment of a zeolite. Therefore, when using rings to determine the properties of a framework, T-site, or oxygen atom, it is important to know the difference in the conventions, and use one that determines the features of interest.

Here we present the Zeolite Simulation Environment (ZSE, \url{https://wwww.github.com/jtcrum/zse}), a Python package that implements the ring finding algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} to find rings up to a user defined cutoff size, and can implement the previously published ring set reduction conventions. We use ZSE to provide an analysis of rings returned from each conventions on the entire set of IZA Database of zeolite frameworks \cite{baerlocher-database-nodate} to compare how they results in different characterizations. We highlight rings that are found by these conventions but not typically discussed in the literature for a number of frameworks. We also show that the vertex symbol, a common approach used to characterize T-sites misses important parts of the stereochemistry around a T-site.

\section{Methods}
\label{sec:org347f423}

\subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
\label{sec:org942aa65}

In this work we implement an efficient algorithm that was presented by Goetzke and Klein to find all the rings associated with a T-site that do not contain a shortcut \cite{goetzke-properties-1991} in a Python package called the Zeolite Simulation Environment (ZSE). In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures. All graph theory functions are performed used the NetworkX Python package \cite{hagberg-exploring-2008}.

First, we convert the ASE atoms object into a connectivity matrix which represents every atom across the columns and rows. If two atoms are bound together, their respective entry in the connectivity matrix contains a 1, else a 0. This connectivity matrix is then converted to a NetworkX graph object, and then a distance dictionary using NetworkX built in functions. Then we implement Step 3 from Geotzke and Klein's algorithm \cite{goetzke-properties-1991} summarized here: to find the rings that pass through a T-site, we iteratively search for every size ring between 3-MR and a maximum ring value that is user specified. For this work we set a cutoff of 18-MRs. A schematic showing the evolution of the ring search is shown in \cref{fig:goetzke}. For ring size \(\lambda\) we start at the T-site of interest (labeled 1) \cref{fig:goetzke}, and search the distance matrix for any T-sites that are \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (for odd \(\lambda\)) distance from the starting T-site (labeled 2). Next we attempt to create to distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by FIGURE. Each node added to each of the paths must be \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (odd \(\lambda\)) from the head of the other path. Also each node added to each path needs to be the correct distance from 1 and 2 for the given step respectively. If either of the previous two conditions are not met, a ring cannot be formed of length \(\lambda\) along the given paths, we backtrack and repeat until all possible options have been explored for \(\lambda\). Then we increase \(\lambda\) and continue until the cutoff ring size is completed.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/goetzke.pdf}
\caption{Diagram showing how the ring finding algorithm evolves. Adapted from Goetzke and Klein \cite{goetzke-properties-1991}. \label{fig:goetzke}}
\end{figure}
\end{figure}

\subsection{Finding Vertex Symbol Rings \label{section:vertex}}
\label{sec:orgf265857}

Starting from the set of all rings found in \cref{section:goetzke}, we can prune the ring list to the set of vertex symbol rings. We find the shortest ring in the set that connects each pair of oxygens bound to our initial T-site. It is possible for there to be multiple rings of the same size connecting each oxygen, in which case all the rings of that size are kept.

\subsection{Finding Shortest Path Rings}
\label{sec:org50f9a65}

Here we prune the set of all rings from \cref{section:goetzke} to a subset of rings the meets the shortest path definition published by Sastre and Corma \cite{sastre-topological-2009}. For each ring, we iterate over every group of O-T-O atoms in the ring, and check if this ring is the shortest path connecting the two oxygen atoms. If so, the loop is broke, because the ring need only be the shortest path for one group of O-T-O atoms to fit the definition. This is the most time consuming process out of all the ring finding conventions we have implemented.

\subsection{Modified Shortcut Rings}
\label{sec:orgde5abba}

In this work we present a modified definition of shortcut to capture a different subset of rings from any of the other ring finding conventions. Traditionally a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}. We propose that a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into at least one smaller ring. This definition does not require that the shortcut between two nodes be shorter than both paths connecting those nodes along the cycle. Our shortcut only needs to be shorter than one of the paths connecting those two nodes along the cycle.

This definition is explained graphically in \cref{fig:mod}, where we present a cutout of the AFI framework showing a portion of the 12-MR channel. There is a 14-MR that traverses seven T-sites in each of the 12-MRs, through the combination of Path 1 (blue) and Path 2 (purple). With the classical definition of a shortcut, this cycle is considered a ring. However, Path 3 (in green) connecting T1 and T2 produces a 12-MR when combined with Path 1, making Path 3 a modified shortcut. This 14-MR would not be counted under our new definition.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/modified-shortcut.pdf}
\caption{Cutout of the AFI framework showing two stacked 12-MRs from the channel. A 14-MR is shown as T-sites replaced with aluminum atoms in gray. The two paths making connecting Al 1 and Al2 that make this 14-MR are highlighted with blue and purple bonds. Path 3 highlighted with green bonds is a modified shortcut connecting Al1 and Al2. \label{fig:mod}}
\end{figure}
\end{figure}

To remove rings containing modified shortcuts from the full set of rings, we iterate over every T-site pair of the ring and check for the shortest path connecting them. If that shortest path is shorter than either of the two paths along the ring connecting the two T-sites, we check if the combination of this shorter path and the shorter of the two ring paths forms a new smaller ring. If so the iteration is broken, and the ring is removed from the counted set.

\subsection{Ordered Vertex Symbols \label{section:ov}}
\label{sec:org678cf59}

To add information about the spatial orientation of the rings around a T-site to the vertex symbol, we have developed a method to order the edges in the vertex symbol. We systematically list the rings by following the edges of the tetrahedron such that each ring listed is connected to one of the oxygens of the next ring listed. After removing all the rings that are not a part of the vertex symbol (\cref{section:vertex}) we use the following process to order them.

\begin{enumerate}
\item List all the possible arrangements of the oxygens bound to the T-site (4! = 24 possible arrangements).
\item Use a predetermined order of edges: [[0,2],[0,1],[1,2],[2,3],[3,0],[1,3]].
\begin{enumerate}
\item Where each of those values represents the index of the oxygen to use.
\end{enumerate}
\item Find the ring size (and multiplicity) connecting each pair of oxygens in this predetermined order.
\item Make a list of weights, where for each pair of oxygens the weight is the ring size \texttimes{} multiplicity.
\item Reverse sort the list of all possible oxygen arrangements by the correlating list of weights.
\item Use the first oxygen arrangement coupled with the predetermined edge order to list the rings and multiplicity for each edge.
\end{enumerate}

\subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
\label{sec:org0e83149}

Finally, to determine all the ring sizes exhibited with in a zeolite framework, we take advantage of T-site symmetry. The rings of a framework are made of T-sites, and if two T-sites are symmetrically identical they will have the same set of rings passing through them. Therefore, we only need to find the rings associated with each symmetry distinct T-sites to know all the possible ring sizes within a framework. For example, AFI only contains one symmetry distinct T-site (T1). Using the basic definition of a shortcut, T1 is a part of 4-, 6-, 12-, and 14-MRs when using a cutoff of 18-MR. Every other T-site in the AFI framework is also a T1, thus the only possible ring sizes in AFI are 4-, 6-, 12-, and 14-MRs.

\section{Results}
\label{sec:orgdb1ba9c}
\subsection{Characterizing Rings in a Zeolite Graph}
\label{sec:org07b595e}
The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site. These rings listed by the IZA are referred to as tabulated rings in the literature \cite{curtis-statistical-2003}. In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the list of tabulated rings. These 'untabulated' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site. \cref{fig:fw-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database \cite{baerlocher-database-nodate}. There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/rings-vs-iza-rings.pdf}
\caption{Counts of IZA frameworks containing each size ring between 3- and 18-MR using the Goetzke algorithm and the tabulated rings listed by the IZA \cite{baerlocher-database-nodate}. \label{fig:fw-counts}}
\end{figure}
\end{figure}

Taking a closer look at some of these untabulated rings, highlights rings not typically listed for some frameworks, but still relevant to describing their topology. Using CHA as an example, \cref{fig:cha-rings} displays a 12-MR (in purple) that exists in CHA that circumferences the cage. This ring is not associated with the vertex symbol of the single symmetry distinct T-site in CHA and does not define a channel. Thus, this ring is not included in the list of tabulated rings. We would argue that this is still a ring that provides an important topological descriptor of CHA because none of the tabulated rings provides information about the size of the CHA cage.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/cha-all-rings.pdf}
\caption{Chabazite cage and double 6-MR (D6R) with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple. The 8-MR in the D6R and the 12-MR are rings not typically discussed in literature. Si atoms have been replaced with Al atoms to help identify those rings in the overal cage strcuture. \label{fig:cha-rings}}
\end{figure}
\end{figure}

Using AFI as another example, we find another type of ring that arises from traversing a pair of stacked rings and is not included in the list of tabulated rings. AFI, like CHA, contains one symmetry distinct T-site. According to the IZA, the AFI framework contains 4-, 6-, and 12-MRs \cite{baerlocher-database-nodate}. When we search for rings using the Goetzke algorithm \cite{goetzke-properties-1991}, we also find that it contains 14-MRs created by using seven T-sites from two 12-MRs that are separated by a distance of one oxygen (\cref{fig:afi-14}). Rings of this nature are prevalent in many frameworks, another example can be seen in the bottom right of \cref{fig:cha-rings}, where an 8-MR is highlighted traversing the two 6-MRs of the D6R. These types of rings may not be of interest depending on which topological feature one intends to describe. This has led us to create a modified definition of a shortcut as explained in METHODS REFERENCE, which excludes these types of rings. The benefit of this new shortcut definition is that larger rings that are missed by the vertex symbol or shortest path rings (i.e., 12-MR in AFI) are still captured, while excluding rings that arise from convolution of staked rings found with the typical shortcut definition.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/afi-14.pdf}
\caption{Cutout of the 12-MR channel in AFI with a 14-MR (yellow) traversing seven T-sites of each 12-MR. The T-sites comprising the 14-MR have been replaced with Al for visibility. \label{fig:afi-14}}
\end{figure}
\end{figure}

\subsection{Characterizing Frameworks by Rings}
\label{sec:org9ac1d1a}
With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database. \cref{fig:ring-counts} shows how many frameworks contain each size ring found using the various ring counting conventions from 3- to 18-MRs. This plot highlights the differences in the conventions and shows that a topological description of a framework based on rings will depend on the way that you define a ring. In general, a hierarchy of ring sizes found by each convention is: all rings not containing a shortcut \textgreater this work \textgreater shortest path rings \textgreater vertex symbol rings. While the IZA listed rings includes the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/ring-counts-2.pdf}
\caption{Number of IZA zeolite frameworks containing each size ring, using the various ring counting conventions, as well as the rings listed by the IZA Database \cite{baerlocher-database-nodate}. \label{fig:ring-counts}}
\end{figure}
\end{figure}

One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of cycles that don't fit this definition, but still exhibit properties like rings. This is a trade-off between well-defined connectivity rules, and the inclusion of particular void environments that may still have important applications. These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings. One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}. Similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}. \cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/MFI-6MC.pdf}
\caption{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green. The 6-membered cycle would not be found by any of the connectivity ring rules outlined in this work. \label{fig:mfi-6}}
\end{figure}
\end{figure}

\subsection{Characterizing T-sites by Rings}
\label{sec:org362568a}

Considering that zeolite frameworks are comprised of one or more symmetry distinct T-sites, it may be of interest to describe those T-sites by the rings that pass through them. Most often the vertex symbol is used to make such a classification \cite{okeeffe-vertex-1997}. Sastre and Corma also provided characterization of T-sites using the shortest path rings that pass through them \cite{sastre-topological-2009}. In their work, they presented the ring index, which lists all the rings passing through a T-site from smallest to largest, and a subscript for each size representing it's multiplicity.  The rings associated with a T-site can provided information about the local void environments around the T-site, and could potentially be correlated to other physicochemical properties of the T-site once enough once sufficient data on those physicochemical properties exists.

Take for example the AFI framework, containing one symmetry distinct T-site. AFI contains 4-, 6-, 12-, and 14-MRs. To describe that T-site we can count how many of each of those rings pass through the T-site. We can also prune this list using our modified definition of a shortcut, the shortest path rings definition \cite{sastre-topological-2009}, or the rings contained within the vertex symbol of this T-site \cite{okeeffe-vertex-1997}. Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}:
\begin{itemize}
\item Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\)
\item This work: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
\item Shortest Path Rings: 4\(\cdot\)6\(_{\text{13}}\)
\item Vertex Symbol Rings: 4\(\cdot\)6\(_{\text{11}}\)
\end{itemize}

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/afi-funnel.pdf}
\caption{Diagram shwoing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
\end{figure}
\end{figure}

With an understanding of how we characterize T-site by counting the rings that pass through, \cref{tab:uninodal} shows the ring index for a selection of T-sites from uninodal (containing only one symmetry distinct T-site) frameworks. This table highlights the differences in rings counts found with each convention, and shows that in general as you move from left to right across the table the largest ring found decreases. The results in the shortest path column were found using ZSE, but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}. The results in the vertex symbol rings column were also found with zse, and agree directly with the vertex symbols listed on the IZA Database website \cite{baerlocher-database-nodate}.

\begin{table}
\centering
\begin{threeparttable}
\caption{Comparison of ring indices for the T-sites in various uninodal zeolite frameworks. \label{tab:uninodal}}
{\scriptsize
\begin{tabular}{lllll}
\hline
Framework & Rings & This Work & Shortest Path Rings \cite{sastre-topological-2009} & Vertex Symbol Rings \cite{baerlocher-database-nodate}\\
\hline
ABW & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{2}}\)\\
ACO & 4\(_{\text{3}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{15}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\)\\
AFI & 4\(_{\text{1}} \cdot\)\(_{\text{13}} \cdot\)12\(_{\text{1}} \cdot\)14\(_{\text{7}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{11}}\)\\
ANA & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{8}}\)\\
ATO & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)8\(_{\text{8}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\)\\
BCT & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{12}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
CHA & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
DFT & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}} \cdot\)10\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{4}} \cdot\)8\(_{\text{6}}\)\\
GIS & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
GME & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{7}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
MER & 4\(_{\text{3}} \cdot\)8\(_{\text{4}} \cdot\)10\(_{\text{10}} \cdot\)14\(_{\text{14}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
MON & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{4}} \cdot\)8\(_{\text{4}}\)\\
NPO & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
\hline
\end{tabular}
\begin{tablenotes}
Vertex symbols have been represented in ring index format for ease of comparison.
}
\end{tablenotes}
\end{threeparttable}
\end{table}

Next, we take an in-depth look at the ring counts for a framework with multiple symmetry distinct T-sites, to show how a ring index can provide information about the local environment around a T-site, and help differentiate them. MOZ is a zeolite framework containing 4-, 6-, 8-, 10-, 12-, 14-, and 18-MRs, 6 symmetry distinct T-sites, and two distinct 12-MR channels. \cref{tab:moz} shows the ring index for each T-site using each ring finding method.

\begin{table}
\centering
\begin{threeparttable}
\caption{Ring indices for each distinct T-site in the MOZ framework using each ring counting convention. \label{tab:moz}}
{\scriptsize
\begin{tabular}{lllll}
\hline
T-Site & Rings & This Work & Shortest Path Rings & Vertex Symbol Rings\\
\hline
T1 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)18\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
T2 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
T3 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{5}} \cdot\)10\(_{\text{4}} \cdot\)12\(_{\text{4}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(\cdot\)12\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
T4 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{6}} \cdot\)12\(\cdot\)18\(_{\text{26}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12\\
T5 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{6}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\)\\
T6 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)10\(_{\text{2}} \cdot\)12\(_{\text{8}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)12\(_{\text{8}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\)\\
\hline
\end{tabular}
\begin{tablenotes}
Vertex symbols have been represented in ring index format for ease of comparison.
}
\end{tablenotes}
\end{threeparttable}
\end{table}

\cref{fig:moz} shows the T-site locations inside a 2-dimensional view of the framework. If you were interested in which T-sites have access to the 12-MR channels, the shortest path rings and vertex symbol rings would only suggest T3 participates in the 12-MR rings. However, all rings and this work both identify T4 and T6 as participating in the 12-MR channels as highlighted in \cref{fig:moz}.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/moz.pdf}
\caption{Cutout of the MOZ framework showing two 12-MR channels, with an example of each distinct T-site highlighted. T1: navy, T2: green, T3: orange, T4: purple, T5: blue, and T6: red. As shown, T3, T4, and T6 are all associted with the 12-MR channels, while T1, T2, and T5 are not connected to the 12-MR channels. \label{fig:moz}}
\end{figure}
\end{figure}

We next used ZSE to find the rings associated with every symmetry distinct T-site in every framework across the IZA Database using each of the four ring counting conventions. For each T-site we used the rings to generate a ring index, and \cref{fig:unique-ts} shows how many unique ring indices are present when using each of the ring counting conventions. The plot follows intuition with the number of unique ring indices decreasing as we use more restrictive ring counting conventions, because less rings are found and provides less room for differentiation. This raises the question, if you want to ascertain chemical or physical properties about a T-site based on its ring count, and differentiate these T-sites from other similar but distinct T-site, which ring counting convention will suffice? The answer will depend on what level of detail is desired. Larger rings can be found with the standard shortcut definition, rings traversing other stacked rings will be excluded with our modified shortcut definition. The shortest path definition and vertex symbol rings will provide the most localized information about a T-site.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/unique-ts.pdf}
\caption{Number of unique T-site ring indices when classified by the rings passing through them using the various ring counting conventions. There are 1460 T-sites across all the frameworks in the IZA Database. As we move from less restrictive to more restrictive (left to right) ring counting conventions, the number of unique ring indices decreases. \label{fig:unique-ts}}
\end{figure}
\end{figure}

To further compare the ring counting conventions, we show a distribution of the number of T-sites containing each size ring between 3- and 18-MR in \cref{fig:tsite-frequency} (right). This plot highlights that more T-sites contain larger sized rings when using the basic definition of a shortcut, and at smaller rings sizes (\textless6-MR) all the ring counting conventions return the same results.  To further emphasize this results, we have provided a cumulative distribution of the same data normalized to the maximum 'rings' value in \cref{fig:tsite-frequency}. At 6-MRs is where we see the cumulative distribution functions deviate from each other. The largest deviation takes place at 12-MRs, and the cumulative distributions start to level out at larger ring sizes.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/dist-cumudist.pdf}
\caption{Frequency of T-sites accross all IZA frameworks containing ring sizes between 3- and 18-MR (left), and cumulative distribution of T-sites containing each ring size normalized to the final 'rings' value (right). \label{fig:tsite-frequency}}
\end{figure}
\end{figure}

To complete the comparison of ring counting conventions, we have developed method to determine how similar the ring counts of one convention are to another. We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}. In this equation, sr is the number of similar rings that are found in both counting conventions, and mr is the maximum number of rings found by either convention. For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\) and 4\(\cdot\)6\(_{\text{13}} \cdot\). The number of similar rings found by both conventions is 14, and the maximum number of rings found by either convention is 22. This would lead to a similarity score of 0.636. We do this for every T-site between two conventions and average the similarity score to get the results in \cref{fig:similarity}. Down the diagonal each method is compared to itself and clearly has a similarity of 1. The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compare to the least restrictive convention (rings) has the lowest similarity score. The two most similar ring counting methods are our modified shortcut definition and the shortest path rings.

\begin{equation}\label{eq:similarity}
s = \mathrm{ \frac{sr}{mr} }
\end{equation}

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/similarity-heat-map.pdf}
\caption{Heat map showing the similarity score for four ring counting methods. Similarity score of 1 means identical set of rings returned, while a similarity of 0 would mean no matching rings are returned. \label{fig:similarity}}
\end{figure}
\end{figure}

One final point we would like to make about T-site characterization, is that the vertex symbol and ring indices only provide information about rings, and do not give any information about the spatial orientation of those rings around the T-site. The vertex symbol comes close to accomplishing that, by finding rings associated with opposite edge pairs of the tetrahedron, and then listing the those pairs from smallest to largest. However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments. For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\). However, the orientation of those rings around each of those T-sites are not identical. \cref{fig:stereo} shows a cutout of each of these frameworks that only includes the atoms that make up the rings of the vertex symbol around the specified T-sites. We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1. The main difference is the location of the 5\(_{\text{2}}\)- and 4-MR edges. They are highlighted in \cref{fig:stereo}.

\begin{figure}
\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{figures/chapter-3/stereo.pdf}
\caption{Cutout of the MOR, EON, and MON frameworks that only shows the rings associated with the vertex symbol of T3, T9, and T1 respectively. The 4-MR (green) and 2\texttimes{} 5-MR (teal) that are in swapped positions are highlighted for emphasis. The 4-MR for MON, and the 2\timex 5-MRs for MOR and EON are into the plane, and not easily shown. \label{fig:stereo}}
\end{figure}
\end{figure}

The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation. This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings.

DEFINE ORDERED VERTEX SYMBOL HERE

With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5. The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol. Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about such as deprotonation energy, T-site substitution energy, or catalytic properties of reactions happening at that T-site.

We have used this new ordered vertex symbol to characterize all 1460 T-sites in the IZA Database. We found that using the standard vertex symbol to characterize T-sites there are 649 unique vertex symbols present. In contrast, there are 666 unique ordered vertex symbols across every T-site. This would imply that not a large amount of vertex symbols contain stereochemical differences. In \cref{tab:ov} we provide a list of some common T-site vertex symbols, and their representative ordered vertex symbols.

\begin{table}
\centering
\begin{threeparttable}
\caption{List of three vertex symbols, the T-sites associated with them, and the representative ordered vertex symbol for those T-sites. \label{tab:ov}}
{\scriptsize
\begin{tabular}{lll}
\hline
Vertex Symbol & Framework T-site & Ordered Vertex Symbol\\
\hline
5\(\cdot\)5\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)10 & EWS T3 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & ITN T9 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & ITN T21 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & OKO T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & OKO T5 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & PCS T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & PCS T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & SFS T10 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & SFV T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & SFV T7 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 & TUN T10 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
\hline
4\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)8\(\cdot\)5\(\cdot\)8 & DAC T3 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 & DAC T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 & EON T10 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 & EPI T1 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 & MOR T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 & RSN T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
 & VNI T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
 & VSV T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
 & YFI T9 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
\hline
4\(\cdot\)6\(\cdot\)4\(\cdot\)6\(\cdot\)6\(\cdot\)8 & ATN T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
 & JSN T3 & 8\(\cdot\)6\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\\
 & PON T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
 & SAS T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
 & ZON T3 & 8\(\cdot\)6\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\\
\hline
\end{tabular}
\begin{tablenotes}
}
\end{tablenotes}
\end{threeparttable}
\end{table}
\section{Conclusions}
\label{sec:orge6f2ecc}
Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}. We find that the commonly reported ring sizes in literature and on the IZA website leave out many rings that fit the classical definition of a cycle that does not contain a shortcut. To completely describe the topology of a zeolite these rings are required, however there are often cases where someone might want to consider only a subset of rings of interest.

We have shown a comparison of three different conventions used to count rings, and highlighted the differences in rings that are found by each convention. The classic definition of a ring identifies the largest set of ring sizes across all the zeolite frameworks, while the shortest path ring and vertex symbol rings only identify smaller ring sizes. We have provided a modified definition of a shortcut that when used to find rings still finds larger rings defining channel openings, but excludes rings that are able to be decomposed into at least one smaller ring. It is important to understand the difference of ring sizes and types found by each convention when discussing the rings of a zeolite framework. A disadvantage to using purely connectivity based definitions of rings is the exclusion of cycles in a framework that behave physicochemically like a ring but contain a shortcut. We have displayed an example case of these geometric rings, and in the future it would be beneficial to develop a computation method of identifying these cycles.

This same methodology was used to describe T-sites in zeolite frameworks by counting all of the rings that pass through the T-site using each of the ring counting conventions described. When using ring finding conventions that find larger rings, we see more diversity in the descriptions of T-sites, which can aid researchers who want to identify similar T-sites across multiple frameworks. We have also shown that the vertex symbol used to describe T-sites leaves out subtle but distinct stereochemical differences in the spatial orientation of the rings around a T-site. To address this shortcoming we have provided a new method for ordering the rings of a vertex symbol that takes into consideration the ring stereochemistry and is able to identify differences in T-sites that have the same vertex symbol. In the future, correlating physicochemical properties of T-sites to the ring descriptors identified with each ring counting convention can help identify sets of frameworks with desired T-site properties.

\section{Acknowledgments}
\label{sec:org0cb7f0a}

We acknowledge financial support provided by the National Science Foundation under Cooperative Agreement No. EEC-1647722, which is an Engineering Research Center for the Innovative and Strategic Transformation of Alkane Resources. JTC thanks the Arthur J. Schmitt Foundation for financial aid in the form of a PhD fellowship. We thank Dr. Christian Baerlocher for the numerous discussions about the topology of zeolites, and methodologies used by the IZA Structure Database. We thank Dr. German Sastre for providing a copy of zeoTsites to compare results, and for the helpful conversions about use cases and zeolite topology. This research was supported in part by the Notre Dame center for Research Computing through access to high performance computing clusters. 
\backmatter

\bibliographystyle{nddiss2e}
\bibliography{ref}
\end{document}

### 2022-02-14 outline/outline.tex
wrote manuscript

@@
-% Created 2022-01-22 Sat 12:56
+% Created 2022-02-05 Sat 09:46
 % Intended LaTeX compiler: pdflatex
@@
 \date{}
-\title{ZSE Manuscript Outline}
+\title{ZSE Manuscript Outline---Needs a Title!!!!}
 \begin{document}
@@
 \section{Abstract}
-\label{sec:org14a7c72}
+\label{sec:orgb67b8b5}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 \section{Highlights}
-\label{sec:orga9a0704}
+\label{sec:orgf8787e1}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section{Introduction}
-\label{sec:org9bb9b10}
+\label{sec:org38f6217}
 
@@
 \section{Software Description}
-\label{sec:org2e90fc2}
+\label{sec:org6a04a45}
 
@@
 
-\section{Results}
-\label{sec:orge3c702c}
-
+\section{Results and Discussion}
+\label{sec:org2e12605}
+\subsection{Characterizing rings in a zeolite graph}
+\label{sec:org9599689}
 \textbf{\textbf{Paragraph 1: IZA doesn't list all rings in a framework}}
@@
 
+\subsection{Characterizing frameworks by rings}
+\label{sec:org90a7fd6}
 \textbf{\textbf{Paragraph 4: Comparing all 4 ringing finding conventions and IZA listed rings by counting the number of frameworks containing each size ring}}
@@
 \item While the IZA lists all vertex symbol rings, and a selection of general rings
+\item \red{How many unique combinations of ring sizes within all known zeolites?}
 \end{itemize}
+
 \begin{center}
@@
 \end{center}
+
+\subsection{Characterizing T-sites by rings}
+\label{sec:org41f4a0a}
+Motivate this\ldots{}.
 
@@
 
-\section{Conclusions}
-\label{sec:org2135ac7}
+\section{Conclusions---Needs to be updated}
+\label{sec:org999ea72}
 \begin{itemize}
+\item \red{Conclusions need to be connected to motivations and results above.
+Should connect to points made in abstract.
+Why do I care about rings in a framework?
+What are their practical significance?}
+
 \item \red{Rings of graph are well defined; here identify all rings up to XXX in YYY frameworks. Find that commonly reported (IZA) ring sizes miss certain rings.}
+
+\item Can characterize Fws by
 \end{itemize}
@@
 \section{Acknowledgments}
-\label{sec:org7b1d020}
+\label{sec:org8d20140}
 \begin{itemize}

### 2022-02-14 outline/outline.tex
added title

@@
-% Created 2022-02-05 Sat 09:46
+% Created 2022-02-14 Mon 09:26
 % Intended LaTeX compiler: pdflatex
@@
 \date{}
-\title{ZSE Manuscript Outline---Needs a Title!!!!}
+\title{Characterization and Analysis of Ring Topology of Zeolite Frameworks}
 \begin{document}
@@
 \section{Abstract}
-\label{sec:orgb67b8b5}
+\label{sec:org12f50fd}
 The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
@@
 \section{Highlights}
-\label{sec:orgf8787e1}
+\label{sec:org87be40d}
 What is the main point of the paper? What problem is it addressing? What are the main insights?
@@
 \section{Introduction}
-\label{sec:org38f6217}
+\label{sec:org54979aa}
 
@@
 \section{Software Description}
-\label{sec:org6a04a45}
+\label{sec:orge438e7d}
 
@@
 \section{Results and Discussion}
-\label{sec:org2e12605}
+\label{sec:orgf17e1d7}
 \subsection{Characterizing rings in a zeolite graph}
-\label{sec:org9599689}
+\label{sec:orga4651b7}
 \textbf{\textbf{Paragraph 1: IZA doesn't list all rings in a framework}}
@@
 \subsection{Characterizing frameworks by rings}
-\label{sec:org90a7fd6}
+\label{sec:org8b6927e}
 \textbf{\textbf{Paragraph 4: Comparing all 4 ringing finding conventions and IZA listed rings by counting the number of frameworks containing each size ring}}
@@
 \subsection{Characterizing T-sites by rings}
-\label{sec:org41f4a0a}
+\label{sec:org98bec9d}
 Motivate this\ldots{}.
@@
 \section{Conclusions---Needs to be updated}
-\label{sec:org999ea72}
+\label{sec:orgd9045f5}
 \begin{itemize}
@@
 \section{Acknowledgments}
-\label{sec:org8d20140}
+\label{sec:orgf859058}
 \begin{itemize}

### 2022-02-14 manuscript/manuscript.tex
added title

@@
-% Created 2022-02-14 Mon 09:23
+% Created 2022-02-14 Mon 09:29
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgf4a8429}
+\label{sec:org69cd8cc}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \section{Methods}
-\label{sec:org347f423}
+\label{sec:org4e3a1e8}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org942aa65}
+\label{sec:org3a6a8dc}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:orgf265857}
+\label{sec:orgc034067}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org50f9a65}
+\label{sec:orge66ba34}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:orgde5abba}
+\label{sec:org34a7c1c}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org678cf59}
+\label{sec:org71add99}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org0e83149}
+\label{sec:orgd45fa37}
 
@@
 \section{Results}
-\label{sec:orgdb1ba9c}
+\label{sec:orgf83bcff}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org07b595e}
+\label{sec:org4f9d4d6}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:org9ac1d1a}
+\label{sec:orgbdafedc}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:org362568a}
+\label{sec:orgc865d95}
 
@@
 \section{Conclusions}
-\label{sec:orge6f2ecc}
+\label{sec:org95528f4}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org0cb7f0a}
+\label{sec:org4b5ad85}
 

### 2022-04-11 elsman/manuscript.tex
added elsevier format

--- 2022-02-14 manuscript/manuscript.tex
@@
-% Created 2022-02-14 Mon 09:29
+% Created 2022-04-11 Mon 09:50
 % Intended LaTeX compiler: pdflatex
-\documentclass[11pt]{article}
+\documentclass[final,numrefs,noinfo,sort&compress]{elsarticle}
   \usepackage[utf8]{inputenc}
-\usepackage{lmodern}
-\usepackage[T1]{fontenc}
-\usepackage{fixltx2e}
+\usepackage{url}
+\usepackage[version=4]{mhchem}
+\usepackage{chemmacros}[2016/05/02]
 \usepackage{graphicx}
+\usepackage{float}
+\usepackage{color}
+\usepackage{adjustbox}
+\usepackage{amsmath}
+\usepackage{siunitx}
+\usepackage{textcomp}
+\usepackage{wasysym}
+\usepackage{latexsym}
+\usepackage{amssymb}
+\usepackage{lineno}
+\usepackage{chemformula}
+\usepackage{xr}
+\usepackage{pifont}
 \usepackage{longtable}
-\usepackage{float}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{marvosym}
-\usepackage{wasysym}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage[version=3]{mhchem}
-\usepackage[numbers,super,sort&compress]{natbib}
-\usepackage{natmove}
-\usepackage{url}
+\usepackage[section]{placeins}
+\usepackage{threeparttable}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+\chemsetup{formula = mhchem ,modules = {reactions,thermodynamics}}
+\usepackage{xcolor}
+\chemsetup[reactions]{tag-open= ( , tag-close = )}
+\usepackage[capitalize]{cleveref}
 \usepackage{minted}
-\usepackage{underscore}
-\usepackage[linktocpage,pdfstartview=FitH,colorlinks,
-linkcolor=blue,anchorcolor=blue,
-citecolor=blue,filecolor=blue,menucolor=blue,urlcolor=blue]{hyperref}
-\usepackage{attachfile}
-\usepackage{geometry}
-\geometry{margin=1.0in,top=.75in,bottom=.75in}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage[numbers,super,sort&compress]{natbib}
-\usepackage{caption}
-\usepackage{subcaption}
-\captionsetup{font=footnotesize}
-\usepackage[version=3]{mhchem}
-\usepackage{siunitx}
-\usepackage{fancyhdr}
-\usepackage{paralist}
-\usepackage{amsmath}
-\usepackage{enumitem}
-\usepackage{mdwlist}
-\usepackage{hyperref}
-\pagestyle{fancy}
-\usepackage{wrapfig}
-\usepackage{nopageno}
-\fancyhf{}
-\fancyhead[LE,RO]{\scriptsize Jerry Crum}
-\fancyhead[RE,LO]{\scriptsize ZSE Outline}
-%\fancyfoot[CE,CO]{\leftmark}
-\fancyfoot[LE,RO]{\thepage}
-%\usepackage{subfig}
-\usepackage{comment}
-\usepackage{titlesec}
-\titlespacing*{\section}
-{0pt}{0.6\baselineskip}{0.2\baselineskip}
-\titlespacing*{\subsection}
-{0pt}{0.6\baselineskip}{0.2\baselineskip}
-\titlespacing*{\subsubsection}
-{0pt}{0.4\baselineskip}{0.1\baselineskip}
-\usepackage{parskip}
-\usepackage[section]{placeins}
-\usepackage{siunitx}
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-\newcommand{\blue}[1]{\textcolor{blue}{#1}}
-\newcommand{\green}[1]{\textcolor{green}{#1}}
-\newcommand{\orange}[1]{\textcolor{orange}{#1}}
-\usepackage[capitalise]{cleveref}
-\author{Jerry T. Crum\(^{\text{a}}\), Justin R. Crum\(^{\text{b}}\), Cameron Taylor\(^{\text{a}}\), William F. Schneider\(^{\text{a,c}}\)}
+\DeclareSIUnit\Td{Td}
+\setminted[python]{frame=lines,fontsize=\scriptsize,xleftmargin=\parindent,linenos,breaklines}
 \date{}
-\title{Characterization and Analysis of Ring Topology of Zeolite Frameworks}
+\title{}
 \begin{document}
 
-\begin{OPTIONS}
-\def\udesoftecoverride\#1\mainmatter\{\%
-  \AfterEndPreamble{#1\mainmatter}
-\end{OPTIONS}
-
+\frontmatter
+\title{Characterization and Analysis of Ring Topology of Zeolite Frameworks}
+\author[nd]{Jerry T. Crum}
+\author[uoa]{Justin R. Crum}
+\author[nd]{Cameron Taylor}
+\author[nd,ndc]{William F. Schneider}
+\address[nd]{Department of Chemical and Biomoledcular Engineering, University of Notre Dame, 250 Nieuwland Science Hall, Notre Dame, IN 46556, USA}
+\address[uoa]{Department of Applied Mathematics, University of Arizona, 617 N Santa Rita Ave, Tucson, AZ 85721, USA}
+\address[ndc]{Department of Chemistry and Biochemistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA}
+\begin{abstract}
+The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
+These ring descriptors have been used to identify feasible zeolite topologies, to describe the similarity and differences between zeolites, to identify sites or voids of catalytic relevance, and as machine learning fingerprints.
+Numerous definitions and algorithms for finding zeolite rings have been proposed and applied throughout the literature.
+Here we report an analysis of rings and T-sites in a large number of zeolite frameworks using Zeolite Simulation Environment, a Python package that implements an efficient algorithm presented by Goetzke and Klein for finding rings in arbitrary frameworks.
+We compare the result of a number of common and new ring definitions applied to a large number of common zeolite frameworks.
+We discover previously unrecognized rings in a number of frameworks.
+We show that the vertex symbol, a common approach used to characterize T-sites, misses important parts of the stereochemistry around a T-site, and propose an alternative definition.
+This tool provides an effective platform for characterizing zeolite and T-site structures useful for building models and doing machine learning.
+\end{abstract}
 \maketitle
 
-\begin{asparaenum}[\expandafter\textsuperscript a ]
-\item Department of Chemical and Biolmolecular Engineering, University of Notre Dame, 250 Nieuwland Science Hall, Notre Dame, IN 46556, USA \\
-\item Department of Applied Mathematics, University of Arizona, 617 N Santa Rita Ave, Tucson, AZ 85721, USA\\
-\item Department of Chemistry and Biochemistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA
-\end{asparaenum}
-
-\newpage
-
+\mainmatter
 \section{Introduction}
-\label{sec:org69cd8cc}
+\label{sec:orgd888b04}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \section{Methods}
-\label{sec:org4e3a1e8}
+\label{sec:org29c0bff}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org3a6a8dc}
+\label{sec:org924f108}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:orgc034067}
+\label{sec:org0ba8199}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:orge66ba34}
+\label{sec:org888dd54}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:org34a7c1c}
+\label{sec:orgfbdc2ae}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org71add99}
+\label{sec:org6ef5356}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:orgd45fa37}
+\label{sec:org48d80de}
 
@@
 \section{Results}
-\label{sec:orgf83bcff}
+\label{sec:orgea2aae3}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org4f9d4d6}
+\label{sec:orge3221a8}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:orgbdafedc}
+\label{sec:org3c516bd}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:orgc865d95}
+\label{sec:orga218832}
 
@@
 \section{Conclusions}
-\label{sec:org95528f4}
+\label{sec:org2639d47}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org4b5ad85}
+\label{sec:orga5cd5e7}
 
@@
 
-\bibliographystyle{nddiss2e}
+\bibliographystyle{elsarticle-num}
 \bibliography{ref}

### 2022-04-11 manuscript/manuscript.tex
added elsevier format

@@
-% Created 2022-02-14 Mon 09:29
+% Created 2022-04-11 Mon 09:05
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org69cd8cc}
+\label{sec:orgb31d910}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \section{Methods}
-\label{sec:org4e3a1e8}
+\label{sec:org477e3e7}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org3a6a8dc}
+\label{sec:orgc1b0da0}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:orgc034067}
+\label{sec:org97ea47e}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:orge66ba34}
+\label{sec:orge9011a0}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:org34a7c1c}
+\label{sec:orgf75517d}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org71add99}
+\label{sec:org75405ff}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:orgd45fa37}
+\label{sec:org11fc0d6}
 
@@
 \section{Results}
-\label{sec:orgf83bcff}
+\label{sec:org371918d}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org4f9d4d6}
+\label{sec:org3b507b3}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:orgbdafedc}
+\label{sec:orgead59d0}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:orgc865d95}
+\label{sec:orgab44aee}
 
@@
 \section{Conclusions}
-\label{sec:org95528f4}
+\label{sec:orgac48af2}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org4b5ad85}
+\label{sec:org868a098}
 

### 2022-07-24 elsman/manuscript.tex
new revisions

@@
-% Created 2022-04-11 Mon 09:50
+% Created 2022-07-23 Sat 18:33
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgd888b04}
+\label{sec:org19dcf8c}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \begin{table}
-\centering
 \begin{threeparttable}
@@
 
-Here we present the Zeolite Simulation Environment (ZSE, \url{https://wwww.github.com/jtcrum/zse}), a Python package that implements the ring finding algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} to find rings up to a user defined cutoff size, and can implement the previously published ring set reduction conventions.
-We use ZSE to provide an analysis of rings returned from each conventions on the entire set of IZA Database of zeolite frameworks \cite{baerlocher-database-nodate} to compare how they results in different characterizations.
+Here we present an analysis of rings captured by Goetzke and Klein's efficient ring finding algorithm \cite{goetzke-properties-1991}, and compare those rings to the rings found by other previously published ring set reduction conventions.
+We have implemented all of these ring finding conventions in a Python package called the Zeolite Simulation Environment (ZSE, \url{https://www.github.com/jtcrum/zse}).
+We use ZSE to provide an analysis of rings captured by each convention for the entire set of zeolite frameworks contained on the IZA Database \cite{baerlocher-database-nodate} to compare how these sets of rings provide different characterizations of said frameworks.
 We highlight rings that are found by these conventions but not typically discussed in the literature for a number of frameworks.
-We also show that the vertex symbol, a common approach used to characterize T-sites misses important parts of the stereochemistry around a T-site.
+We also show that the vertex symbol, a common approach used to characterize T-sites \cite{okeeffe-vertex-1997}, based on the shortest rings connecting the neighboring oxygen, misses important parts of the stereochemistry around a T-site.
+Finally, we provide an alternative method for listing the vertex symbol rings that takes into account their orientation and connectivity around the T-site..
 
 \section{Methods}
-\label{sec:org29c0bff}
+\label{sec:org6054a21}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org924f108}
+\label{sec:org08fc5d2}
 
@@
 For ring size \(\lambda\) we start at the T-site of interest (labeled 1) \cref{fig:goetzke}, and search the distance matrix for any T-sites that are \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (for odd \(\lambda\)) distance from the starting T-site (labeled 2).
-Next we attempt to create to distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by FIGURE.
+Next we attempt to create to distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by \cref{fig:goetzke}.
 Each node added to each of the paths must be \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (odd \(\lambda\)) from the head of the other path.
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org0ba8199}
+\label{sec:orgb6fb9da}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org888dd54}
+\label{sec:org5531fca}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:orgfbdc2ae}
+\label{sec:org1c61cf5}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org6ef5356}
+\label{sec:orgb76cef5}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org48d80de}
+\label{sec:org860dd8f}
 
@@
 \section{Results}
-\label{sec:orgea2aae3}
+\label{sec:org6ea0f4c}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:orge3221a8}
+\label{sec:org361aea8}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:org3c516bd}
+\label{sec:orgac790ce}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:orga218832}
+\label{sec:org31bfe20}
 
@@
 \section{Conclusions}
-\label{sec:org2639d47}
+\label{sec:org2ae0775}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:orga5cd5e7}
+\label{sec:org0f0b307}
 

### 2022-08-29 elsman/manuscript.tex
fixed reference to methods section:

@@
-% Created 2022-07-23 Sat 18:33
+% Created 2022-08-28 Sun 20:43
 % Intended LaTeX compiler: pdflatex
-\documentclass[final,numrefs,noinfo,sort&compress]{elsarticle}
+\documentclass[preprint,numrefs,noinfo,sort&compress]{elsarticle}
   \usepackage[utf8]{inputenc}
@@
 \section{Introduction}
-\label{sec:org19dcf8c}
+\label{sec:org357df7f}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \section{Methods}
-\label{sec:org6054a21}
+\label{sec:orgd626155}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org08fc5d2}
+\label{sec:org18323f6}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:orgb6fb9da}
+\label{sec:org6489a90}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org5531fca}
+\label{sec:org79b8ee2}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:org1c61cf5}
+\label{sec:orgab747f4}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:orgb76cef5}
+\label{sec:org0ad3d6b}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org860dd8f}
+\label{sec:orge68662a}
 
@@
 \section{Results}
-\label{sec:org6ea0f4c}
+\label{sec:orgd3657a1}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org361aea8}
+\label{sec:org6ba0672}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:orgac790ce}
+\label{sec:org0da4ee6}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:org31bfe20}
+\label{sec:org5362aa3}
 
@@
 The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
-This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings.
-
-DEFINE ORDERED VERTEX SYMBOL HERE
+This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings described in \cref{section:ov}.
 
@@
 \section{Conclusions}
-\label{sec:org2ae0775}
+\label{sec:orgc2d46e4}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org0f0b307}
+\label{sec:orgfa56224}
 

### 2022-08-29 elsman/manuscript.tex
fixed reference

@@
-% Created 2022-08-28 Sun 20:43
+% Created 2022-08-29 Mon 07:25
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org357df7f}
+\label{sec:org2a67457}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \begin{figure}
-\begin{figure}[!H]
+\begin{figure}[H]
 \centering
@@
 \section{Methods}
-\label{sec:orgd626155}
+\label{sec:org6431034}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org18323f6}
+\label{sec:org42ddcba}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org6489a90}
+\label{sec:org20e1097}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org79b8ee2}
+\label{sec:org42016a9}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:orgab747f4}
+\label{sec:orgd5be89b}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org0ad3d6b}
+\label{sec:org8f45030}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:orge68662a}
+\label{sec:org79d764e}
 
@@
 \section{Results}
-\label{sec:orgd3657a1}
+\label{sec:orgcd96db0}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org6ba0672}
+\label{sec:org15e507d}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:org0da4ee6}
+\label{sec:orgb71506d}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:org5362aa3}
+\label{sec:orge809831}
 
@@
 \section{Conclusions}
-\label{sec:orgc2d46e4}
+\label{sec:org0be7bec}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:orgfa56224}
+\label{sec:org805a78b}
 

### 2022-10-03 elsman/manuscript.tex
submitted

@@
-% Created 2022-08-29 Mon 07:25
+% Created 2022-10-03 Mon 12:32
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org2a67457}
+\label{sec:org7af04cb}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol.
-An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.
+An example would be the vertex symbol of 4\(\cdot\)6\textsubscript{2}\(\cdot\)6\(\cdot\)6\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)6\textsubscript{3} for T1 in AFI.
 
@@
 \section{Methods}
-\label{sec:org6431034}
+\label{sec:org83ee037}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org42ddcba}
+\label{sec:orge9b7691}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org20e1097}
+\label{sec:org89c7798}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org42016a9}
+\label{sec:org0530c72}
 
@@
 \subsection{Modified Shortcut Rings}
-\label{sec:orgd5be89b}
+\label{sec:org7cfe2eb}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org8f45030}
+\label{sec:org54f88e7}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org79d764e}
+\label{sec:orgb8842aa}
 
@@
 \section{Results}
-\label{sec:orgcd96db0}
+\label{sec:org3113701}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org15e507d}
+\label{sec:org11cd53a}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:orgb71506d}
+\label{sec:orgc427fb9}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings.
-One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}.
-Similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}.
+One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\textsuperscript{2+} uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}.
+Similar to Co\textsuperscript{2+} uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}.
 \cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:orge809831}
+\label{sec:orgc94a591}
 
@@
 \begin{itemize}
-\item Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\)
-\item This work: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
-\item Shortest Path Rings: 4\(\cdot\)6\(_{\text{13}}\)
-\item Vertex Symbol Rings: 4\(\cdot\)6\(_{\text{11}}\)
+\item Rings: 4\(\cdot\)6\textsubscript{13}\(\cdot\)12\(\cdot\)14\textsubscript{7}
+\item This work: 4\(\cdot\)6\textsubscript{13}\(\cdot\)12
+\item Shortest Path Rings: 4\(\cdot\)6\textsubscript{13}
+\item Vertex Symbol Rings: 4\(\cdot\)6\textsubscript{11}
 \end{itemize}
@@
 \hline
-ABW & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{2}}\)\\
-ACO & 4\(_{\text{3}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{15}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\)\\
-AFI & 4\(_{\text{1}} \cdot\)\(_{\text{13}} \cdot\)12\(_{\text{1}} \cdot\)14\(_{\text{7}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{11}}\)\\
-ANA & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{8}}\)\\
-ATO & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)8\(_{\text{8}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\)\\
-BCT & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{12}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
-CHA & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
-DFT & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}} \cdot\)10\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{4}} \cdot\)8\(_{\text{6}}\)\\
-GIS & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
-GME & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{7}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
-MER & 4\(_{\text{3}} \cdot\)8\(_{\text{4}} \cdot\)10\(_{\text{10}} \cdot\)14\(_{\text{14}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
-MON & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{4}} \cdot\)8\(_{\text{4}}\)\\
-NPO & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
+ABW & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{2}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{2}\\
+ACO & 4\textsubscript{3}\(\cdot\)6\textsubscript{3}\(\cdot\)8\textsubscript{6}\(\cdot\)10\textsubscript{15} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6} & 4\textsubscript{3}\(\cdot\)8\textsubscript{6}\\
+AFI & 4\textsubscript{1}\(\cdot\)\textsubscript{13}\(\cdot\)12\textsubscript{1}\(\cdot\)14\textsubscript{7} & 4\textsubscript{1}\(\cdot\)6\textsubscript{13}\(\cdot\)12\textsubscript{1} & 4\textsubscript{1}\(\cdot\)6\textsubscript{13} & 4\textsubscript{1}\(\cdot\)6\textsubscript{11}\\
+ANA & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{16} & 4\textsubscript{2}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{8}\\
+ATO & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\(\cdot\)8\textsubscript{8}\(\cdot\)12\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\(\cdot\)12\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9} & 4\textsubscript{1}\(\cdot\)6\textsubscript{9}\\
+BCT & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{20} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{12} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6} & 4\textsubscript{1}\(\cdot\)6\textsubscript{6}\\
+CHA & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{6}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\\
+DFT & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10}\(\cdot\)10\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{6}\(\cdot\)8\textsubscript{10} & 4\textsubscript{2}\(\cdot\)6\textsubscript{4}\(\cdot\)8\textsubscript{6}\\
+GIS & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\\
+GME & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{6}\(\cdot\)12\textsubscript{7} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\(\cdot\)12\textsubscript{1} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2} & 4\textsubscript{3}\(\cdot\)6\textsubscript{1}\(\cdot\)8\textsubscript{2}\\
+MER & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\(\cdot\)10\textsubscript{10}\(\cdot\)14\textsubscript{14} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4} & 4\textsubscript{3}\(\cdot\)8\textsubscript{4}\\
+MON & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{5}\(\cdot\)8\textsubscript{6} & 4\textsubscript{1}\(\cdot\)5\textsubscript{4}\(\cdot\)8\textsubscript{4}\\
+NPO & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)12\textsubscript{40} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\(\cdot\)12\textsubscript{40} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6} & 3\textsubscript{1}\(\cdot\)6\textsubscript{6}\\
 \hline
@@
 \hline
-T1 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)18\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T2 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T3 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{5}} \cdot\)10\(_{\text{4}} \cdot\)12\(_{\text{4}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(\cdot\)12\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T4 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{6}} \cdot\)12\(\cdot\)18\(_{\text{26}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12\\
-T5 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{6}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\)\\
-T6 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)10\(_{\text{2}} \cdot\)12\(_{\text{8}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)12\(_{\text{8}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\)\\
+T1 & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{7}\(\cdot\)10\textsubscript{7}\(\cdot\)18\textsubscript{5} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{3} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{3} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\\
+T2 & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{7}\(\cdot\)10\textsubscript{7}\(\cdot\)14\textsubscript{5} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{3} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{3} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\\
+T3 & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\textsubscript{5}\(\cdot\)10\textsubscript{4}\(\cdot\)12\textsubscript{4}\(\cdot\)14\textsubscript{5} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\(\cdot\)12\textsubscript{4} & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8 & 4\textsubscript{3}\(\cdot\)6\textsubscript{2}\(\cdot\)8\\
+T4 & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{6}\(\cdot\)10\textsubscript{6}\(\cdot\)12\(\cdot\)18\textsubscript{26} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{6}\(\cdot\)12 & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{6}\(\cdot\)12 & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{6}\(\cdot\)12\\
+T5 & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{7}\(\cdot\)10\textsubscript{6}\(\cdot\)14\textsubscript{18} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{7} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{7} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{7}\\
+T6 & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{3}\(\cdot\)10\textsubscript{2}\(\cdot\)12\textsubscript{8}\(\cdot\)14\textsubscript{18} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{3}\(\cdot\)12\textsubscript{8} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{3} & 4\textsubscript{2}\(\cdot\)6\(\cdot\)8\textsubscript{3}\\
 \hline
@@
 In this equation, sr is the number of similar rings that are found in both counting conventions, and mr is the maximum number of rings found by either convention.
-For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\) and 4\(\cdot\)6\(_{\text{13}} \cdot\).
+For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\textsubscript{13}\(\cdot\)12\(\cdot\)14\textsubscript{7} and 4\(\cdot\)6\textsubscript{13}\(\cdot\).
 The number of similar rings found by both conventions is 14, and the maximum number of rings found by either convention is 22.
@@
 However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments.
-For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\).
+For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\textsubscript{2}\(\cdot\)5\(\cdot\)8\textsubscript{2}\(\cdot\)5\(\cdot\)8\textsubscript{2}.
 However, the orientation of those rings around each of those T-sites are not identical.
@@
 We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1.
-The main difference is the location of the 5\(_{\text{2}}\)- and 4-MR edges.
+The main difference is the location of the 5\textsubscript{2}- and 4-MR edges.
 They are highlighted in \cref{fig:stereo}.
@@
 
-With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
+With this new descriptor MOR T3 and EON T9 would be labeled as: 8\textsubscript{2}\(\cdot\)8\textsubscript{2}\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\textsubscript{2}\(\cdot\)8\textsubscript{2}\(\cdot\)4\(\cdot\)5\(\cdot\)5\textsubscript{2}\(\cdot\)5.
 The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol.
@@
 \hline
-5\(\cdot\)5\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)10 & EWS T3 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & ITN T9 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & ITN T21 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & OKO T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & OKO T5 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & PCS T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & PCS T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFS T10 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFV T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFV T7 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & TUN T10 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+5\(\cdot\)5\(\cdot\)5\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)10 & EWS T3 & 10\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & ITN T9 & 10\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & ITN T21 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & OKO T2 & 10\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & OKO T5 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & PCS T2 & 10\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & PCS T3 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & SFS T10 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & SFV T3 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & SFV T7 & 5\textsubscript{2}\(\cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
+ & TUN T10 & 10\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
 \hline
-4\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)8\(\cdot\)5\(\cdot\)8 & DAC T3 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & DAC T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & EON T10 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & EPI T1 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & MOR T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & RSN T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & VNI T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & VSV T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & YFI T9 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+4\(\cdot\)5\textsubscript{2}\(\cdot\)5\(\cdot\)8\(\cdot\)5\(\cdot\)8 & DAC T3 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+ & DAC T4 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+ & EON T10 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+ & EPI T1 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+ & MOR T4 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
+ & RSN T4 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
+ & VNI T2 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
+ & VSV T2 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
+ & YFI T9 & 5\textsubscript{2}\(\cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
 \hline
@@
 \section{Conclusions}
-\label{sec:org0be7bec}
+\label{sec:org3c5b493}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org805a78b}
+\label{sec:org5f43fe4}
 

### 2022-10-03 highlights.tex
submitted

--- 2022-08-29 elsman/manuscript.tex
@@
-% Created 2022-08-29 Mon 07:25
+% Created 2022-10-03 Mon 11:34
 % Intended LaTeX compiler: pdflatex
-\documentclass[preprint,numrefs,noinfo,sort&compress]{elsarticle}
-  \usepackage[utf8]{inputenc}
-\usepackage{url}
-\usepackage[version=4]{mhchem}
-\usepackage{chemmacros}[2016/05/02]
-\usepackage{graphicx}
-\usepackage{float}
-\usepackage{color}
-\usepackage{adjustbox}
-\usepackage{amsmath}
-\usepackage{siunitx}
-\usepackage{textcomp}
-\usepackage{wasysym}
-\usepackage{latexsym}
-\usepackage{amssymb}
-\usepackage{lineno}
-\usepackage{chemformula}
-\usepackage{xr}
-\usepackage{pifont}
-\usepackage{longtable}
-\usepackage[section]{placeins}
-\usepackage{threeparttable}
-\newcommand{\red}[1]{\textcolor{red}{#1}}
-\chemsetup{formula = mhchem ,modules = {reactions,thermodynamics}}
-\usepackage{xcolor}
-\chemsetup[reactions]{tag-open= ( , tag-close = )}
-\usepackage[capitalize]{cleveref}
-\usepackage{minted}
-\DeclareSIUnit\Td{Td}
-\setminted[python]{frame=lines,fontsize=\scriptsize,xleftmargin=\parindent,linenos,breaklines}
-\date{}
+\documentclass{elsarticle}
+  \usepackage[left=1in, right=1in, top=1in, bottom=1in, nohead]{geometry}
+\author{William F. Schneider}
+\date{\today}
 \title{}
@@
 
-\frontmatter
+\begin{OPTIONS:}
+\end{OPTIONS:}
+
 \title{Characterization and Analysis of Ring Topology of Zeolite Frameworks}
@@
 \address[ndc]{Department of Chemistry and Biochemistry, University of Notre Dame, 251 Nieuwland Science Hall, Notre Dame, IN 46556, USA}
-\begin{abstract}
-The topology of zeolite frameworks and of associated tetrahedral sites (T-sites) are commonly characterized by their associated rings, typically defined as some set of closed paths or cycles through a framework that cannot be decomposed into shorter cycles.
-These ring descriptors have been used to identify feasible zeolite topologies, to describe the similarity and differences between zeolites, to identify sites or voids of catalytic relevance, and as machine learning fingerprints.
-Numerous definitions and algorithms for finding zeolite rings have been proposed and applied throughout the literature.
-Here we report an analysis of rings and T-sites in a large number of zeolite frameworks using Zeolite Simulation Environment, a Python package that implements an efficient algorithm presented by Goetzke and Klein for finding rings in arbitrary frameworks.
-We compare the result of a number of common and new ring definitions applied to a large number of common zeolite frameworks.
-We discover previously unrecognized rings in a number of frameworks.
-We show that the vertex symbol, a common approach used to characterize T-sites, misses important parts of the stereochemistry around a T-site, and propose an alternative definition.
-This tool provides an effective platform for characterizing zeolite and T-site structures useful for building models and doing machine learning.
-\end{abstract}
 \maketitle
 
-\mainmatter
-\section{Introduction}
-\label{sec:org2a67457}
-Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
-The International Zeolite Association (IZA) lists over 200 known zeolite frameworks that are described by dimensionality, pore shape and size, and Si/Al ratios \cite{baerlocher-database-nodate}.
-It is natural to characterize differences in zeolites by the size and shapes of the features present in the crystal.
-Rings are one common type of feature widely reported and used, both to characterize a zeolite crystal and as descriptors of the individual tetrahedral sites (T-sites) of a zeolite.
-Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ring-examples2.pdf}
-\caption{Cutout of the Chabazite framework showing a path (5-6-7-8-9) highlighted with purple bonds, a cycle (3-4-18-19-20-17) highlighted with blue bonds, an 8-MR filled in with pink, and a 4-MR filled in with green.
-Yellow atoms are Si (T-sites), and red atoms are oxygen.
-\label{fig:3.1}}
-\end{figure}
-\end{figure}
-
-Many researchers have presented methods to define and count the rings present in zeolites \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001}.
-Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges.
-In some cases researchers will use T-sites as nodes and oxygen atoms as edges, since the oxygen atoms are only connected to two T-sites \cite{goetzke-properties-1991}.
-Some basic definitions of graphs that will be used going forward are highlighted in a cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}, and described in \cref{tab:graphs}.
-The size of each of these features can either be defined by the total number of atoms contained in them, or by the number of T-sites contained in them.
-The latter method is the standard convention which we will use.
-A path is a sequence of edges which joins a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}).
-A cycle is a path that starts and ends at the same node, and does not repeat any other nodes (shown by blue bonds in \cref{fig:3.1}).
-In the most basic form a ring is a cycle that does not contain any shortcuts \cite{guttman-ring-1990,goetzke-properties-1991}.
-\cref{fig:3.1} shows an 8-membered ring and a 4-MR filled in with pink and green respectively.
-
-\begin{table}
-\begin{threeparttable}
-\caption{List of graph based features, their descriptions, and how they apply to frameworks, T-sites, and oxygen atoms. \label{tab:graphs}}
-{\tiny
-\begin{tabular}{lp{2cm}p{2cm}p{2cm}p{2cm}}
-\hline
-Feature & Description & Application to Frameworks & Application to T-sites & Application to Oxygen Atoms\\
-\hline
-Node & T-site or oxygen atom & Contains some set of symmetry distinct nodes &  & \\
-Path & A sequence of nodes that are connected to each other & Traverses through a framework & T-sites and oxygen atoms connected in series forms a path & T-sites and oxygen atoms connected in series forms a path\\
-Cycle & A path that starts and ends at the same node, and does not repeat any other nodes & Contains some set of cycles & Formed by alternating connected T-sites and oxygen atoms & Formed by alternating connected T-sites and oxygen atoms\\
-Rings & A cycle that does not contain a shortcut \cite{guttman-ring-1990,goetzke-properties-1991} & Can describe channels of a framework, or other distinct void environments & Can be described by counting the number of rings that pass through it & Can be described by the number of rings that pass through it\\
-Modified Shortcut Rings & A cycle that does not contain a modified shortcut & A subset of the rings will be found in the framework & Used to describe a T-site & Used to describe an oxygen atom\\
-Shortest Path Rings & A ring that is the shortest ring for at one least one set of O-T-O along the cycle & Another subset of the rings of a framework & Used to describe a T-site & Used to describe an oxygen atom\\
-Vertex Symbol Rings & The rings making up the vertex symbol of a T-site & Yet another subset of the rings of the framework & Used to describe a T-site & Not applicable to oxygen atoms\\
-\hline
-\end{tabular}
-
-\begin{tablenotes}
-}
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-A shortcut is defined as a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}, or in other words, the path is a shortcut if it is shorter than both paths between these nodes along the cycle.
-An example is the path connecting T17 and T18 in \cref{fig:3.1}, which is a shortcut of the blue cycle.
-This path contains two T-sites, splitting the 8-MR cycle into two 4-MRs.
-Many definitions build off this idea that rings do not contain shortcuts.
-In a zeolite, rings are used to define the topology of the framework, such as the channels and void environments that are present.
-Counting the rings that pass through a T-site can also be used to describe and differentiate the local void environment around symmetry distinct T-sites.
-Like T-sites, ring counts can be used to differentiate symmetry distinct oxygen atoms.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/cha-vertex.pdf}
-\caption{Cutout of the CHA framework, showing the rings that make up the vertex symbol of the single symmetry distinct T-site.
-a) Example of the tetrahedron formed by the T-site and four connected oxygen atoms, with labeled edges of the tetrahedron.
-b) Rings associated with opposite edges E1 and E3.
-c) Rings associated with opposite edges E5 and E6.
-d) Ring associated with opposited edges E2 and E4.
-Rings are colored as: 4-MR (green), 6-MR (blue), and 8-MR (pink).
-\label{fig:cha-vertex}}
-\end{figure}
-\end{figure}
-
-Vertex symbols are a common way to describe the T-sites of a zeolite framework, and were first used to do so in 1997 by O'Keeffe and Hyde in 1997 \cite{okeeffe-vertex-1997}.
-The vertex symbol of a T-site contains the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms.
-The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together.
-These grouped pairs are listed from smallest to largest forming the vertex symbol.
-\cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework.
-The edges of the tetrahedron are labeled to aid in identification.
-\cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges.
-The vertex symbol of T1 in CHA is thus able to be determined as 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8.
-For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol.
-An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.
-
-Another ring counting convention presented by Sastre and Corma is to count only the shortest path connecting an O-T-O in a framework.
-With this definition, they can find and count all the rings in a framework that are the shortest path for at least one set of O-T-O along the cycle \cite{sastre-topological-2009}.
-Using AFI as an example again (\cref{fig:3.3}), we see that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom.
-The 12-MR would not be included because for every O-T-O along the 12-MR, the shortest path connecting them is not the 12-MR.
-The difference between this method and the vertex symbol rings is subtle, but with this shortest path convention any ring belonging to the vertex symbol of any T-site in a framework will be included in the ring count for each of the T-sites that ring passes through.
-\cref{fig:3.4} shows a cutout of the TON framework including a 6- (blue) and 10-MR (orange).
-For T1, only the 6-MR would be counted in the vertex symbol because it is the shortest path connecting O2 and O3.
-The 10-MR is part of the vertex symbol for T3 because it is the shortest path connecting O2 and O14.
-Since this 10-MR is the shortest path for at least one set of O-T-O along the ring, and passes through T1, it does get counted in the shortest path rings for T1.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/afi-12-6.pdf}
-\caption{Cutout of the 12-MR channel in AFI highlighting a 12-MR in purple, and a 6-MR in blue.
-The 6-MR is included in the vertex symbol of labeled T1 because it is the shortest path connecting O1 and O2.
-The 12-MR would not be included in the vertex symbol or shortest path ring list because for each O-T-O along the 12-MR there is a shorter path connecting them.
-\label{fig:3.3}}
-\end{figure}
-\end{figure}
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ton-6-10.pdf}
-\caption{Cutout of the TON framework showing a 6- (blue) and 10-MR (orange).
-The 10-MR is the shortest path connecting O14-T3-O2, and passes through T1, so it is counted in the shortest path rings for T1.
-\label{fig:3.4}}
-\end{figure}
-\end{figure}
-
-We can use ring counts to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites.
-Since various conventions exist that can reduce the set of rings in a zeolite to more strictly defined properties, the ring counts returned by the various conventions will differ.
-Differences in ring counts leads to differences in how we might describe the topological environment of a zeolite.
-Therefore, when using rings to determine the properties of a framework, T-site, or oxygen atom, it is important to know the difference in the conventions, and use one that determines the features of interest.
-
-Here we present an analysis of rings captured by Goetzke and Klein's efficient ring finding algorithm \cite{goetzke-properties-1991}, and compare those rings to the rings found by other previously published ring set reduction conventions.
-We have implemented all of these ring finding conventions in a Python package called the Zeolite Simulation Environment (ZSE, \url{https://www.github.com/jtcrum/zse}).
-We use ZSE to provide an analysis of rings captured by each convention for the entire set of zeolite frameworks contained on the IZA Database \cite{baerlocher-database-nodate} to compare how these sets of rings provide different characterizations of said frameworks.
-We highlight rings that are found by these conventions but not typically discussed in the literature for a number of frameworks.
-We also show that the vertex symbol, a common approach used to characterize T-sites \cite{okeeffe-vertex-1997}, based on the shortest rings connecting the neighboring oxygen, misses important parts of the stereochemistry around a T-site.
-Finally, we provide an alternative method for listing the vertex symbol rings that takes into account their orientation and connectivity around the T-site..
-
-\section{Methods}
-\label{sec:org6431034}
-
-\subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org42ddcba}
-
-In this work we implement an efficient algorithm that was presented by Goetzke and Klein to find all the rings associated with a T-site that do not contain a shortcut \cite{goetzke-properties-1991} in a Python package called the Zeolite Simulation Environment (ZSE).
-In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures.
-All graph theory functions are performed used the NetworkX Python package \cite{hagberg-exploring-2008}.
-
-First, we convert the ASE atoms object into a connectivity matrix which represents every atom across the columns and rows.
-If two atoms are bound together, their respective entry in the connectivity matrix contains a 1, else a 0.
-This connectivity matrix is then converted to a NetworkX graph object, and then a distance dictionary using NetworkX built in functions.
-Then we implement Step 3 from Geotzke and Klein's algorithm \cite{goetzke-properties-1991} summarized here: to find the rings that pass through a T-site, we iteratively search for every size ring between 3-MR and a maximum ring value that is user specified.
-For this work we set a cutoff of 18-MRs.
-A schematic showing the evolution of the ring search is shown in \cref{fig:goetzke}.
-For ring size \(\lambda\) we start at the T-site of interest (labeled 1) \cref{fig:goetzke}, and search the distance matrix for any T-sites that are \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (for odd \(\lambda\)) distance from the starting T-site (labeled 2).
-Next we attempt to create to distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by \cref{fig:goetzke}.
-Each node added to each of the paths must be \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (odd \(\lambda\)) from the head of the other path.
-Also each node added to each path needs to be the correct distance from 1 and 2 for the given step respectively.
-If either of the previous two conditions are not met, a ring cannot be formed of length \(\lambda\) along the given paths, we backtrack and repeat until all possible options have been explored for \(\lambda\).
-Then we increase \(\lambda\) and continue until the cutoff ring size is completed.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/goetzke.pdf}
-\caption{Diagram showing how the ring finding algorithm evolves. Adapted from Goetzke and Klein \cite{goetzke-properties-1991}. \label{fig:goetzke}}
-\end{figure}
-\end{figure}
-
-\subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org20e1097}
-
-Starting from the set of all rings found in \cref{section:goetzke}, we can prune the ring list to the set of vertex symbol rings.
-We find the shortest ring in the set that connects each pair of oxygens bound to our initial T-site.
-It is possible for there to be multiple rings of the same size connecting each oxygen, in which case all the rings of that size are kept.
-
-\subsection{Finding Shortest Path Rings}
-\label{sec:org42016a9}
-
-Here we prune the set of all rings from \cref{section:goetzke} to a subset of rings the meets the shortest path definition published by Sastre and Corma \cite{sastre-topological-2009}.
-For each ring, we iterate over every group of O-T-O atoms in the ring, and check if this ring is the shortest path connecting the two oxygen atoms.
-If so, the loop is broke, because the ring need only be the shortest path for one group of O-T-O atoms to fit the definition.
-This is the most time consuming process out of all the ring finding conventions we have implemented.
-
-\subsection{Modified Shortcut Rings}
-\label{sec:orgd5be89b}
-
-In this work we present a modified definition of shortcut to capture a different subset of rings from any of the other ring finding conventions.
-Traditionally a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}.
-We propose that a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into at least one smaller ring.
-This definition does not require that the shortcut between two nodes be shorter than both paths connecting those nodes along the cycle.
-Our shortcut only needs to be shorter than one of the paths connecting those two nodes along the cycle.
-
-This definition is explained graphically in \cref{fig:mod}, where we present a cutout of the AFI framework showing a portion of the 12-MR channel.
-There is a 14-MR that traverses seven T-sites in each of the 12-MRs, through the combination of Path 1 (blue) and Path 2 (purple).
-With the classical definition of a shortcut, this cycle is considered a ring.
-However, Path 3 (in green) connecting T1 and T2 produces a 12-MR when combined with Path 1, making Path 3 a modified shortcut.
-This 14-MR would not be counted under our new definition.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/modified-shortcut.pdf}
-\caption{Cutout of the AFI framework showing two stacked 12-MRs from the channel.
-A 14-MR is shown as T-sites replaced with aluminum atoms in gray.
-The two paths making connecting Al 1 and Al2 that make this 14-MR are highlighted with blue and purple bonds.
-Path 3 highlighted with green bonds is a modified shortcut connecting Al1 and Al2.
-\label{fig:mod}}
-\end{figure}
-\end{figure}
-
-To remove rings containing modified shortcuts from the full set of rings, we iterate over every T-site pair of the ring and check for the shortest path connecting them.
-If that shortest path is shorter than either of the two paths along the ring connecting the two T-sites, we check if the combination of this shorter path and the shorter of the two ring paths forms a new smaller ring.
-If so the iteration is broken, and the ring is removed from the counted set.
-
-\subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org8f45030}
-
-To add information about the spatial orientation of the rings around a T-site to the vertex symbol, we have developed a method to order the edges in the vertex symbol.
-We systematically list the rings by following the edges of the tetrahedron such that each ring listed is connected to one of the oxygens of the next ring listed.
-After removing all the rings that are not a part of the vertex symbol (\cref{section:vertex}) we use the following process to order them.
-
+\section{Highlights}
+\label{sec:org1e12ba5}
 \begin{enumerate}
-\item List all the possible arrangements of the oxygens bound to the T-site (4! = 24 possible arrangements).
-\item Use a predetermined order of edges: [[0,2],[0,1],[1,2],[2,3],[3,0],[1,3]].
-\begin{enumerate}
-\item Where each of those values represents the index of the oxygen to use.
+\item Classes of zeolite rings identified
+\item Algorithm and software to identify and enumerate rings developed
+\item Software used to analyze ring patterns across all zeolites
 \end{enumerate}
-\item Find the ring size (and multiplicity) connecting each pair of oxygens in this predetermined order.
-\item Make a list of weights, where for each pair of oxygens the weight is the ring size \texttimes{} multiplicity.
-\item Reverse sort the list of all possible oxygen arrangements by the correlating list of weights.
-\item Use the first oxygen arrangement coupled with the predetermined edge order to list the rings and multiplicity for each edge.
-\end{enumerate}
-
-\subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org79d764e}
-
-Finally, to determine all the ring sizes exhibited with in a zeolite framework, we take advantage of T-site symmetry.
-The rings of a framework are made of T-sites, and if two T-sites are symmetrically identical they will have the same set of rings passing through them.
-Therefore, we only need to find the rings associated with each symmetry distinct T-sites to know all the possible ring sizes within a framework.
-For example, AFI only contains one symmetry distinct T-site (T1).
-Using the basic definition of a shortcut, T1 is a part of 4-, 6-, 12-, and 14-MRs when using a cutoff of 18-MR.
-Every other T-site in the AFI framework is also a T1, thus the only possible ring sizes in AFI are 4-, 6-, 12-, and 14-MRs.
-
-\section{Results}
-\label{sec:orgcd96db0}
-\subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org15e507d}
-The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
-These rings listed by the IZA are referred to as tabulated rings in the literature \cite{curtis-statistical-2003}.
-In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the list of tabulated rings.
-These 'untabulated' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site.
-\cref{fig:fw-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database \cite{baerlocher-database-nodate}.
-There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/rings-vs-iza-rings.pdf}
-\caption{Counts of IZA frameworks containing each size ring between 3- and 18-MR using the Goetzke algorithm and the tabulated rings listed by the IZA \cite{baerlocher-database-nodate}.
-\label{fig:fw-counts}}
-\end{figure}
-\end{figure}
-
-Taking a closer look at some of these untabulated rings, highlights rings not typically listed for some frameworks, but still relevant to describing their topology.
-Using CHA as an example, \cref{fig:cha-rings} displays a 12-MR (in purple) that exists in CHA that circumferences the cage.
-This ring is not associated with the vertex symbol of the single symmetry distinct T-site in CHA and does not define a channel.
-Thus, this ring is not included in the list of tabulated rings.
-We would argue that this is still a ring that provides an important topological descriptor of CHA because none of the tabulated rings provides information about the size of the CHA cage.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/cha-all-rings.pdf}
-\caption{Chabazite cage and double 6-MR (D6R) with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
-The 8-MR in the D6R and the 12-MR are rings not typically discussed in literature.
-Si atoms have been replaced with Al atoms to help identify those rings in the overal cage strcuture.
-\label{fig:cha-rings}}
-\end{figure}
-\end{figure}
-
-Using AFI as another example, we find another type of ring that arises from traversing a pair of stacked rings and is not included in the list of tabulated rings.
-AFI, like CHA, contains one symmetry distinct T-site.
-According to the IZA, the AFI framework contains 4-, 6-, and 12-MRs \cite{baerlocher-database-nodate}.
-When we search for rings using the Goetzke algorithm \cite{goetzke-properties-1991}, we also find that it contains 14-MRs created by using seven T-sites from two 12-MRs that are separated by a distance of one oxygen (\cref{fig:afi-14}).
-Rings of this nature are prevalent in many frameworks, another example can be seen in the bottom right of \cref{fig:cha-rings}, where an 8-MR is highlighted traversing the two 6-MRs of the D6R.
-These types of rings may not be of interest depending on which topological feature one intends to describe.
-This has led us to create a modified definition of a shortcut as explained in METHODS REFERENCE, which excludes these types of rings.
-The benefit of this new shortcut definition is that larger rings that are missed by the vertex symbol or shortest path rings (i.e., 12-MR in AFI) are still captured, while excluding rings that arise from convolution of staked rings found with the typical shortcut definition.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/afi-14.pdf}
-\caption{Cutout of the 12-MR channel in AFI with a 14-MR (yellow) traversing seven T-sites of each 12-MR.
-The T-sites comprising the 14-MR have been replaced with Al for visibility.
-\label{fig:afi-14}}
-\end{figure}
-\end{figure}
-
-\subsection{Characterizing Frameworks by Rings}
-\label{sec:orgb71506d}
-With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
-\cref{fig:ring-counts} shows how many frameworks contain each size ring found using the various ring counting conventions from 3- to 18-MRs.
-This plot highlights the differences in the conventions and shows that a topological description of a framework based on rings will depend on the way that you define a ring.
-In general, a hierarchy of ring sizes found by each convention is: all rings not containing a shortcut \textgreater this work \textgreater shortest path rings \textgreater vertex symbol rings.
-While the IZA listed rings includes the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ring-counts-2.pdf}
-\caption{Number of IZA zeolite frameworks containing each size ring, using the various ring counting conventions, as well as the rings listed by the IZA Database \cite{baerlocher-database-nodate}.
-\label{fig:ring-counts}}
-\end{figure}
-\end{figure}
-
-One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of cycles that don't fit this definition, but still exhibit properties like rings.
-This is a trade-off between well-defined connectivity rules, and the inclusion of particular void environments that may still have important applications.
-These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings.
-One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}.
-Similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}.
-\cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/MFI-6MC.pdf}
-\caption{Cutout of MFI framework showing the structure referred to as an \(\alpha\)-6-MR in blue, and the two 5-MRs that compose it in green.
-The 6-membered cycle would not be found by any of the connectivity ring rules outlined in this work.
-\label{fig:mfi-6}}
-\end{figure}
-\end{figure}
-
-\subsection{Characterizing T-sites by Rings}
-\label{sec:orge809831}
-
-Considering that zeolite frameworks are comprised of one or more symmetry distinct T-sites, it may be of interest to describe those T-sites by the rings that pass through them.
-Most often the vertex symbol is used to make such a classification \cite{okeeffe-vertex-1997}.
-Sastre and Corma also provided characterization of T-sites using the shortest path rings that pass through them \cite{sastre-topological-2009}.
-In their work, they presented the ring index, which lists all the rings passing through a T-site from smallest to largest, and a subscript for each size representing it's multiplicity.
-The rings associated with a T-site can provided information about the local void environments around the T-site, and could potentially be correlated to other physicochemical properties of the T-site once enough once sufficient data on those physicochemical properties exists.
-
-Take for example the AFI framework, containing one symmetry distinct T-site.
-AFI contains 4-, 6-, 12-, and 14-MRs.
-To describe that T-site we can count how many of each of those rings pass through the T-site.
-We can also prune this list using our modified definition of a shortcut, the shortest path rings definition \cite{sastre-topological-2009}, or the rings contained within the vertex symbol of this T-site \cite{okeeffe-vertex-1997}.
-Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}:
-\begin{itemize}
-\item Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\)
-\item This work: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
-\item Shortest Path Rings: 4\(\cdot\)6\(_{\text{13}}\)
-\item Vertex Symbol Rings: 4\(\cdot\)6\(_{\text{11}}\)
-\end{itemize}
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/afi-funnel.pdf}
-\caption{Diagram shwoing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
-\end{figure}
-\end{figure}
-
-With an understanding of how we characterize T-site by counting the rings that pass through, \cref{tab:uninodal} shows the ring index for a selection of T-sites from uninodal (containing only one symmetry distinct T-site) frameworks.
-This table highlights the differences in rings counts found with each convention, and shows that in general as you move from left to right across the table the largest ring found decreases.
-The results in the shortest path column were found using ZSE, but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}.
-The results in the vertex symbol rings column were also found with zse, and agree directly with the vertex symbols listed on the IZA Database website \cite{baerlocher-database-nodate}.
-
-\begin{table}
-\centering
-\begin{threeparttable}
-\caption{Comparison of ring indices for the T-sites in various uninodal zeolite frameworks. \label{tab:uninodal}}
-{\scriptsize
-\begin{tabular}{lllll}
-\hline
-Framework & Rings & This Work & Shortest Path Rings \cite{sastre-topological-2009} & Vertex Symbol Rings \cite{baerlocher-database-nodate}\\
-\hline
-ABW & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{2}}\)\\
-ACO & 4\(_{\text{3}} \cdot\)6\(_{\text{3}} \cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{15}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{6}}\)\\
-AFI & 4\(_{\text{1}} \cdot\)\(_{\text{13}} \cdot\)12\(_{\text{1}} \cdot\)14\(_{\text{7}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{13}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{11}}\)\\
-ANA & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{16}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{8}}\)\\
-ATO & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)8\(_{\text{8}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}} \cdot\)12\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{9}}\)\\
-BCT & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{20}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{12}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
-CHA & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
-DFT & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}} \cdot\)10\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{6}} \cdot\)8\(_{\text{10}}\) & 4\(_{\text{2}} \cdot\)6\(_{\text{4}} \cdot\)8\(_{\text{6}}\)\\
-GIS & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
-GME & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{6}} \cdot\)12\(_{\text{7}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}} \cdot\)12\(_{\text{1}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{1}} \cdot\)8\(_{\text{2}}\)\\
-MER & 4\(_{\text{3}} \cdot\)8\(_{\text{4}} \cdot\)10\(_{\text{10}} \cdot\)14\(_{\text{14}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)8\(_{\text{4}}\)\\
-MON & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{5}} \cdot\)8\(_{\text{6}}\) & 4\(_{\text{1}} \cdot\)5\(_{\text{4}} \cdot\)8\(_{\text{4}}\)\\
-NPO & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}} \cdot\)12\(_{\text{40}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\) & 3\(_{\text{1}} \cdot\)6\(_{\text{6}}\)\\
-\hline
-\end{tabular}
-\begin{tablenotes}
-Vertex symbols have been represented in ring index format for ease of comparison.
-}
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-Next, we take an in-depth look at the ring counts for a framework with multiple symmetry distinct T-sites, to show how a ring index can provide information about the local environment around a T-site, and help differentiate them.
-MOZ is a zeolite framework containing 4-, 6-, 8-, 10-, 12-, 14-, and 18-MRs, 6 symmetry distinct T-sites, and two distinct 12-MR channels.
-\cref{tab:moz} shows the ring index for each T-site using each ring finding method.
-
-\begin{table}
-\centering
-\begin{threeparttable}
-\caption{Ring indices for each distinct T-site in the MOZ framework using each ring counting convention. \label{tab:moz}}
-{\scriptsize
-\begin{tabular}{lllll}
-\hline
-T-Site & Rings & This Work & Shortest Path Rings & Vertex Symbol Rings\\
-\hline
-T1 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)18\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T2 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{7}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{3}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T3 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(_{\text{5}} \cdot\)10\(_{\text{4}} \cdot\)12\(_{\text{4}} \cdot\)14\(_{\text{5}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\(\cdot\)12\(_{\text{4}}\) & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8 & 4\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)8\\
-T4 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)10\(_{\text{6}} \cdot\)12\(\cdot\)18\(_{\text{26}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{6}} \cdot\)12\\
-T5 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}} \cdot\)10\(_{\text{6}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{7}}\)\\
-T6 & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)10\(_{\text{2}} \cdot\)12\(_{\text{8}} \cdot\)14\(_{\text{18}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}} \cdot\)12\(_{\text{8}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\) & 4\(_{\text{2}} \cdot\)6\(\cdot\)8\(_{\text{3}}\)\\
-\hline
-\end{tabular}
-\begin{tablenotes}
-Vertex symbols have been represented in ring index format for ease of comparison.
-}
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\cref{fig:moz} shows the T-site locations inside a 2-dimensional view of the framework.
-If you were interested in which T-sites have access to the 12-MR channels, the shortest path rings and vertex symbol rings would only suggest T3 participates in the 12-MR rings.
-However, all rings and this work both identify T4 and T6 as participating in the 12-MR channels as highlighted in \cref{fig:moz}.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/moz.pdf}
-\caption{Cutout of the MOZ framework showing two 12-MR channels, with an example of each distinct T-site highlighted.
-T1: navy, T2: green, T3: orange, T4: purple, T5: blue, and T6: red.
-As shown, T3, T4, and T6 are all associted with the 12-MR channels, while T1, T2, and T5 are not connected to the 12-MR channels.
-\label{fig:moz}}
-\end{figure}
-\end{figure}
-
-We next used ZSE to find the rings associated with every symmetry distinct T-site in every framework across the IZA Database using each of the four ring counting conventions.
-For each T-site we used the rings to generate a ring index, and \cref{fig:unique-ts} shows how many unique ring indices are present when using each of the ring counting conventions.
-The plot follows intuition with the number of unique ring indices decreasing as we use more restrictive ring counting conventions, because less rings are found and provides less room for differentiation.
-This raises the question, if you want to ascertain chemical or physical properties about a T-site based on its ring count, and differentiate these T-sites from other similar but distinct T-site, which ring counting convention will suffice?
-The answer will depend on what level of detail is desired.
-Larger rings can be found with the standard shortcut definition, rings traversing other stacked rings will be excluded with our modified shortcut definition.
-The shortest path definition and vertex symbol rings will provide the most localized information about a T-site.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/unique-ts.pdf}
-\caption{Number of unique T-site ring indices when classified by the rings passing through them using the various ring counting conventions.
-There are 1460 T-sites across all the frameworks in the IZA Database.
-As we move from less restrictive to more restrictive (left to right) ring counting conventions, the number of unique ring indices decreases.
-\label{fig:unique-ts}}
-\end{figure}
-\end{figure}
-
-To further compare the ring counting conventions, we show a distribution of the number of T-sites containing each size ring between 3- and 18-MR in \cref{fig:tsite-frequency} (right).
-This plot highlights that more T-sites contain larger sized rings when using the basic definition of a shortcut, and at smaller rings sizes (\textless6-MR) all the ring counting conventions return the same results.
-To further emphasize this results, we have provided a cumulative distribution of the same data normalized to the maximum 'rings' value in \cref{fig:tsite-frequency}.
-At 6-MRs is where we see the cumulative distribution functions deviate from each other.
-The largest deviation takes place at 12-MRs, and the cumulative distributions start to level out at larger ring sizes.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/dist-cumudist.pdf}
-\caption{Frequency of T-sites accross all IZA frameworks containing ring sizes between 3- and 18-MR (left), and cumulative distribution of T-sites containing each ring size normalized to the final 'rings' value (right).
-\label{fig:tsite-frequency}}
-\end{figure}
-\end{figure}
-
-To complete the comparison of ring counting conventions, we have developed method to determine how similar the ring counts of one convention are to another.
-We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}.
-In this equation, sr is the number of similar rings that are found in both counting conventions, and mr is the maximum number of rings found by either convention.
-For example: the ring index of AFI using the classic shortcut definition and the shortest path definition are 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\) and 4\(\cdot\)6\(_{\text{13}} \cdot\).
-The number of similar rings found by both conventions is 14, and the maximum number of rings found by either convention is 22.
-This would lead to a similarity score of 0.636.
-We do this for every T-site between two conventions and average the similarity score to get the results in \cref{fig:similarity}.
-Down the diagonal each method is compared to itself and clearly has a similarity of 1.
-The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compare to the least restrictive convention (rings) has the lowest similarity score.
-The two most similar ring counting methods are our modified shortcut definition and the shortest path rings.
-
-\begin{equation}\label{eq:similarity}
-s = \mathrm{ \frac{sr}{mr} }
-\end{equation}
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/similarity-heat-map.pdf}
-\caption{Heat map showing the similarity score for four ring counting methods.
-Similarity score of 1 means identical set of rings returned, while a similarity of 0 would mean no matching rings are returned.
-\label{fig:similarity}}
-\end{figure}
-\end{figure}
-
-One final point we would like to make about T-site characterization, is that the vertex symbol and ring indices only provide information about rings, and do not give any information about the spatial orientation of those rings around the T-site.
-The vertex symbol comes close to accomplishing that, by finding rings associated with opposite edge pairs of the tetrahedron, and then listing the those pairs from smallest to largest.
-However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments.
-For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\).
-However, the orientation of those rings around each of those T-sites are not identical.
-\cref{fig:stereo} shows a cutout of each of these frameworks that only includes the atoms that make up the rings of the vertex symbol around the specified T-sites.
-We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1.
-The main difference is the location of the 5\(_{\text{2}}\)- and 4-MR edges.
-They are highlighted in \cref{fig:stereo}.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/stereo.pdf}
-\caption{Cutout of the MOR, EON, and MON frameworks that only shows the rings associated with the vertex symbol of T3, T9, and T1 respectively.
-The 4-MR (green) and 2\texttimes{} 5-MR (teal) that are in swapped positions are highlighted for emphasis.
-The 4-MR for MON, and the 2\timex 5-MRs for MOR and EON are into the plane, and not easily shown.
-\label{fig:stereo}}
-\end{figure}
-\end{figure}
-
-The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
-This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings described in \cref{section:ov}.
-
-With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
-The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol.
-Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about such as deprotonation energy, T-site substitution energy, or catalytic properties of reactions happening at that T-site.
-
-We have used this new ordered vertex symbol to characterize all 1460 T-sites in the IZA Database.
-We found that using the standard vertex symbol to characterize T-sites there are 649 unique vertex symbols present.
-In contrast, there are 666 unique ordered vertex symbols across every T-site.
-This would imply that not a large amount of vertex symbols contain stereochemical differences.
-In \cref{tab:ov} we provide a list of some common T-site vertex symbols, and their representative ordered vertex symbols.
-
-\begin{table}
-\centering
-\begin{threeparttable}
-\caption{List of three vertex symbols, the T-sites associated with them, and the representative ordered vertex symbol for those T-sites. \label{tab:ov}}
-{\scriptsize
-\begin{tabular}{lll}
-\hline
-Vertex Symbol & Framework T-site & Ordered Vertex Symbol\\
-\hline
-5\(\cdot\)5\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)10 & EWS T3 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & ITN T9 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & ITN T21 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & OKO T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & OKO T5 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & PCS T2 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & PCS T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFS T10 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFV T3 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & SFV T7 & 5\(_{\text{2}} \cdot\)10\(\cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
- & TUN T10 & 10\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)5\(\cdot\)5\(\cdot\)5\\
-\hline
-4\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)8\(\cdot\)5\(\cdot\)8 & DAC T3 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & DAC T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & EON T10 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & EPI T1 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & MOR T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
- & RSN T4 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & VNI T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & VSV T2 & 5\(_{\text{2}} \cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)8\(\cdot\)4\\
- & YFI T9 & 5\(_{\text{2}} \cdot\)8\(\cdot\)8\(\cdot\)5\(\cdot\)5\(\cdot\)4\\
-\hline
-4\(\cdot\)6\(\cdot\)4\(\cdot\)6\(\cdot\)6\(\cdot\)8 & ATN T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
- & JSN T3 & 8\(\cdot\)6\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\\
- & PON T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
- & SAS T1 & 8\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\(\cdot\)6\\
- & ZON T3 & 8\(\cdot\)6\(\cdot\)6\(\cdot\)4\(\cdot\)4\(\cdot\)6\\
-\hline
-\end{tabular}
-\begin{tablenotes}
-}
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-\section{Conclusions}
-\label{sec:org0be7bec}
-Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
-We find that the commonly reported ring sizes in literature and on the IZA website leave out many rings that fit the classical definition of a cycle that does not contain a shortcut.
-To completely describe the topology of a zeolite these rings are required, however there are often cases where someone might want to consider only a subset of rings of interest.
-
-We have shown a comparison of three different conventions used to count rings, and highlighted the differences in rings that are found by each convention.
-The classic definition of a ring identifies the largest set of ring sizes across all the zeolite frameworks, while the shortest path ring and vertex symbol rings only identify smaller ring sizes.
-We have provided a modified definition of a shortcut that when used to find rings still finds larger rings defining channel openings, but excludes rings that are able to be decomposed into at least one smaller ring.
-It is important to understand the difference of ring sizes and types found by each convention when discussing the rings of a zeolite framework.
-A disadvantage to using purely connectivity based definitions of rings is the exclusion of cycles in a framework that behave physicochemically like a ring but contain a shortcut.
-We have displayed an example case of these geometric rings, and in the future it would be beneficial to develop a computation method of identifying these cycles.
-
-This same methodology was used to describe T-sites in zeolite frameworks by counting all of the rings that pass through the T-site using each of the ring counting conventions described.
-When using ring finding conventions that find larger rings, we see more diversity in the descriptions of T-sites, which can aid researchers who want to identify similar T-sites across multiple frameworks.
-We have also shown that the vertex symbol used to describe T-sites leaves out subtle but distinct stereochemical differences in the spatial orientation of the rings around a T-site.
-To address this shortcoming we have provided a new method for ordering the rings of a vertex symbol that takes into consideration the ring stereochemistry and is able to identify differences in T-sites that have the same vertex symbol.
-In the future, correlating physicochemical properties of T-sites to the ring descriptors identified with each ring counting convention can help identify sets of frameworks with desired T-site properties.
-
-\section{Acknowledgments}
-\label{sec:org805a78b}
-
-We acknowledge financial support provided by the National Science Foundation under Cooperative Agreement No.
-EEC-1647722, which is an Engineering Research Center for the Innovative and Strategic Transformation of Alkane Resources.
-JTC thanks the Arthur J.
-Schmitt Foundation for financial aid in the form of a PhD fellowship.
-We thank Dr.
-Christian Baerlocher for the numerous discussions about the topology of zeolites, and methodologies used by the IZA Structure Database.
-We thank Dr.
-German Sastre for providing a copy of zeoTsites to compare results, and for the helpful conversions about use cases and zeolite topology.
-This research was supported in part by the Notre Dame center for Research Computing through access to high performance computing clusters.
-\backmatter
-
-\bibliographystyle{elsarticle-num}
-\bibliography{ref}
 \end{document}

### 2022-12-05 elsman/manuscript.tex
fixed some of the reviewers comments

--- 2022-08-29 elsman/manuscript.tex
@@
-% Created 2022-08-29 Mon 07:25
+% Created 2022-12-04 Sun 20:29
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org2a67457}
+\label{sec:org6bb8471}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 Here we present an analysis of rings captured by Goetzke and Klein's efficient ring finding algorithm \cite{goetzke-properties-1991}, and compare those rings to the rings found by other previously published ring set reduction conventions.
-We have implemented all of these ring finding conventions in a Python package called the Zeolite Simulation Environment (ZSE, \url{https://www.github.com/jtcrum/zse}).
+We have implemented all of these ring finding conventions in a Python package called the Zeolite Simulation Environment (ZSE) \cite{crum-jtcrumzse-2022}.
 We use ZSE to provide an analysis of rings captured by each convention for the entire set of zeolite frameworks contained on the IZA Database \cite{baerlocher-database-nodate} to compare how these sets of rings provide different characterizations of said frameworks.
@@
 \section{Methods}
-\label{sec:org6431034}
+\label{sec:org738721e}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org42ddcba}
+\label{sec:orgdf60ef5}
 
-In this work we implement an efficient algorithm that was presented by Goetzke and Klein to find all the rings associated with a T-site that do not contain a shortcut \cite{goetzke-properties-1991} in a Python package called the Zeolite Simulation Environment (ZSE).
+In this work we implement the efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} in a Python package called the Zeolite Simulation Environment (ZSE) \cite{crum-jtcrumzse-2022} to find all the rings associated with a T-site that do not contain a shortcut.
 In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures.
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org20e1097}
+\label{sec:org73f1098}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org42016a9}
+\label{sec:org284d066}
 
@@
 
-\subsection{Modified Shortcut Rings}
-\label{sec:orgd5be89b}
+\subsection{Modified Shortcut Rings \label{section:modified}}
+\label{sec:org43cb558}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org8f45030}
+\label{sec:org44354c5}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org79d764e}
+\label{sec:org1887ac4}
 
@@
 \section{Results}
-\label{sec:orgcd96db0}
+\label{sec:org83f1d83}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org15e507d}
+\label{sec:org4bd7b88}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 These types of rings may not be of interest depending on which topological feature one intends to describe.
-This has led us to create a modified definition of a shortcut as explained in METHODS REFERENCE, which excludes these types of rings.
+This has led us to create a modified definition of a shortcut as explained in \cref{section:modified}, which excludes these types of rings.
 The benefit of this new shortcut definition is that larger rings that are missed by the vertex symbol or shortest path rings (i.e., 12-MR in AFI) are still captured, while excluding rings that arise from convolution of staked rings found with the typical shortcut definition.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:orgb71506d}
+\label{sec:org9a58f2f}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 
-One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of cycles that don't fit this definition, but still exhibit properties like rings.
+One drawback to using a ring convention based on connectivity and shortcuts is the exclusion of non-ring cycles that exhibit geometric properties similar to those designated as rings.
 This is a trade-off between well-defined connectivity rules, and the inclusion of particular void environments that may still have important applications.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:orge809831}
+\label{sec:org777c54c}
 
@@
 \section{Conclusions}
-\label{sec:org0be7bec}
+\label{sec:org32d48ad}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org805a78b}
+\label{sec:org89b8250}
 

### 2022-12-05 elsman/fix_ref.py

@@ -1,0 +1,13 @@
+file = open('references.bib',"r")
+lines = file.readlines()
+file.close()
+for i,line in enumerate(lines):
+    if '@' in line:
+        if '_' in line:
+            line = line.replace('_','-')
+            print(line)
+        lines[i]=line
+
+file = open('ref.bib',"w")
+file.writelines(lines)
+file.close()

### 2022-12-11 elsman/manuscript.tex
addressed most of the reviewer comments

@@
-% Created 2022-12-04 Sun 20:29
+% Created 2022-12-10 Sat 19:04
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org6bb8471}
+\label{sec:orgd4436af}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 
-Many researchers have presented methods to define and count the rings present in zeolites \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001}.
+Many researchers have presented methods to define and count the rings present in zeolites \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001}, and various software tools have been created for this purpose, such as ZeoTSites \cite{sastre-zeotsites-2001} and ToposPro \cite{blatov-applied-2014}.
 Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges.
@@
 Many definitions build off this idea that rings do not contain shortcuts.
+A more restrictive definition for a ring is that of a "strong ring", which is not the sum of a set of smaller rings \cite{goetzke-properties-1991,blatov-three-periodic-2007}.
 In a zeolite, rings are used to define the topology of the framework, such as the channels and void environments that are present.
@@
 These grouped pairs are listed from smallest to largest forming the vertex symbol.
+Since the vertex symbol only lists rings by opposite edges and size, it is possible for one vertex symbol to describe two T-sites who's rings have differing connectivity, and are effectively constitutional isomers.
+To accurately describe the stereochemistry of the rings around a T-site, we propose a new method for ordering the rings of a vertex symbol in this work.
 \cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework.
@@
 
-Another ring counting convention presented by Sastre and Corma is to count only the shortest path connecting an O-T-O in a framework.
+Another ring counting convention presented by Sastre and Corma is to count only the shortest path connecting any O-T-O in a framework.
+This convention provides an univocal count of the rings that pass through a T-site without the ambiguity of needing to limit the search to a certain ring size.
 With this definition, they can find and count all the rings in a framework that are the shortest path for at least one set of O-T-O along the cycle \cite{sastre-topological-2009}.
-Using AFI as an example again (\cref{fig:3.3}), we see that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom.
+Using AFI as an example again (\cref{fig:3.3}(a)), we see that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom.
 The 12-MR would not be included because for every O-T-O along the 12-MR, the shortest path connecting them is not the 12-MR.
@@
 
+When used to count rings in zeolites, the shortest path convention generally finds less (and smaller) rings than the convention counting all non-shortcut bearing rings.
+When compared to the IZA Structure Database, the all rings convention generally includes rings larger than what is listed for a given framework, and the shortest path convention generally doesn't include the larger rings listed for a given framework.
+In this work we provide an alternative definition for shortcuts that only require the shortcut to be shorter than one of the paths connecting two nodes along the ring, not both paths.
+When this modified shortcut is used for ring counting a set of rings is returned which is closer to what one would find listed in the IZA Structure Database as rings of interest, with counts falling between the sets of all rings and shortest path rings.
+This modified shortcut is explained graphically in \cref{fig:3.3}(b), where we present a cutout of the AFI framework showing a portion of the 12-MR channel.
+There is a 14-MR that traverses seven T-sites in each of the 12-MRs, through the combination of Path 1 (blue) and Path 2 (purple).
+With the classical definition of a shortcut, this cycle is considered a ring.
+However, Path 3 (in green) connecting T1 and T2 produces a 12-MR when combined with Path 1, making Path 3 a modified shortcut.
+This 14-MR would not be counted under our new definition.
+
 \begin{figure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/afi-12-6.pdf}
-\caption{Cutout of the 12-MR channel in AFI highlighting a 12-MR in purple, and a 6-MR in blue.
+\includegraphics[width=\textwidth]{figures/chapter-3/afi-14-and-modified.pdf}
+\caption{Cutouts of the 12-MR channel in AFI: a) Highlighting a 12-MR in purple, and a 6-MR in blue.
 The 6-MR is included in the vertex symbol of labeled T1 because it is the shortest path connecting O1 and O2.
 The 12-MR would not be included in the vertex symbol or shortest path ring list because for each O-T-O along the 12-MR there is a shorter path connecting them.
+b) A 14-MR is shown as T-sites replaced with aluminum atoms in gray.
+The two paths connecting Al1 and Al2 that make this 14-MR are highlighted with blue and purple bonds.
+Path 3 highlighted with green bonds is a modified shortcut connecting Al1 and Al2.
 \label{fig:3.3}}
@@
 We can use ring counts to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites.
+These rings can be used to describe the channel sizes for understanding shape selectivity in catalysis, zeolite building blocks for solid state chemists, and the sizes of framework cages and windows to provide insights into adsorption properties.
 Since various conventions exist that can reduce the set of rings in a zeolite to more strictly defined properties, the ring counts returned by the various conventions will differ.
@@
 \section{Methods}
-\label{sec:org738721e}
+\label{sec:orgd3de46b}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:orgdf60ef5}
+\label{sec:org73ee410}
 
-In this work we implement the efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} in a Python package called the Zeolite Simulation Environment (ZSE) \cite{crum-jtcrumzse-2022} to find all the rings associated with a T-site that do not contain a shortcut.
+In this work we implement the efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} in ZSE \cite{crum-jtcrumzse-2022} to find all the rings associated with a T-site that do not contain a shortcut.
 In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures.
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org73f1098}
+\label{sec:org5c1b103}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:org284d066}
+\label{sec:orgc6e6cb9}
 
@@
 \subsection{Modified Shortcut Rings \label{section:modified}}
-\label{sec:org43cb558}
+\label{sec:orgb03fdff}
 
@@
 Our shortcut only needs to be shorter than one of the paths connecting those two nodes along the cycle.
-
-This definition is explained graphically in \cref{fig:mod}, where we present a cutout of the AFI framework showing a portion of the 12-MR channel.
-There is a 14-MR that traverses seven T-sites in each of the 12-MRs, through the combination of Path 1 (blue) and Path 2 (purple).
-With the classical definition of a shortcut, this cycle is considered a ring.
-However, Path 3 (in green) connecting T1 and T2 produces a 12-MR when combined with Path 1, making Path 3 a modified shortcut.
-This 14-MR would not be counted under our new definition.
-
-\begin{figure}
-\begin{figure}[H]
-\centering
-\includegraphics[width=\textwidth]{figures/chapter-3/modified-shortcut.pdf}
-\caption{Cutout of the AFI framework showing two stacked 12-MRs from the channel.
-A 14-MR is shown as T-sites replaced with aluminum atoms in gray.
-The two paths making connecting Al 1 and Al2 that make this 14-MR are highlighted with blue and purple bonds.
-Path 3 highlighted with green bonds is a modified shortcut connecting Al1 and Al2.
-\label{fig:mod}}
-\end{figure}
-\end{figure}
+This definition has been described graphically in \cref{fig:3.3}(b).
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:org44354c5}
+\label{sec:orgec814e5}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:org1887ac4}
+\label{sec:orgd18c4e3}
 
@@
 \section{Results}
-\label{sec:org83f1d83}
+\label{sec:org9a859e6}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org4bd7b88}
+\label{sec:org198e873}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:org9a58f2f}
+\label{sec:org54cc411}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:org777c54c}
+\label{sec:org98f8218}
 
@@
 The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
+While a vertex symbol includes the shortest rings connecting the four oxygen atoms adjacent to a T-site, and listing them by size and opposite edges, there is still ambiguity in the exact connectivity of those rings.
 This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings described in \cref{section:ov}.
@@
 \section{Conclusions}
-\label{sec:org32d48ad}
+\label{sec:orgb5c3952}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org89b8250}
+\label{sec:org577f172}
 

### 2022-12-11 elsman/manuscript.tex
addressed most reviewer comments

@@
-% Created 2022-12-10 Sat 19:04
+% Created 2022-12-10 Sat 19:19
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgd4436af}
+\label{sec:org497996e}
 Zeolites are three dimensional crystalline structures containing tetrahedral Si or Al atoms connected by oxygen bridges.
@@
 \section{Methods}
-\label{sec:orgd3de46b}
+\label{sec:org568c810}
 
 \subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
-\label{sec:org73ee410}
+\label{sec:orgeb1d79d}
 
@@
 \subsection{Finding Vertex Symbol Rings \label{section:vertex}}
-\label{sec:org5c1b103}
+\label{sec:org415adcd}
 
@@
 \subsection{Finding Shortest Path Rings}
-\label{sec:orgc6e6cb9}
+\label{sec:org6be3a64}
 
@@
 \subsection{Modified Shortcut Rings \label{section:modified}}
-\label{sec:orgb03fdff}
+\label{sec:org9232478}
 
@@
 \subsection{Ordered Vertex Symbols \label{section:ov}}
-\label{sec:orgec814e5}
+\label{sec:org7db42da}
 
@@
 \subsection{Determining All Ring Sizes Contained in a Zeolite Framework}
-\label{sec:orgd18c4e3}
+\label{sec:org6a29aa7}
 
@@
 \section{Results}
-\label{sec:org9a859e6}
+\label{sec:org47268f3}
 \subsection{Characterizing Rings in a Zeolite Graph}
-\label{sec:org198e873}
+\label{sec:orgdbc479c}
 The IZA Database \cite{baerlocher-database-nodate} is a common reference used to identify all the rings in a zeolite framework, however it only lists the rings that define a channel (ex: 12-MR in AFI), or rings associated with the symbol of a T-site.
@@
 \subsection{Characterizing Frameworks by Rings}
-\label{sec:org54cc411}
+\label{sec:org3c3c2d3}
 With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
@@
 \subsection{Characterizing T-sites by Rings}
-\label{sec:org98f8218}
+\label{sec:org6b7283b}
 
@@
 \section{Conclusions}
-\label{sec:orgb5c3952}
+\label{sec:org9840720}
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
@@
 \section{Acknowledgments}
-\label{sec:org577f172}
+\label{sec:org5946b2f}
 

### 2022-12-27 elsman/manuscript.tex
WFS edits

@@
 Rings have been used to identify feasible zeolites topologies \cite{li-why-2014}, to describe the similarity and differences between zeolites \cite{curtis-statistical-2003,bermudez-calculation-2017}, to identify sites or voids of catalytic relevance \cite{li-first-principles-2018,kester-effects-2021}, and as machine learning finger prints of physicochemical properties \cite{helfrecht-new-2019}.
+Rings can be used to describe the channel sizes for understanding shape selectivity in catalysis, zeolite building blocks for solid state chemists, and the sizes of framework cages and windows to provide insights into adsorption properties.
+The dimension and number of rings in a framework can be used to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites.
+The number and dimension  of rings that pass through a T-site can be used to describe and differentiate the local void environment around symmetry distinct T-sites and to  differentiate symmetry distinct oxygen atoms.
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ring-examples2.pdf}
+\includegraphics[width=0.5\textwidth]{figures/chapter-3/ring-examples2.pdf}
 \caption{Cutout of the Chabazite framework showing a path (5-6-7-8-9) highlighted with purple bonds, a cycle (3-4-18-19-20-17) highlighted with blue bonds, an 8-MR filled in with pink, and a 4-MR filled in with green.
@@
 
-Many researchers have presented methods to define and count the rings present in zeolites \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001}, and various software tools have been created for this purpose, such as ZeoTSites \cite{sastre-zeotsites-2001} and ToposPro \cite{blatov-applied-2014}.
-Zeolites can naturally be represented using graph theory, where atoms are nodes, and bonds are edges.
-In some cases researchers will use T-sites as nodes and oxygen atoms as edges, since the oxygen atoms are only connected to two T-sites \cite{goetzke-properties-1991}.
-Some basic definitions of graphs that will be used going forward are highlighted in a cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}, and described in \cref{tab:graphs}.
+A number of definitions and enumerations of rings in zeolites have been presented   \cite{guttman-ring-1990,goetzke-properties-1991,franzblau-computation-1991,yuan-efficient-2002-1,wooten-structure-2002,le-roux-ring-2010,sastre-zeotsites-2001} and implemented in software tools, including ZeoTSites \cite{sastre-zeotsites-2001} and ToposPro \cite{blatov-applied-2014}.
+Zeolites are naturally  represented using graph theory, in which  atoms are nodes and bonds are edges.
+Because the oxygen atoms are only connected to exactly two T-sites, their inclusion in a graph does not change the graph structure, and thus T-sites can be  defined as nodes and the intervening oxygen atoms as edges \cite{goetzke-properties-1991}.
+The cutout of the chabazite (CHA) framework shown in \cref{fig:3.1}  and  \cref{tab:graphs} summarize basic definitions of graphs that will be used here.
 The size of each of these features can either be defined by the total number of atoms contained in them, or by the number of T-sites contained in them.
-The latter method is the standard convention which we will use.
-A path is a sequence of edges which joins a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}).
+The latter method is the standard convention, which we will use.
+A path is a sequence of edges that connects a sequence of nodes in which every edge is distinct (shown with purple highlighted bonds in \cref{fig:3.1}).
 A cycle is a path that starts and ends at the same node, and does not repeat any other nodes (shown by blue bonds in \cref{fig:3.1}).
-In the most basic form a ring is a cycle that does not contain any shortcuts \cite{guttman-ring-1990,goetzke-properties-1991}.
-\cref{fig:3.1} shows an 8-membered ring and a 4-MR filled in with pink and green respectively.
+A zeolite crystal contains an infinite number of cycles, but the symmetry of the crystal reduces this to a finite number of symmetry-distrinct rings whose exact identity is subject to definition.
+\cref{fig:3.1} shows an 8-membered ring (8-MR) and a 4-MR filled in pink and green, respectively.
 
@@
 {\tiny
-\begin{tabular}{lp{2cm}p{2cm}p{2cm}p{2cm}}
+\begin{tabular}{lp{1.5cm}p{2cm}p{2cm}p{2cm}}
 \hline
@@
 
-A shortcut is defined as a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}, or in other words, the path is a shortcut if it is shorter than both paths between these nodes along the cycle.
+% These rings are relevant to ion exchange and shape selectivity in catalysis, to describing the zeolite building blocks relevant to zeolite assembly, and to capturing void spaces relevant to  adsorption properties.
+
+ In the most basic definition, a ring is any cycle that does not contain any shortcuts \cite{guttman-ring-1990,goetzke-properties-1991}.
+A shortcut is defined as a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}; i.e., a path is a shortcut if it is shorter than both paths between these nodes along the cycle.
 An example is the path connecting T17 and T18 in \cref{fig:3.1}, which is a shortcut of the blue cycle.
 This path contains two T-sites, splitting the 8-MR cycle into two 4-MRs.
-Many definitions build off this idea that rings do not contain shortcuts.
-A more restrictive definition for a ring is that of a "strong ring", which is not the sum of a set of smaller rings \cite{goetzke-properties-1991,blatov-three-periodic-2007}.
-In a zeolite, rings are used to define the topology of the framework, such as the channels and void environments that are present.
-Counting the rings that pass through a T-site can also be used to describe and differentiate the local void environment around symmetry distinct T-sites.
-Like T-sites, ring counts can be used to differentiate symmetry distinct oxygen atoms.
+While general, this definition reports as distinct rings features of a zeolite framework that do not intuitively relate to local structure or accessibility or usefully ``fingerprint'' frameworks.
+Thus, more restrictive defintions of zeolite rings have been proposed.
+For instance, a ``strong ring'' is defined as a ring that is not the sum of a set of smaller rings.
+\red{A little unclear.
+How does this relate to shortcuts?} \cite{goetzke-properties-1991,blatov-three-periodic-2007}.
+
+ Another pruning method is based off of the concept of a vertex symbol.
+The vertex symbol is a common way to describe the structure around the T-sites of zeolite frameworks and were first used to do so in 1997 by O'Keeffe and Hyde \cite{okeeffe-vertex-1997}.
+The vertex symbol of a T-site is determined by the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms.
+The symbol is formed by grouping rings on opposite edges of the tetrahedron and listing from smallest to largest.
+\cref{fig:cha-vertex}(a) shows as example the tetrahedron formed at the single symmetry-distinct T-site in the CHA framework.
+The edges of the tetrahedron are labeled to aid in identification.
+\cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges.
+The vertex symbol of T1 in CHA is thus 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8.
+For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol.
+An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.
+The rings of a framework are then the union of the rings of all the distinct T-sites, a definition that eliminates some rings that satisfy the no shortcut criterion.
 
@@
 
-Vertex symbols are a common way to describe the T-sites of a zeolite framework, and were first used to do so in 1997 by O'Keeffe and Hyde in 1997 \cite{okeeffe-vertex-1997}.
-The vertex symbol of a T-site contains the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms.
-The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together.
-These grouped pairs are listed from smallest to largest forming the vertex symbol.
-Since the vertex symbol only lists rings by opposite edges and size, it is possible for one vertex symbol to describe two T-sites who's rings have differing connectivity, and are effectively constitutional isomers.
-To accurately describe the stereochemistry of the rings around a T-site, we propose a new method for ordering the rings of a vertex symbol in this work.
-\cref{fig:cha-vertex}(a) shows an example of the tetrahedron formed at the single symmetry distinct T-site in the CHA framework.
-The edges of the tetrahedron are labeled to aid in identification.
-\cref{fig:cha-vertex}(b-d) show the rings associated with each opposite pair of edges.
-The vertex symbol of T1 in CHA is thus able to be determined as 4\(\cdot\)4\(\cdot\)4\(\cdot\)8\(\cdot\)6\(\cdot\)8.
-For a T-site that contains a multiplicity of rings at one edge, that multiplicity would be represented as a subscript in the vertex symbol.
-An example would be the vertex symbol of 4\(\cdot\)6\(_{\text{2}} \cdot\)6\(\cdot\)6\(_{\text{3}} \cdot\)6\(_{\text{2}} \cdot\)6\(_{\text{3}}\) for T1 in AFI.
-
-Another ring counting convention presented by Sastre and Corma is to count only the shortest path connecting any O-T-O in a framework.
-This convention provides an univocal count of the rings that pass through a T-site without the ambiguity of needing to limit the search to a certain ring size.
-With this definition, they can find and count all the rings in a framework that are the shortest path for at least one set of O-T-O along the cycle \cite{sastre-topological-2009}.
-Using AFI as an example again (\cref{fig:3.3}(a)), we see that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom.
-The 12-MR would not be included because for every O-T-O along the 12-MR, the shortest path connecting them is not the 12-MR.
-The difference between this method and the vertex symbol rings is subtle, but with this shortest path convention any ring belonging to the vertex symbol of any T-site in a framework will be included in the ring count for each of the T-sites that ring passes through.
+Sastre and Corma \cite{sastre-topological-2009} proposed an alternative convetion, in which distinct rings are only those that form the shortest path connecting any \ce{O-T-O} in a framework \red{do you mean cycle?}.
+This convention provides an univocal count of the rings that pass through a T-site without the ambiguity of limiting the search to a certain ring size.
+Using AFI as an example,  (\cref{fig:3.3}(a)) shows that both a 12-MR (purple) and a 6-MR (blue) pass through the labeled T1 atom.
+The 12-MR would not be included in the Sastre and Corma enumeration because for every \ce{O-T-O} along the 12-MR, the shortest path connecting them is not the 12-MR.
+The difference between this criterion and the vertex symbol rings is subtle, but with this shortest path convention any ring belonging to the vertex symbol of any T-site in a framework will be included in the ring count for each of the T-sites that ring passes through.
 \cref{fig:3.4} shows a cutout of the TON framework including a 6- (blue) and 10-MR (orange).
-For T1, only the 6-MR would be counted in the vertex symbol because it is the shortest path connecting O2 and O3.
+For T1, only the 6-MR would be counted within the vertex symbol convention because it is the shortest path connecting O2 and O3.
 The 10-MR is part of the vertex symbol for T3 because it is the shortest path connecting O2 and O14.
-Since this 10-MR is the shortest path for at least one set of O-T-O along the ring, and passes through T1, it does get counted in the shortest path rings for T1.
-
-When used to count rings in zeolites, the shortest path convention generally finds less (and smaller) rings than the convention counting all non-shortcut bearing rings.
-When compared to the IZA Structure Database, the all rings convention generally includes rings larger than what is listed for a given framework, and the shortest path convention generally doesn't include the larger rings listed for a given framework.
-In this work we provide an alternative definition for shortcuts that only require the shortcut to be shorter than one of the paths connecting two nodes along the ring, not both paths.
-When this modified shortcut is used for ring counting a set of rings is returned which is closer to what one would find listed in the IZA Structure Database as rings of interest, with counts falling between the sets of all rings and shortest path rings.
-This modified shortcut is explained graphically in \cref{fig:3.3}(b), where we present a cutout of the AFI framework showing a portion of the 12-MR channel.
-There is a 14-MR that traverses seven T-sites in each of the 12-MRs, through the combination of Path 1 (blue) and Path 2 (purple).
-With the classical definition of a shortcut, this cycle is considered a ring.
-However, Path 3 (in green) connecting T1 and T2 produces a 12-MR when combined with Path 1, making Path 3 a modified shortcut.
-This 14-MR would not be counted under our new definition.
+Because this 10-MR is the shortest path for at least one set of \ce{O-T-O} along the ring, and passes through T1, it does get counted in the shortest path rings for T1.
+\red{The end of this paragraph is unclear as to whether the Vertex symbol or Sastre definition is being used.
+Please reread and clarify.}
 
@@
 
+% The vertex symbol is a property of a T-site that is determined by the rings that pass through that site.
+Because the vertex symbol convention  lists rings in an order determined by size and locations on opposite edges of the tetrahedron, it is possible for two T-sites of differing connectivity (effectively, local constitutional isomers) to be described by the same  vertex symbol
+% The ring sizes, and  their multiplicity, for opposite edges (edges of the tetrahedron that do not connect to the same oxygen atom) are grouped together.
+These grouped pairs are listed from smallest to largest forming the vertex symbol.
+%  the number and type of rings Because various conventions exist that can reduce the set of rings in a zeolite to more strictly defined properties, the ring counts returned by the various conventions will differ.
+Differences in ring counts leads to differences in how one might describe the topological environment of a zeolite.
+Therefore, when using rings to describe the properties of a framework, T-site, or oxygen atom, it is of value to characterize the differences between the conventions and to have a use one that determines the features of interest.
+
 \begin{figure}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ton-6-10.pdf}
+\includegraphics[width=0.5\textwidth]{figures/chapter-3/ton-6-10.pdf}
 \caption{Cutout of the TON framework showing a 6- (blue) and 10-MR (orange).
@@
 
-We can use ring counts to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites.
-These rings can be used to describe the channel sizes for understanding shape selectivity in catalysis, zeolite building blocks for solid state chemists, and the sizes of framework cages and windows to provide insights into adsorption properties.
-Since various conventions exist that can reduce the set of rings in a zeolite to more strictly defined properties, the ring counts returned by the various conventions will differ.
-Differences in ring counts leads to differences in how we might describe the topological environment of a zeolite.
-Therefore, when using rings to determine the properties of a framework, T-site, or oxygen atom, it is important to know the difference in the conventions, and use one that determines the features of interest.
-
-Here we present an analysis of rings captured by Goetzke and Klein's efficient ring finding algorithm \cite{goetzke-properties-1991}, and compare those rings to the rings found by other previously published ring set reduction conventions.
+Thus, the dimensions and numbers of rings present in a framework will differ depending on precise convention, with potential consequences for the ability to relate properties to rings.
+Here we present an analysis of rings captured by Goetzke and Klein's efficient ring finding algorithm \cite{goetzke-properties-1991} and compare those rings to the rings found by other previously published ring set reduction conventions.
 We have implemented all of these ring finding conventions in a Python package called the Zeolite Simulation Environment (ZSE) \cite{crum-jtcrumzse-2022}.
-We use ZSE to provide an analysis of rings captured by each convention for the entire set of zeolite frameworks contained on the IZA Database \cite{baerlocher-database-nodate} to compare how these sets of rings provide different characterizations of said frameworks.
+We use ZSE to analyze the sets of rings captured by each convention across the entire set of zeolite frameworks contained in the IZA Database \cite{baerlocher-database-nodate}.
 We  highlight rings that are found by these conventions but not typically discussed in the literature for a number of frameworks.
-We also show that the vertex symbol, a common approach used to characterize T-sites \cite{okeeffe-vertex-1997}, based on the shortest rings connecting the neighboring oxygen, misses important parts of the stereochemistry around a T-site.
-Finally, we provide an alternative method for listing the vertex symbol rings that takes into account their orientation and connectivity around the T-site..
+A comparison of the rings reported by various conventions to those listed in the IZA highlights differences related to fused rings illustrated in \cref{fig:3.3}}b.
+We propose a modification of the shortcut definition that results in a set of rings closer to those listed in the IZA database than do existing conventions.
+Lastly, we consider the vertex symbol \cite{okeeffe-vertex-1997} characterization of T-sites and show that T-sites described by the same vertex symbol are not necessarily superimposable but rather can have distinct local stereochemistries.
+We identify and enumerate these cases of constitutional isomerization within T-sites of identical vertex symbol and propose an alternative convention that uniquely takes into account their orientation and connectivity around the T-site.
 
@@
 
-\subsection{Finding Rings That do not Contain Shortcuts \label{section:goetzke}}
+\subsection{Rings That Do Not Contain Shortcuts \label{section:goetzke}}
 \label{sec:orgeb1d79d}
 
-In this work we implement the efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} in ZSE \cite{crum-jtcrumzse-2022} to find all the rings associated with a T-site that do not contain a shortcut.
+We implemented the efficient algorithm presented by Goetzke and Klein \cite{goetzke-properties-1991} in ZSE \cite{crum-jtcrumzse-2022} to find all the rings associated with a T-site that do not contain a shortcut.
 In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/goetzke.pdf}
+\includegraphics[width=0.5\textwidth]{figures/chapter-3/goetzke.pdf}
 \caption{Diagram showing how the ring finding algorithm evolves. Adapted from Goetzke and Klein \cite{goetzke-properties-1991}. \label{fig:goetzke}}
@@
 
-\subsection{Finding Vertex Symbol Rings \label{section:vertex}}
+\subsection{Vertex Symbol Rings \label{section:vertex}}
 \label{sec:org415adcd}
 
-Starting from the set of all rings found in \cref{section:goetzke}, we can prune the ring list to the set of vertex symbol rings.
+Starting from the set of all rings found in \cref{section:goetzke}, we prune the ring list to the set of vertex symbol rings.
 We find the shortest ring in the set that connects each pair of oxygens bound to our initial T-site.
@@
 
-\subsection{Finding Shortest Path Rings}
+\subsection{Shortest Path Rings}
 \label{sec:org6be3a64}
 
-Here we prune the set of all rings from \cref{section:goetzke} to a subset of rings the meets the shortest path definition published by Sastre and Corma \cite{sastre-topological-2009}.
-For each ring, we iterate over every group of O-T-O atoms in the ring, and check if this ring is the shortest path connecting the two oxygen atoms.
-If so, the loop is broke, because the ring need only be the shortest path for one group of O-T-O atoms to fit the definition.
+We prune the set of all rings from \cref{section:goetzke} to a subset of rings the meets the shortest path definition published by Sastre and Corma \cite{sastre-topological-2009}.
+For each ring, we iterate over every group of \ce{O-T-O} atoms in the ring, and check whether this ring is the shortest path connecting the two oxygen atoms.
+If so, the loop is broke, because the ring need only be the shortest path for one group of \ce{O-T-O} atoms to fit the definition.
 This is the most time consuming process out of all the ring finding conventions we have implemented.
@@
 \label{sec:org9232478}
+The shortest path convention generally finds fewer (and smaller) rings than does an all non-shortcut bearing ring convention.
+As we show below, these two conventions bracket the set of rings captured in the IZA Structure Database, the all-rings convention generally including even larger rings and the shortest path convention missing large rings.
+In seeking a convention of intermediate stringency, we identified a convention that would require the shortcut to be shorter than one, rather than both, of the paths connecting two nodes along the ring.
+This modified shortcut definition is illustrated in \cref{fig:3.3}(b).
+By the standard shortcut convention, AFI would be considered to contain a 14-MR formed from the union of Path 1 (blue) and Path 2 (purple) through T1 and T2.
+However, because T1 and T2 are also connected by Path 3, which, while the same length as Path 1, is shorter than Path 2, the 14-MR does not satisfy the modified shortcut criterion.
+As we show below, the modified shortcut rule identifies rings of a given zeolite that are closer to those identified in the IZA Structure Database as rings of interest than do either all-ring or shortest-path-ring rules.
 
-In this work we present a modified definition of shortcut to capture a different subset of rings from any of the other ring finding conventions.
-Traditionally a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into smaller rings \cite{guttman-ring-1990,goetzke-properties-1991}.
-We propose that a shortcut is a path connecting two nodes of a cycle that decomposes the cycle into at least one smaller ring.
-This definition does not require that the shortcut between two nodes be shorter than both paths connecting those nodes along the cycle.
-Our shortcut only needs to be shorter than one of the paths connecting those two nodes along the cycle.
-This definition has been described graphically in \cref{fig:3.3}(b).
-
-To remove rings containing modified shortcuts from the full set of rings, we iterate over every T-site pair of the ring and check for the shortest path connecting them.
-If that shortest path is shorter than either of the two paths along the ring connecting the two T-sites, we check if the combination of this shorter path and the shorter of the two ring paths forms a new smaller ring.
+Algorithmically, to remove rings containing modified shortcuts from the full set of rings, we iterate over every T-site pair of the ring and check for the shortest path connecting them.
+If that shortest path is shorter than either of the two paths along the ring connecting the two T-sites, we check whether the combination of this shorter path and the shorter of the two ring paths forms a new smaller ring.
 If so the iteration is broken, and the ring is removed from the counted set.
@@
 
-To complete the comparison of ring counting conventions, we have developed method to determine how similar the ring counts of one convention are to another.
+To complete the comparison of ring counting conventions, we have developed a method to determine the similarity between ring count  conventions.
 We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}.
@@
 
-One final point we would like to make about T-site characterization, is that the vertex symbol and ring indices only provide information about rings, and do not give any information about the spatial orientation of those rings around the T-site.
-The vertex symbol comes close to accomplishing that, by finding rings associated with opposite edge pairs of the tetrahedron, and then listing the those pairs from smallest to largest.
-However, the vertex symbol does not capture subtle but distinct differences in the orientation of the rings around the T-site that can lead to varying local void environments.
-For example: MOR T3, MON T1, and EON T9 all have the same vertex symbol of: 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\).
-However, the orientation of those rings around each of those T-sites are not identical.
-\cref{fig:stereo} shows a cutout of each of these frameworks that only includes the atoms that make up the rings of the vertex symbol around the specified T-sites.
-We can see that MOR T3 and EON T9 have the same ring orientation, and that orientation is different from the rings making up the vertex symbol of MON T1.
-The main difference is the location of the 5\(_{\text{2}}\)- and 4-MR edges.
-They are highlighted in \cref{fig:stereo}.
+The conventional definition of a vertex symbol, which lists pairs of rings along opposite edges of the tetrahedron from smallest to largest, only partially captures  the stereochemistry around a T-site.
+As example, \cref{fig:stereo} illustrates the rings associated  with T3, T9, and T1 or MOR, MON, and EON, respectively, all of which share the same 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\) vertex symbol.
+While MOR T3 and EON T9 are similar, MON T1 differs in the location of the 5\(_{\text{2}}\)- and 4-MR edges, as highlighted in the lower panel of \cref{fig:stereo}.
+To remove this ambiguity, an alternative approach is to \red{need to explain the basis of the modified procedure in words here.
+2.5 describes the algorithm but not its physical basis.} Within this algorithm MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
+The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol.
+We computed the ordered vertex symbols for  all 1460 T-sites in the IZA Database.
+The 649 unique vertex symbols in the database increases to 666 unique ordered vertex symbols.
+Thus, the consideration of stereochemistry only modestly increases the space of unique  tetrahedral sites.
+\cref{tab:ov}  lists some common T-site vertex symbols and associated ordered vertex symbols.
+\red{Complete results are presented in the Supplementary?}
 
@@
 
-The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
+% The structural differences shown in \cref{fig:stereo} that are not able to be captured by the vertex symbol leads us to believe that the vertex symbol is not a complete descriptor, and there is room to define a new descriptor that takes into consideration ring orientation.
 While a vertex symbol includes the shortest rings connecting the four oxygen atoms adjacent to a T-site, and listing them by size and opposite edges, there is still ambiguity in the exact connectivity of those rings.
 This has led us to create a new method for listing the rings in the vertex symbol that considers the structural connection of the rings described in \cref{section:ov}.
-
-With this new descriptor MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
-The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol.
-Stereochemistry of the rings associated with a T-site could influence the chemical properties we care about such as deprotonation energy, T-site substitution energy, or catalytic properties of reactions happening at that T-site.
-
-We have used this new ordered vertex symbol to characterize all 1460 T-sites in the IZA Database.
-We found that using the standard vertex symbol to characterize T-sites there are 649 unique vertex symbols present.
-In contrast, there are 666 unique ordered vertex symbols across every T-site.
-This would imply that not a large amount of vertex symbols contain stereochemical differences.
-In \cref{tab:ov} we provide a list of some common T-site vertex symbols, and their representative ordered vertex symbols.
 

### 2023-01-04 elsman/manuscript.tex
updated text for reviewer responses

@@
 This path contains two T-sites, splitting the 8-MR cycle into two 4-MRs.
-While general, this definition reports as distinct rings features of a zeolite framework that do not intuitively relate to local structure or accessibility or usefully ``fingerprint'' frameworks.
+While general, this definition reports, as distinct, ring features of a zeolite framework that do not intuitively relate to local structure, accessibility, or usefully ``fingerprint'' frameworks.
 Thus, more restrictive defintions of zeolite rings have been proposed.
-For instance, a ``strong ring'' is defined as a ring that is not the sum of a set of smaller rings.
-\red{A little unclear.
-How does this relate to shortcuts?} \cite{goetzke-properties-1991,blatov-three-periodic-2007}.
+For instance, a ``strong ring'' is defined as a ring that is not the sum of a set of smaller rings \cite{goetzke-properties-1991,blatov-three-periodic-2007}.
 
- Another pruning method is based off of the concept of a vertex symbol.
-The vertex symbol is a common way to describe the structure around the T-sites of zeolite frameworks and were first used to do so in 1997 by O'Keeffe and Hyde \cite{okeeffe-vertex-1997}.
-The vertex symbol of a T-site is determined by the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bounded oxygen atoms.
+ Another method for pruning the list of rings found in a zeolite frameworks is based off of the concept of a vertex symbol.
+The vertex symbol is a common way to describe the structure around the T-sites of zeolite frameworks and was first used to do so in 1997 by O'Keeffe and Hyde \cite{okeeffe-vertex-1997}.
+The vertex symbol of a T-site is determined by the shortest rings associated with each of the six edges of the tetrahedron formed by the T-atom and its four bound oxygen atoms.
 The symbol is formed by grouping rings on opposite edges of the tetrahedron and listing from smallest to largest.
@@
 
-Sastre and Corma \cite{sastre-topological-2009} proposed an alternative convetion, in which distinct rings are only those that form the shortest path connecting any \ce{O-T-O} in a framework \red{do you mean cycle?}.
+Sastre and Corma \cite{sastre-topological-2009} proposed an alternative convetion, in which distinct rings are only those that form the shortest cycle connecting any \ce{O-T-O} in a framework.
 This convention provides an univocal count of the rings that pass through a T-site without the ambiguity of limiting the search to a certain ring size.
@@
 For T1, only the 6-MR would be counted within the vertex symbol convention because it is the shortest path connecting O2 and O3.
-The 10-MR is part of the vertex symbol for T3 because it is the shortest path connecting O2 and O14.
-Because this 10-MR is the shortest path for at least one set of \ce{O-T-O} along the ring, and passes through T1, it does get counted in the shortest path rings for T1.
-\red{The end of this paragraph is unclear as to whether the Vertex symbol or Sastre definition is being used.
-Please reread and clarify.}
+The 10-MR is the shortest path connecting O2, and since this 10-MR is the shortest path for at least one set of \ce{O-T-O} along the cycle, and passes through T1, it does get counted in the shortest path rings for T1.
 
@@
 In ZSE we use the framework put in place by the Atomic Simulation Environment (ASE) \cite{hjorth-larsen-atomic-2017} to handle routine analysis zeolite crystal structures.
-All graph theory functions are performed used the NetworkX Python package \cite{hagberg-exploring-2008}.
+All graph theory functions are performed using the NetworkX Python package \cite{hagberg-exploring-2008}.
 
@@
 This connectivity matrix is then converted to a NetworkX graph object, and then a distance dictionary using NetworkX built in functions.
-Then we implement Step 3 from Geotzke and Klein's algorithm \cite{goetzke-properties-1991} summarized here: to find the rings that pass through a T-site, we iteratively search for every size ring between 3-MR and a maximum ring value that is user specified.
+Then we implement Step 3 from Geotzke and Klein's algorithm \cite{goetzke-properties-1991} summarized here: to find the rings that pass through a T-site, we iteratively search for every size ring between 3-MR and a maximum ring value that is user-specified.
 For this work we set a cutoff of 18-MRs.
 A schematic showing the evolution of the ring search is shown in \cref{fig:goetzke}.
-For ring size \(\lambda\) we start at the T-site of interest (labeled 1) \cref{fig:goetzke}, and search the distance matrix for any T-sites that are \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (for odd \(\lambda\)) distance from the starting T-site (labeled 2).
-Next we attempt to create to distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by \cref{fig:goetzke}.
+For ring size \(\lambda\) we start at the T-site of interest (labeled 1), and search the distance matrix for any T-sites that are \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (for odd \(\lambda\)) distance from the starting T-site (labeled 2).
+Next we attempt to create two distinct paths from 1 \(\rightarrow\) 2 and from 2 \(\rightarrow\) 1 alternating adding a node to each path as indicated by \cref{fig:goetzke}.
 Each node added to each of the paths must be \(\lambda\)/2 (even \(\lambda\)) or (\(\lambda\)-1)/2 (odd \(\lambda\)) from the head of the other path.
@@
 The rings of a framework are made of T-sites, and if two T-sites are symmetrically identical they will have the same set of rings passing through them.
-Therefore, we only need to find the rings associated with each symmetry distinct T-sites to know all the possible ring sizes within a framework.
+Therefore, we only need to find the rings associated with each symmetry distinct T-site to know all the possible ring sizes within a framework.
 For example, AFI only contains one symmetry distinct T-site (T1).
@@
 There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.
+We have limited our search to 18-MR because the differences in ring finding conventions are captured with the smaller rings, requiring less computation time, and there is a lower probability of finding larger rings in zeolites \cite{li-why-2014}.
 
@@
 
-Taking a closer look at some of these untabulated rings, highlights rings not typically listed for some frameworks, but still relevant to describing their topology.
-Using CHA as an example, \cref{fig:cha-rings} displays a 12-MR (in purple) that exists in CHA that circumferences the cage.
+Taking a closer look at some of these untabulated rings highlights rings not typically listed for some frameworks, but are still relevant to describing their topology.
+Using CHA as an example, \cref{fig:cha-rings} displays a 12-MR (in purple) that circumferences the CHA cage.
 This ring is not associated with the vertex symbol of the single symmetry distinct T-site in CHA and does not define a channel.
 Thus, this ring is not included in the list of tabulated rings.
-We would argue that this is still a ring that provides an important topological descriptor of CHA because none of the tabulated rings provides information about the size of the CHA cage.
+We would argue that this is still a ring that provides an important topological descriptor of CHA because none of the tabulated rings provide information about the size of the CHA cage.
 
@@
 When we search for rings using the Goetzke algorithm \cite{goetzke-properties-1991}, we also find that it contains 14-MRs created by using seven T-sites from two 12-MRs that are separated by a distance of one oxygen (\cref{fig:afi-14}).
-Rings of this nature are prevalent in many frameworks, another example can be seen in the bottom right of \cref{fig:cha-rings}, where an 8-MR is highlighted traversing the two 6-MRs of the D6R.
+Rings of this nature are prevalent in many frameworks; another example can be seen in the bottom right of \cref{fig:cha-rings}, where an 8-MR is highlighted traversing the two 6-MRs of the D6R.
 These types of rings may not be of interest depending on which topological feature one intends to describe.
 This has led us to create a modified definition of a shortcut as explained in \cref{section:modified}, which excludes these types of rings.
-The benefit of this new shortcut definition is that larger rings that are missed by the vertex symbol or shortest path rings (i.e., 12-MR in AFI) are still captured, while excluding rings that arise from convolution of staked rings found with the typical shortcut definition.
+The benefit of this new shortcut definition is that larger rings that are missed by the vertex symbol or shortest path rings (i.e., 12-MR in AFI) are still captured, while excluding rings that arise from convolution of stacked rings found with the typical shortcut definition.
 
@@
 \label{sec:org3c3c2d3}
-With the addition of our modified shortcut definition, we have four ring finding conventions to compare, as well as including the tabulated rings from the IZA Database.
+With the addition of our modified shortcut definition, we can compare the results of four ring finding conventions to the tabulated rings listed in the IZA Database.
 \cref{fig:ring-counts} shows how many frameworks contain each size ring found using the various ring counting conventions from 3- to 18-MRs.
@@
 In general, a hierarchy of ring sizes found by each convention is: all rings not containing a shortcut \textgreater this work \textgreater shortest path rings \textgreater vertex symbol rings.
-While the IZA listed rings includes the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.
+The IZA listed rings include the vertex symbol rings, and a selection of general rings \cite{baerlocher-database-nodate}.
 
@@
 These shortcut containing cycles can display chemical and/or geometric properties consistent with rings and are of interest to catalysis researchers even though they are not classically considered rings.
-One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}, which is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}.
-Similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}.
+One example is the 6-membered cycle referred to as the \(\alpha\)-6-MR in literature (\cref{fig:mfi-6}) and is present in a number of frameworks including but not limited to MOR, FER, MFI, and BEA \cite{dedecek-siting-2012,bernauer-proton-2016}.
+This \(\alpha\)-6-MR is a potential location for Co\(^{\text{2+}}\) uptake when two Al atoms are 3rd nearest neighbor (NN) in the cycle  \cite{bernauer-proton-2016,nimlos-experimental-2020}, similar to Co\(^{\text{2+}}\) uptake at 3NN Al atoms in 6-MRs in other frameworks such as CHA \cite{di-iorio-cooperative-2020}.
 \cref{fig:mfi-6} shows that this particular structure would be considered two 5-MRs using connectivity rules based on a shortcut.
@@
 In their work, they presented the ring index, which lists all the rings passing through a T-site from smallest to largest, and a subscript for each size representing it's multiplicity.
-The rings associated with a T-site can provided information about the local void environments around the T-site, and could potentially be correlated to other physicochemical properties of the T-site once enough once sufficient data on those physicochemical properties exists.
+The rings associated with a T-site can provide information about the local void environments around the T-site, and could potentially be correlated to other physicochemical properties of the T-site once sufficient data on those physicochemical properties exists.
 
@@
 We can also prune this list using our modified definition of a shortcut, the shortest path rings definition \cite{sastre-topological-2009}, or the rings contained within the vertex symbol of this T-site \cite{okeeffe-vertex-1997}.
-Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}:
+Using the ring index outlined above, each of these conventions will provide a different description of the zeolite (highlighted in \cref{fig:afi-funnel}):
 \begin{itemize}
@@
 
-With an understanding of how we characterize T-site by counting the rings that pass through, \cref{tab:uninodal} shows the ring index for a selection of T-sites from uninodal (containing only one symmetry distinct T-site) frameworks.
-This table highlights the differences in rings counts found with each convention, and shows that in general as you move from left to right across the table the largest ring found decreases.
+With an understanding of how we characterize T-sites by counting the rings that pass through them, \cref{tab:uninodal} shows the ring index for a selection of T-sites from uninodal (containing only one symmetry distinct T-site) frameworks.
+This table highlights the differences in ring counts found with each convention, and shows that in general as you move from left to right across the table the largest ring size found decreases.
 The results in the shortest path column were found using ZSE, but agree directly with the results shown by Sastre and Corma \cite{sastre-topological-2009}.
-The results in the vertex symbol rings column were also found with zse, and agree directly with the vertex symbols listed on the IZA Database website \cite{baerlocher-database-nodate}.
+The results in the vertex symbol rings column were also found with ZSE, and agree directly with the vertex symbols listed on the IZA Database website \cite{baerlocher-database-nodate}.
 
@@
 If you were interested in which T-sites have access to the 12-MR channels, the shortest path rings and vertex symbol rings would only suggest T3 participates in the 12-MR rings.
-However, all rings and this work both identify T4 and T6 as participating in the 12-MR channels as highlighted in \cref{fig:moz}.
+However, the all rings convention and the modified shortcut convention both identify T4 and T6 as participating in the 12-MR channels as highlighted in \cref{fig:moz}.
 
@@
 T1: navy, T2: green, T3: orange, T4: purple, T5: blue, and T6: red.
-As shown, T3, T4, and T6 are all associted with the 12-MR channels, while T1, T2, and T5 are not connected to the 12-MR channels.
+As shown, T3, T4, and T6 are all associated with the 12-MR channels, while T1, T2, and T5 are not connected to the 12-MR channels.
 \label{fig:moz}}
@@
 For each T-site we used the rings to generate a ring index, and \cref{fig:unique-ts} shows how many unique ring indices are present when using each of the ring counting conventions.
-The plot follows intuition with the number of unique ring indices decreasing as we use more restrictive ring counting conventions, because less rings are found and provides less room for differentiation.
-This raises the question, if you want to ascertain chemical or physical properties about a T-site based on its ring count, and differentiate these T-sites from other similar but distinct T-site, which ring counting convention will suffice?
+The plot follows intuition with the number of unique ring indices decreasing as we use more restrictive ring counting conventions, because fewer rings are found providing less room for differentiation.
+This raises the question, if you want to ascertain chemical or physical properties about a T-site based on its ring count, and differentiate these T-sites from other similar but distinct T-sites, which ring counting convention will suffice?
 The answer will depend on what level of detail is desired.
-Larger rings can be found with the standard shortcut definition, rings traversing other stacked rings will be excluded with our modified shortcut definition.
+Larger rings can be found with the standard shortcut definition, while rings traversing other stacked rings will be excluded with our modified shortcut definition.
 The shortest path definition and vertex symbol rings will provide the most localized information about a T-site.
@@
 This plot highlights that more T-sites contain larger sized rings when using the basic definition of a shortcut, and at smaller rings sizes (\textless6-MR) all the ring counting conventions return the same results.
-To further emphasize this results, we have provided a cumulative distribution of the same data normalized to the maximum 'rings' value in \cref{fig:tsite-frequency}.
+To further emphasize these results, we have provided a cumulative distribution of the same data normalized to the maximum 'rings' value in \cref{fig:tsite-frequency}.
 At 6-MRs is where we see the cumulative distribution functions deviate from each other.
@@
 
-To complete the comparison of ring counting conventions, we have developed a method to determine the similarity between ring count  conventions.
+To complete the comparison of ring counting conventions, we have developed a method to determine their similarity.
 We do this by comparing the ring index for a T-site using each convention, where the similarity of the ring indices is scored with \cref{eq:similarity}.
@@
 Down the diagonal each method is compared to itself and clearly has a similarity of 1.
-The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compare to the least restrictive convention (rings) has the lowest similarity score.
+The remainder of the table follows intuition, in that the  most restrictive ring counting convention (vertex symbol rings) compared to the least restrictive convention (rings) has the lowest similarity score.
 The two most similar ring counting methods are our modified shortcut definition and the shortest path rings.
@@
 The conventional definition of a vertex symbol, which lists pairs of rings along opposite edges of the tetrahedron from smallest to largest, only partially captures  the stereochemistry around a T-site.
-As example, \cref{fig:stereo} illustrates the rings associated  with T3, T9, and T1 or MOR, MON, and EON, respectively, all of which share the same 4\codt5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\) vertex symbol.
+As an example, \cref{fig:stereo} illustrates the rings associated  with T3, T9, and T1 or MOR, MON, and EON, respectively, all of which share the same 4\(\cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}} \cdot\)5\(\cdot\)8\(_{\text{2}}\) vertex symbol.
 While MOR T3 and EON T9 are similar, MON T1 differs in the location of the 5\(_{\text{2}}\)- and 4-MR edges, as highlighted in the lower panel of \cref{fig:stereo}.
-To remove this ambiguity, an alternative approach is to \red{need to explain the basis of the modified procedure in words here.
-2.5 describes the algorithm but not its physical basis.} Within this algorithm MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
+To remove this ambiguity, an alternative approach is to list the rings of the vertex symbol by starting with the largest one and systematically adding the remaining rings following the edges of the tetrahedron such that each ring listed is connected to one of the oxygen atoms of the next ring listed.
+This provides a physical meaning to the order of the vertex symbol rings which accounts for differences in the connectivity of those rings around the T-site.
+Within this algorithm MOR T3 and EON T9 would be labeled as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)5\(_{\text{2}} \cdot\)5\(\cdot\)4\(\cdot\)5, and MON T1 as: 8\(_{\text{2}} \cdot\)8\(_{\text{2}} \cdot\)4\(\cdot\)5\(\cdot\)5\(_{\text{2}} \cdot\)5.
 The difference is subtle but highlights the distinct structural difference between the two types of T-sites that is not otherwise captured by a vertex symbol.
 We computed the ordered vertex symbols for  all 1460 T-sites in the IZA Database.
-The 649 unique vertex symbols in the database increases to 666 unique ordered vertex symbols.
+The 649 unique vertex symbols in the database increase to 666 unique ordered vertex symbols.
 Thus, the consideration of stereochemistry only modestly increases the space of unique  tetrahedral sites.
 \cref{tab:ov}  lists some common T-site vertex symbols and associated ordered vertex symbols.
-\red{Complete results are presented in the Supplementary?}
+Complete results are presented in the Supplementary Information.
 
@@
 Rings of a graph are well defined; here we identified all rings up to 18-MR in every zeolite framework listed on the IZA Structure Database \cite{baerlocher-database-nodate}.
-We find that the commonly reported ring sizes in literature and on the IZA website leave out many rings that fit the classical definition of a cycle that does not contain a shortcut.
+We find that the commonly reported ring sizes in literature and on the IZA website leave out many rings that fit the classical definition of a cycle that do not contain a shortcut.
 To completely describe the topology of a zeolite these rings are required, however there are often cases where someone might want to consider only a subset of rings of interest.
 
-We have shown a comparison of three different conventions used to count rings, and highlighted the differences in rings that are found by each convention.
+We have shown a comparison of three existing conventions used to count rings, and highlighted the differences in rings that are found by each convention.
 The classic definition of a ring identifies the largest set of ring sizes across all the zeolite frameworks, while the shortest path ring and vertex symbol rings only identify smaller ring sizes.
-We have provided a modified definition of a shortcut that when used to find rings still finds larger rings defining channel openings, but excludes rings that are able to be decomposed into at least one smaller ring.
+We have provided a modified definition of a shortcut that finds larger rings defining channel openings, but excludes rings that are able to be decomposed into at least one smaller ring.
 It is important to understand the difference of ring sizes and types found by each convention when discussing the rings of a zeolite framework.
@@
 
-This same methodology was used to describe T-sites in zeolite frameworks by counting all of the rings that pass through the T-site using each of the ring counting conventions described.
+We describe T-sites in zeolite frameworks by counting all of the rings that pass through the T-site using each of the ring counting conventions described.
 When using ring finding conventions that find larger rings, we see more diversity in the descriptions of T-sites, which can aid researchers who want to identify similar T-sites across multiple frameworks.
@@
 In the future, correlating physicochemical properties of T-sites to the ring descriptors identified with each ring counting convention can help identify sets of frameworks with desired T-site properties.
+
+\section{Data Statement}
+Complete ring finding results are available for download via the Supplementary Information.
+This folder contains a file for each ring counting convention: all rings, modified shortcut rings, shortest path rings, vertex symbol rings, and ordered vertex symbol rings.
+These files contain the rings associated with each oxygen atom and T-site in every zeolite framework listed on the IZA Structure Database.
 

### 2023-01-04 elsman/manuscript.tex
WFS edits

@@
 \item Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12\(\cdot\)14\(_{\text{7}}\)
-\item This work: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
+\item Modified Shortcut Rings: 4\(\cdot\)6\(_{\text{13}} \cdot\)12
 \item Shortest Path Rings: 4\(\cdot\)6\(_{\text{13}}\)
@@
 \includegraphics[width=\textwidth]{figures/chapter-3/afi-funnel.pdf}
-\caption{Diagram shwoing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
+\caption{Diagram showing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
 \end{figure}

### 2023-01-04 elsman/manuscript.tex
updated response for R2 Comment 10

@@
 The dimension and number of rings in a framework can be used to characterize entire zeolite frameworks, T-sites that make up these frameworks, or even the oxygen atoms that connect the T-sites.
+Take for example the CHA framework, \cref{fig:3.1} shows an 8-membered ring (8-MR) and a 4-MR filled in with pink and green respectively.
+As an illustration of the significance of those rings, consider that the T-site labeled 20 is a Br\o nsted acid site, i.e., the site contains an Al and a charge-compensating proton bound to any one of the four adjacent oxygen atoms.
+While the T-site is symmetrically identical to all other T-sites in CHA, the proton sites are inequivalent, distinguished by the dimensions of the rings that pass through the oxygen atoms, a distinction that is experimentally observable in IR spectra \cite{kester-effects-2021}.
 The number and dimension  of rings that pass through a T-site can be used to describe and differentiate the local void environment around symmetry distinct T-sites and to  differentiate symmetry distinct oxygen atoms.
@@
 A zeolite crystal contains an infinite number of cycles, but the symmetry of the crystal reduces this to a finite number of symmetry-distrinct rings whose exact identity is subject to definition.
-\cref{fig:3.1} shows an 8-membered ring (8-MR) and a 4-MR filled in pink and green, respectively.
+\cref{fig:3.1} shows an 8-MR and a 4-MR filled in pink and green, respectively.
 

### 2023-01-11 elsman/manuscript.tex
revised submitted

--- 2023-01-04 elsman/manuscript.tex
@@
 
-\begin{figure}
-\begin{figure}[H]
+\cref{fig:3.1} shows a cutout from the CHA framework, highlighting an 8-membered ring (8-MR) and a 4-MR filled in pink and green, respectively.
+As an illustration of the significance of those rings for a T-site, consider that the T-site labeled 20 is a Br\o nsted acid site, i.e., the site contains an Al.
+It will then have associated with it a charge-compensating proton that will bind to one of the four adjacent oxygen atoms.
+While the T-site is symmetrically identical to all other T-sites in CHA, the proton sites are inequivalent, distinguished by the dimensions of the rings that pass through them.
+These differences lead to differences in proton energies and vibrational frequenices, a distinction that is experimentally observable in infrared spectra \cite{kester-effects-2021}.
+
+\begin{figure}[t]
 \centering
@@
 \caption{Cutout of the Chabazite framework showing a path (5-6-7-8-9) highlighted with purple bonds, a cycle (3-4-18-19-20-17) highlighted with blue bonds, an 8-MR filled in with pink, and a 4-MR filled in with green.
-Yellow atoms are Si (T-sites), and red atoms are oxygen.
+Yellow and red spheres are T-sites (typically Si) and oxygen, respectively.
 \label{fig:3.1}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
@@
 \label{fig:cha-vertex}}
-\end{figure}
 \end{figure}
@@
 \end{figure}
-\end{figure}
 
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/chapter-3/ton-6-10.pdf}
+\includegraphics[width=0.4\textwidth]{figures/chapter-3/ton-6-10.pdf}
 \caption{Cutout of the TON framework showing a 6- (blue) and 10-MR (orange).
@@
 \label{fig:3.4}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.5\textwidth]{figures/chapter-3/goetzke.pdf}
+\includegraphics[width=0.4\textwidth]{figures/chapter-3/goetzke.pdf}
 \caption{Diagram showing how the ring finding algorithm evolves. Adapted from Goetzke and Klein \cite{goetzke-properties-1991}. \label{fig:goetzke}}
-\end{figure}
 \end{figure}
@@
 In some frameworks, other rings (cycles not containing shortcuts) exist that are not included in the list of tabulated rings.
-These 'untabulated' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site.
+These ``untabulated'' rings may still provide important topological information about a zeolite framework, or the local void environment around a T-site.
 \cref{fig:fw-counts} shows counts of frameworks containing each size ring from 3- to 18-MR using the Goetzke algorithm and the listed rings on the IZA database \cite{baerlocher-database-nodate}.
-There are slight differences in the counts up to 6-MRs, but the main divergence takes place as we get to ring sizes \textgreater6-MR.
-We have limited our search to 18-MR because the differences in ring finding conventions are captured with the smaller rings, requiring less computation time, and there is a lower probability of finding larger rings in zeolites \cite{li-why-2014}.
+There are slight differences in the counts up to 6-MRs, but the main divergence takes place at ring sizes greater than six.
+We limit the search to 18-MR and smaller because the differences in ring finding conventions are captured within the smaller ring sets, because probabilities of occurrence become small with increasing ring size \cite{li-why-2014}, and for computational expediency.
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t']
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/rings-vs-iza-rings.pdf}
+\includegraphics[width=0.7\textwidth]{figures/chapter-3/rings-vs-iza-rings.pdf}
 \caption{Counts of IZA frameworks containing each size ring between 3- and 18-MR using the Goetzke algorithm and the tabulated rings listed by the IZA \cite{baerlocher-database-nodate}.
 \label{fig:fw-counts}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/cha-all-rings.pdf}
+\includegraphics[width=0.6\textwidth]{figures/chapter-3/cha-all-rings.pdf}
 \caption{Chabazite cage and double 6-MR (D6R) with highlighted rings: 4-MR in green, 8-MR in pink, and 12-MR in purple.
@@
 \label{fig:cha-rings}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/afi-14.pdf}
+\includegraphics[width=0.5\textwidth]{figures/chapter-3/afi-14.pdf}
 \caption{Cutout of the 12-MR channel in AFI with a 14-MR (yellow) traversing seven T-sites of each 12-MR.
@@
 \label{fig:afi-14}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/ring-counts-2.pdf}
+\includegraphics[width=0.7\textwidth]{figures/chapter-3/ring-counts-2.pdf}
 \caption{Number of IZA zeolite frameworks containing each size ring, using the various ring counting conventions, as well as the rings listed by the IZA Database \cite{baerlocher-database-nodate}.
 \label{fig:ring-counts}}
-\end{figure}
 \end{figure}
@@
 \label{fig:mfi-6}}
-\end{figure}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{figures/chapter-3/afi-funnel.pdf}
-\caption{Diagram showing the ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
+\caption{Ring counts of each size ring that pass through the single symmetry distinct T-site in AFI for each of the various ring finding conventions. \label{fig:afi-funnel}}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/moz.pdf}
+\includegraphics[width=0.8\textwidth]{figures/chapter-3/moz.pdf}
 \caption{Cutout of the MOZ framework showing two 12-MR channels, with an example of each distinct T-site highlighted.
@@
 \label{fig:moz}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/unique-ts.pdf}
+\includegraphics[width=0.7\textwidth]{figures/chapter-3/unique-ts.pdf}
 \caption{Number of unique T-site ring indices when classified by the rings passing through them using the various ring counting conventions.
@@
 \label{fig:unique-ts}}
-\end{figure}
 \end{figure}
@@
 \label{fig:tsite-frequency}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
-\begin{figure}[H]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{figures/chapter-3/similarity-heat-map.pdf}
+\includegraphics[width=0.9\textwidth]{figures/chapter-3/similarity-heat-map.pdf}
 \caption{Heat map showing the similarity score for four ring counting methods.
@@
 \label{fig:similarity}}
-\end{figure}
 \end{figure}
@@
 
-\begin{figure}
 \begin{figure}[H]
@@
 \label{fig:stereo}}
-\end{figure}
 \end{figure}
@@
 Schmitt Foundation for financial aid in the form of a PhD fellowship.
-We thank Dr.
-Christian Baerlocher for the numerous discussions about the topology of zeolites, and methodologies used by the IZA Structure Database.
-We thank Dr.
-German Sastre for providing a copy of zeoTsites to compare results, and for the helpful conversions about use cases and zeolite topology.
+We thank Dr.\ Christian Baerlocher for the numerous discussions about the topology of zeolites, and methodologies used by the IZA Structure Database.
+We thank Dr.\ German Sastre for providing a copy of zeoTsites to compare results, and for the helpful conversions about use cases and zeolite topology.
 This research was supported in part by the Notre Dame center for Research Computing through access to high performance computing clusters.
